---
title: "Landbird Data Processing Workflow"
name: Hannah
date: "`r format(Sys.time(), '%B %Y')`"
output:
  html_notebook:
    toc: yes
    number_sections: yes
  html_document:
    toc: yes
    df_print: paged
address: Alaska Regional Office 1011 E Tudor Rd, Anchorage, AK
email: hannah_vincelette@fws.gov
github: "https://github.com/hdvincelette/landbird-dp-workflow"
surname: Vincelette
position: Wildlife Biologist
editor_options:
  chunk_output_type: console
---

# Overview

The purpose of this document is to provide a reproducible workflow for processing data collected annually by the Landbird section of USFWS Alaska Migratory Bird Management. Data processing is just one component of data management. For complete guidance, view the [Alaska Region Interim Data Management User Guide.](https://ak-region-dst.gitbook.io/alaska-region-interim-data-management-user-guide/)

#### What is an R Notebook?

["An R Notebook is an R Markdown document with chunks that can be executed independently and interactively, with output visible immediately beneath the input." - R Markdown: The Definitive Guide](https://bookdown.org/yihui/rmarkdown/notebook.html)

#### How to use this guide

This document is meant to be interactive, with text prompts and "code chunks" which may require input before being executed. The following workflow was created for tabular data collected annually for projects archived on the Regional Data Repository. While effort is made to generalize code to all actively collected data, further edits may be required to process data in an standardized format. In this case, it is recommended users duplicate and rename the "landbird-dp-workflow.Rmd" before making changes.

# Workspace setup

### Load packages

Run the following installation code.

```{r include=FALSE, results=TRUE}

# Clear environment
rm(list=ls())

# Install CRAN packages
if (!require("pacman"))
  install.packages("pacman")
pacman::p_load("devtools", "move", "auk", "magrittr", "shiny","bslib")

# Load project
setwd(dirname(rstudioapi::documentPath()))
rstudioapi::filesPaneNavigate(getwd())
devtools::load_all()

theme.selection<- shinythemes::shinytheme("simplex")

# # Install GitHub packages
# ## Uninstall packages (if re-installing)
# remove.packages("FWSAkRDRtools")
# remove.packages("mdJSONdictio")
# .rs.restartR()
# 
# ## Set GitHub download file method (optional)
# options(download.file.method = "wininet")
# 
# ## Install FWSAkRDRtools from GitHub
# devtools::install_github("hdvincelette/FWSAkRDRtools")
# 
# ## Install mdJSONdictio from GitHub
# devtools::install_github("hdvincelette/mdJSONdictio")


```

If the installations fail, refer to [Install mdJSONdictio R package](https://hdvincelette.github.io/mdJSONdictio/articles/03_Setup_mdJSONdictio.html#install-mdjsondictio-r-package) for troubleshooting tips.

### Create subfolders

Run the following installation code to install necessary subfolders

```{r}

fs::dir_create(names(which(!fs::dir_exists(
  c(
    "downloads/dictionaries",
    "downloads/data",
    "downloads/templates",
    "data-raw",
    "data-processed/bander_portal",
    "data-processed/movebank",
    "data-processed/to_archive"
  )
))))

```

# Data preprocessing

### Download authoritative files

Open the [Migratory Bird Management Regional Data Repository](file://ifw7ro-file.fws.doi.net/datamgt/mbm/) and navigate to the project folder. If you are working from a remote location, you must be connected to the VPN (Virtual Private Network) to access this folder.

Paste the names of the project folder, data dictionary(ies), and data entry template(s).

```{r include=FALSE}

# Provide the relevent information
RDR.keyword <- "Eielson"

data.keyword <- "survey"
data.subfolder <- "data/final_data"

template.keyword <- NA
template.subfolder <- NA

dictionary.keyword <- "mdeditor"
dictionary.subfolder <- "metadata"

# Provide a directory path to save files
data.download.path <- "downloads/data"
template.download.path <- "downloads/template"
dxnry.download.path <- "downloads/dictionaries"

```

Now run the following code chunk.

```{r include=FALSE, results=FALSE}

if(dir.exists(paste0("//ifw7ro-file.fws.doi.net/datamgt/mbm/"))==FALSE) {
  stop("Check the VPN connection and try again.")
  
}

project.folder <- dir(
  path = paste0("//ifw7ro-file.fws.doi.net/datamgt/mbm/"),
  pattern = paste0("^mbmlb.*", RDR.keyword),
  full.names = FALSE,
  recursive = FALSE,
  ignore.case = TRUE
)

if(length(project.folder)>1) {
  folder.choice <- utils::menu(
    c(project.folder),
    graphics = TRUE,
    title = cat(
      paste0(
        "\nMore than one project folder was found.\nSelect one to access.\n"
      )
    )
  )
}

keywords <- c(data.keyword, 
              template.keyword, 
              dictionary.keyword)
subfolders <- c(data.subfolder, 
                template.subfolder, 
                dictionary.subfolder)
download.paths <- c(data.download.path,
                    template.download.path,
                    dxnry.download.path)

for(x in seq_along(keywords)) {
  if (length(keywords[x]) != 0) {
    if (is.na(keywords[x])==FALSE) {
      
      FWSAkRDRtools::download.files(
        pattern = keywords[x],
        project = project.folder,
        subfolder.path = subfolders[x],
        local.path = download.paths[x],
        main = TRUE,
        incoming = TRUE,
        recursive = TRUE
      )
      
    }
  }
}


utils::browseURL(download.path)

```

View and verify the downloaded data dictionary(ies) and data entry template(s) (the default location is the "downloads" sub-folder of this R project).

*Optional: You can skip the above steps and download everything manually from the [Migratory Bird Management Regional Data Repository](file://ifw7ro-file.fws.doi.net/datamgt/mbm/) project folder. Data dictionaries and data entry templates should be available in the "metadata" sub-folder. Save a copy of data dictionary(ies) and data entry template(s) in an accessible location (i.e. the "downloads" sub-folder of this R project)*

### Enter data

Copy and save each data entry template in a preferred location for data entry (this can be in the "data-raw" sub-folder of this R project). Use a unique and descriptive name - refer to [Best Practices in Naming Conventions](https://ak-region-dst.gitbook.io/alaska-region-interim-data-management-user-guide/alaska-data-management-101/file-organization-and-best-practices/best-practices-in-naming-conventions). It's okay to save this file in an Excel file format to preserve validation rules.

Enter scanned or hard copy data according to the respective data dictionary(ies). Review entered data to check for errors. The Excel Filter function (Home\>Editing\>Filter) is especially useful for viewing unique values and value ranges.

Save the entered data file(s) in the "data-raw" sub-folder of this R project. Convert the file(s) to comma-separated values (CSV) format, if necessary. Make sure the correct sheet is selected when converting if more than one sheet is used (i.e., validation rules).

# Data quality checks

Before beginning, make sure all entered data and data dictionary(ies) are saved in the sub-folders of this R project ("data-raw" and "downloads", respectively).

### Provide inputs

```{r}
# Provide directory paths for the data file(s) and dictionary
raw.data.path <- "data-raw/"
dictionary.path <- "downloads/dictionaries"

# Provide the relevant information (if applicable)
missing_values<- c(NA)
species.col <- "SpeciesCode"
date.col <- "Date"
date.format <-  "%m/%d/%Y"
time.col <- "CaptureTime_2400"
time.format <- "%H:%M"

```

### Import

```{r}

raw.data.list <- importer(
  path = raw.data.path,
  multiple = TRUE,
  graphics = TRUE,
  column.format = list(
    c(column = date.col, format = date.format),
    c(column = time.col, format = time.format)
  )
)

ref.dictionary <- importer(path = dictionary.path,
                                multiple = FALSE,
                                graphics = TRUE)

```

### Filter data

```{r}

#!!# Translate and correct species codes
aou_code <- read.csv("reference/IBP-AOS-LIST23.csv", na = "")

raw.data.list <- raw.data.list %>%
  purrr::map(
    .f = ~ .x %>%
      dplyr::inner_join(
        .,
        aou_code[, c("COMMONNAME", "SPEC")],
        by = c("CommonName" = "COMMONNAME"),
        keep = FALSE,
        relationship = "many-to-many"
      ) %>%
      dplyr::mutate("SpeciesCode" = `SPEC`) %>%
      dplyr::select(-"SPEC")
  )

if(any(c(species.col, date.col) %in% unlist(purrr::map(raw.data.list, colnames)))) {
  # Make filter selections
  if (species.col %in% unlist(purrr::map(raw.data.list, colnames))) {
    species <- unique(unlist(sapply(raw.data.list, "[", species.col)))
  } else {
    species<- character(0)
  }
  
  if (date.col %in% unlist(purrr::map(raw.data.list, colnames))) {
    date.ranges <- raw.data.list %>% purrr::map(
      .f = ~ .x %>%
        dplyr::select(all_of(date.col)) %>%
        dplyr::filter(.data[[date.col]] != "") %>%
        dplyr::mutate({{date.col}} := as.Date(.data[[date.col]], format = date.format))
      
      %>%
        append(list(
          "Min" = min(.[[date.col]]), "Max" = max(.[[date.col]])
        ))
    )
    
    dates.df <- raw.data.list %>% purrr::map(
      .f = ~ .x %>%
        dplyr::select(all_of(date.col)) %>%
        dplyr::filter(.data[[date.col]] != "") %>%
        dplyr::mutate({{date.col}} := as.Date(.data[[date.col]], format = date.format))
    ) %>%
      dplyr::bind_rows(., .id = "id")
    
    min.date <- min(do.call("c", lapply(date.ranges, "[[", "Min")))
    max.date <- max(do.call("c", lapply(date.ranges, "[[", "Max")))
    
  } else {
    min.date <- as.Date("0001-01-01")
    max.date <- as.Date("0001-01-01")
  }


  filter.df.app(species.col, date.col, species, min.date, max.date)


# Apply filters
  id.var <- dplyr::last(make.names(c(unique(
    unlist(purrr::map(raw.data.list, colnames))
  ), ".id"), unique = TRUE))
  
  raw.data.list <- raw.data.list %>%
    purrr::map(.f = ~ .x %>%
                 tibble::rowid_to_column(var = id.var))
  
  other.data.list <- raw.data.list
  
  if (length(species) != 0) {
    raw.data.list <- raw.data.list %>%
      purrr::map(.f = ~ .x %>%
                   dplyr::filter(if_any(
                     matches(species.col), ~ .data[[species.col]] %in% selected.species
                   )))
  }
  
  
  if (min.date != "0001-01-01" |
      max.date != "0001-01-01") {
    raw.data.list <- raw.data.list %>%
      purrr::map(
        .f = ~ .x %>%
          dplyr::mutate_at(dplyr::vars(tidyselect::any_of(date.col)), as.Date, format = date.format) %>%
          dplyr::filter(if_any(
            matches(date.col),
            ~ data.table::between(.data[[date.col]], min(selected.dates), max(selected.dates), NAbounds =
                                    FALSE)
          ))
      )
  }
  
  rm.row <-  purrr::map(raw.data.list, id.var)
  
  other.data.list <- lapply(seq_along(rm.row), function(i) {
    if (rlang::is_empty(rm.row[[i]]) == FALSE) {
      other.data.list[[i]] %>%
        dplyr::slice(-rm.row[[i]])
    } else {
      other.data.list[[i]]
    }
  }) %>%
    purrr::set_names(names(raw.data.list))
  
    raw.data.list <- purrr::map(raw.data.list, ~ (.x %>% dplyr::select(-tidyselect::any_of(id.var))))
    
      other.data.list <- purrr::map(other.data.list, ~ (.x %>% dplyr::select(-tidyselect::any_of(id.var))))
   

# Remove empty dataframes
  raw.data.list <- raw.data.list %>%
    purrr::discard(., ~ nrow(.) == 0) %>%
    purrr::map(.f = ~ .x %>%
                 dplyr::mutate({{date.col}} := format(.data[[date.col]], format = date.format)))
  
  
  other.data.list <- other.data.list %>%
    purrr::discard(., ~ nrow(.) == 0) %>%
    purrr::map(.f = ~ .x %>%
                 dplyr::mutate({{date.col}} := format(.data[[date.col]], format = date.format)))
  
}
```

### Merge data

```{r}

if(length(raw.data.list)>1) {
  # Make data frame selections
  merge.df.app(raw.data.list)
  
  # Execute forced rbind
  if (is.null(selected.df) == FALSE) {
    raw.data.list <-  c(list(merged_data = data.table::rbindlist(raw.data.list[names(raw.data.list) %in% selected.df], fill = TRUE)),
                        raw.data.list[!names(raw.data.list) %in% selected.df])
  }
  
  raw.data.list <- raw.data.list %>%
    purrr::map(
      .f = ~ .x %>%
        dplyr::mutate_if(is.character, ~ dplyr::na_if(., "")) %>%
        dplyr::mutate_if(is.character, ~ dplyr::na_if(., "NULL")) %>%
        dplyr::mutate_if(is.character, ~ dplyr::na_if(., missing_values))
    )
}

```

### Validate data

```{r}

# Validate data file(s) against the dictionary
warning.list <- list()
assign("input.dxnry", ref.dictionary)
if (class(ref.dictionary) == "list") {
  input.dxnry <- mdJSONdictio::extract.mdJSON(x = input.dxnry,
                                              record.type = "dictionaries",
                                              multiple = FALSE)
  
  record.num <-  which(sapply(ref.dictionary[["data"]], "[", "id") == input.dxnry[["data"]][[1]][["id"]])
  
  
  dxnry.colnames <- unlist(sapply(rjson::fromJSON(input.dxnry[["data"]][[1]][["attributes"]][["json"]])[["dataDictionary"]][["entity"]][[1]][["attribute"]], "[", "codeName"))
} else {
  dxnry.colnames <- unique(input.dxnry$codeName)
}
   
for (a in 1:length(raw.data.list)) {
  assign("input.data", raw.data.list[[a]])
  
  # Check colnames
  invalid.cols <- colnames(input.data)[!colnames(input.data) %in% dxnry.colnames]
  
  
  if (length(invalid.cols) != 0) {
    for (invalid.col in invalid.cols) {
      unused.cols <- as.vector(dxnry.colnames[!dxnry.colnames %in% colnames(input.data)])
      
      correct.colname.app(input.data, invalid.col, unused.cols)
      
      
      if (decision == "remove") {
        input.data <- input.data %>% dplyr::select(., -invalid.col)
        
      } else if (decision == "replace") {
        input.data <- input.data %>% dplyr::rename(., !!col.choice := !!invalid.col)
        unused.cols <- unused.cols[!unused.cols %in% col.choice]
        
      } else if (decision == "rename") {
        input.data <- input.data %>% dplyr::rename(., !!col.choice := !!invalid.col)
      }
    }
  }
  
  # Create warning table
  if (class(ref.dictionary) == "list") {
    warning.list[[a]] <-
      mdJSONdictio::validate.mdJSON(x = input.dxnry, y = input.data)
  } else {
    warning.list[[a]] <-
      mdJSONdictio::validate.table(x = input.dxnry, y = input.data)
  }
  
  raw.data.list[[a]] <- input.data
  
}

names(warning.list) <- names(raw.data.list)


```

### Correct data

```{r}

for (a in 1:length(raw.data.list)) {
  assign("input.data", raw.data.list[[a]])
  
  if (class(ref.dictionary) == "list") {
    ref.dxnry <- mdJSONdictio::build.table(input.dxnry)
  } else {
    assign("ref.dxnry", input.dxnry)
  }
  
  warn.table <- warning.list[[a]]
  
  # Correct entry errors
  domainItem_value.table <-
    warn.table %>% dplyr::filter(Category == "domainItem_value")
  
  if (nrow(domainItem_value.table) != 0) {
    for (b in 1:nrow(domainItem_value.table)) {
      invalid.values <-
        scan(
          text = noquote(gsub(
            ".*: ", "", domainItem_value.table$Message[b]
          )),
          what = "",
          quiet = TRUE,
          sep = ",",
          strip.white = TRUE
        )
      
      
      variable <- domainItem_value.table$Variable[b]
      
      select.value.app(variable, invalid.values, ref.dxnry)
      
      for (invalid.value in selected.values) {
        correct.value.app(variable, invalid.value, input.data, ref.dxnry)
        
        if (decision %in% c("correct", "replace")) {
          class <- class(input.data[[variable]])
          
          input.data <- input.data %>%
            dplyr::mutate_at(
              variable,
              ~ as.character(.) %>% dplyr::recode(., !!invalid.value := value.choice)
            )
          
          # if (numeric == TRUE &
          #     is.numeric(input.data[[variable]]) == TRUE) {
          #   input.data <- input.data %>%
          #     dplyr::mutate_at(variable, ~ as.numeric(.))
          #   
          # }
          # 
          # if (is.na(match.fun(paste0("as.", class))(value.choice))
          #   == TRUE) {
          #     
          #   }
          
        }
        
        
      }
    }
  }
  
      
  #     test<- summary(arsenal::comparedf(
  #   dplyr::mutate(df, across(
  #     where(is.character), ~ dplyr::na_if(.x, "")
  #   )),
  #   dplyr::mutate(input.data, across(
  #     where(is.character), ~ dplyr::na_if(.x, "")
  #   ))
  # ))
      
  
  raw.data.list[[a]] <- input.data
}

raw.data.list <- purrr::map(raw.data.list, ~ (.x %>% dplyr::select(-tidyselect::any_of(id.var))))


```

### Correct dictionary

```{r}

req.attribute.fields <-
      c("codeName", "allowNull", "dataType", "definition", "domainId")

for (a in 1:length(warning.list)) {
  warn.table <- warning.list[[a]]
  
  
  # codeName
  codeName.warn <-
    warn.table %>% dplyr::filter(Category == "codeName")
  
  if (nrow(codeName.warn) != 0) {
    codeName.choice <- utils::select.list(
      c(codeName.warn$Variable),
      multiple = TRUE,
      graphics = TRUE,
      title = cat(
        paste0(
          "\nSome variables in the data file '",
          names(raw.data.list[a]),
          "' were not found in the data dictionary.\nSelect which, if any, to add to the data dictionary.\n"
        )
      )
    )
    
    if (length(codeName.choice)!=0) {
      for (b in 1:length(codeName.choice)) {
        if (class(ref.dictionary) == "list") {
          input.dxnry <-
            mdJSONdictio::modify.mdJSON(
              x = input.dxnry,
              how = "add_attribute",
              attribute_codeName = codeName.choice[b],
              attribute_units = NULL,
              attribute_unitsResolution = NULL,
              attribute_isCaseSensitive = NULL,
              attribute_missingValue = NULL,
              attribute_minValue = NULL,
              attribute_maxValue = NULL,
              attribute_fieldWidth = NULL,
              quiet = FALSE
            )
        } else {
          input.dxnry <-
            mdJSONdictio::modify.table(
              x = input.dxnry,
              how = "add_attribute",
              attribute_codeName = codeName.choice[b],
              attribute_units = NULL,
              attribute_unitsResolution = NULL,
              attribute_isCaseSensitive = NULL,
              attribute_missingValue = NULL,
              attribute_minValue = NULL,
              attribute_maxValue = NULL,
              attribute_fieldWidth = NULL,
              quiet = FALSE
            )
        }
      }
    }
  }
  
  # domainItem_value
  domainItem_value.table <-
    warn.table %>% dplyr::filter(Category == "domainItem_value")
  
  if (nrow(domainItem_value.table) != 0) {
    for (b in 1:nrow(domainItem_value.table)) {
      domainItem.vector <-
        scan(
          text = noquote(gsub(
            ".*: ", "", domainItem_value.table$Message[b]
          )),
          what = "",
          quiet = TRUE,
          sep = ",",
          strip.white = TRUE
        )
      
      domainItem.choice <-
        utils::select.list(
          c(domainItem.vector),
          multiple = TRUE,
          graphics = TRUE,
          title = cat(
            paste0(
              "\nThe attribute '",
              domainItem_value.table$Variable[b] ,
              "' in the data file '",
              names(raw.data.list[a]),
              "' contains entry values not found in the data dictionary.\nSelect which, if any, to add to the domain item list in the data dictionary.\n"
            )
          )
        )
      
      if (length(domainItem.choice)!=0) {
        for (c in 1:length(domainItem.choice)) {
          
          if (class(ref.dictionary) == "list") {
            input.dxnry <-
              mdJSONdictio::modify.mdJSON(
                x = input.dxnry,
                how = "add_domainItem",
                domain_codeName = domainItem_value.table$Variable[b],
                domainItem_name = domainItem.choice[c],
                domainItem_value = domainItem.choice[c],
                quiet = FALSE
              )
          } else {
            input.dxnry <-
              mdJSONdictio::modify.table(
                x = input.dxnry,
                how = "add_domainItem",
                domain_codeName = domainItem_value.table$Variable[b],
                domainItem_name = domainItem.choice[c],
                domainItem_value = domainItem.choice[c],
                quiet = FALSE
              )
          }
        }
        
      }
    }
  }
  
  # dataType
  
  dataType.table <-
    warn.table %>% dplyr::filter(Category == "dataType_RDatatype")
  
  if (nrow(dataType.table) != 0) {
    for (b in 1:nrow(dataType.table)) {
      detected.dataType <-
        gsub("[\\(\\)]",
             "",
             regmatches(
               dataType.table$Message[b],
               gregexpr("\\(.*?\\)", dataType.table$Message[b])
             )[[1]])[1]
      dxnry.dataType <-
        gsub("[\\(\\)]",
             "",
             regmatches(
               dataType.table$Message[b],
               gregexpr("\\(.*?\\)", dataType.table$Message[b])
             )[[1]])[2]
      
      dataType.choice <-
        utils::select.list(
          c("yes", "no"),
          multiple = FALSE,
          graphics = TRUE,
          title = cat(
            paste0(
              "\nThe attribute '",
              dataType.table$Variable[b] ,
              "' in the data file '",
              names(raw.data.list[a]),
              "' has entry values with a different dataType (",
              detected.dataType,
              ") than indicated in the dictionary (",
              dxnry.dataType,
              ").\nWould you like to update the dataType in the dictionary?\n"
            )
          )
        )
      
      if ("yes" %in% dataType.choice) {
        input.dxnry <-
          mdJSONdictio::modify.mdJSON(
            x = input.dxnry,
            how = "update_attribute",
            attribute_codeName = dataType.table$Variable[b],
            attribute_allowNull = NULL,
            attribute_dataType = detected.dataType,
            attribute_definition = NULL,
            attribute_units = NULL,
            attribute_unitsResolution = NULL,
            attribute_isCaseSensitive = NULL,
            attribute_missingValue = NULL,
            attribute_minValue = NULL,
            attribute_maxValue = NULL,
            attribute_fieldWidth = NULL,
            domain_codeName = NULL,
            new.attribute_codeName = NULL,
            quiet = FALSE
          )
        
      }
    }
  }

}

ref.dictionary[["data"]][[record.num]] <- input.dxnry[["data"]][[1]]

  
```

### Quality checks

```{r}
# continuous: box plots
# discrete: range of values (interactive table)
# plot lat/long

# All data
for (x in 1:length(raw.data.list)) {
  assign("input.data", raw.data.list[[x]])
  
  ## Duplicate rows

  
  # All capture data
  
  ## Unique values
  ### BBLDisposition = 1 & BandNumberOut
  ### Flag code & BandNumberOut
  ### Colorband combo & BandNumberOut

  ## Species v AOU codes
  
  ## Status codes
 
  ## "Out" leg_cols contains an X
  leg_cols <-   c("ULIn",
                  "LLIn",
                  "URIn",
                  "LRIn",
                  "ULOut",
                  "LLOut",
                  "UROut",
                  "LROut")
  
  leg_cols <- leg_cols[leg_cols %in% colnames(input.data)]
  
  which(rowSums(is.na(input.data[, leg_cols[grepl("Out", leg_cols)]])) == 4)
  
  # Recapture data
  ## "In" leg_cols contains an X
  which(rowSums(is.na(input.data[, leg_cols[grepl("In", leg_cols)]])) == 4 &
          input.data$BBLDisposition == "R")
  
  ## BandNumberIn == BandNumberOut
  which(is.na(input.data$BandNumberIn) &
          input.data$BBLDisposition == "R")
  
  raw.data.list[[x]] <- input.data
  
}

```

# Data postprocessing

### Add uuids

```{r}
raw.data.list <- raw.data.list %>%   purrr::map(
  .f = ~ .x %>%       dplyr::mutate(!!!setNames(
    rep(NA, length(setdiff(
      "OccurenceID", colnames(.x)
    ))), setdiff("OccurenceID", colnames(.x))
  )) %>% dplyr::relocate("OccurenceID") %>%       dplyr::rowwise() %>%       dplyr::mutate_at(
    "OccurenceID",
    ~ ifelse(is.na(.) == TRUE ||
               . == "", uuid::UUIDgenerate(use.time = FALSE), .)
  )
)    
```

### Join ancillary data

#### Location

```{r}
# Provide directory paths for the location data file(s)
ref.data.path <- "reference/ALMS/"

# Provide the relevent information
ref.point.prefix <- "EIEL_"
data.point.col <- "UnitID"
data.lat.col <- "Latitude_dec"
data.long.col <- "Longitude_dec"

```

```{r}

# Import reference data file(s)
ref.data.files <- importer(
  path = ref.data.path,
  multiple = TRUE,
  graphics = TRUE,
  pattern = c("*.gpx$", "*.kml$")
)

locations<- data.table::rbindlist(ref.data.files, fill = TRUE)

# Join location(s)
raw.data.list <- raw.data.list %>%
  purrr::map(
    .f = ~ .x %>%
      dplyr::left_join(
        .,
        locations,
        by = c(setNames("temp:Name", data.point.col)), 
               # "Year" = "temp:year"),
        keep = FALSE
      ) %>%
      dplyr::mutate_at(
        data.lat.col,
        ~ dplyr::case_when(is.na(.) ~ `temp:Latitude`, TRUE ~ !!as.name(data.lat.col))
      ) %>%
      dplyr::mutate_at(
        data.long.col,
        ~ dplyr::case_when(is.na(.) ~ `temp:Longitude`, TRUE ~ !!as.name(data.long.col))
      ) %>%
      dplyr::select(-tidyselect::starts_with("temp:"))
  )
    
```

#### Sex results

```{r}
file.path = "C:/Users/hvincelette/OneDrive - DOI/Documents/Fieldwork_wrap/2023/Sexing_Analysis/"
```

```{r}

# Import sex results
sex_results <- importer(path = file.path,
                        multiple = FALSE,
                        graphics = TRUE)



# Correct results table
sex_results<- sex_results %>%
  dplyr::mutate_at("BandNumber", as.character) %>%
  tidyr::drop_na("OccurenceID") %>%
  dplyr::filter(Sex %in% c("F","M"))


# Check for duplicate results (creates duplicate rows in left_join)
sex_results[c(
  anyDuplicated(sex_results$BandNumber, fromLast = TRUE),
  anyDuplicated(sex_results$BandNumber, fromLast = FALSE)
), ]


# Join sex results
raw.data.list <- raw.data.list %>%
  purrr::map(
    .f = ~ .x %>%
      dplyr::mutate(!!!setNames(
        rep(NA, length(setdiff(
          "CHDSex", colnames(.x)
        ))), setdiff("CHDSex", colnames(.x))
      )) %>%
      dplyr::relocate("CHDSex", .before = "FieldSex") %>%
      dplyr::left_join(.,dplyr::distinct(sex_results[,c("BandNumber","Sex")]), by = c("BandNumberOut" = "BandNumber"), keep = FALSE, relationship = "many-to-many") %>%
    dplyr::mutate_at("CHDSex", ~dplyr::case_when(is.na(.) ~ Sex,TRUE ~ CHDSex)) %>%
      dplyr::select(-"Sex")
    )


```

### Fill missing info

#### Bird Status

```{r}

ref_data <-
  lapply(list.files(path = "reference/bander_portal/", full.names = TRUE),
         readxl::read_excel)

names(ref_data) <-
  gsub(
    list.files(path = "reference/bander_portal/"),
    pattern = ".xlsx$",
    replacement = ""
  )

list2env(ref_data, globalenv())

for (x in 1:length(raw.data.list)) {
  assign("input.data", raw.data.list[[x]])
  
  # Add missing columns
  miss_cols <- setdiff(
    c(
      na.omit(capture_cols$my_col[!grepl("color", capture_cols$my_col)]),
      "TagType1",
      "TagType2",
      "TagStatus1",
      "TagStatus2"
    ),
    colnames(input.data)
  )
  
  input.data[miss_cols] <-
    as.character(NA)
  
  
  my_colors <- c(
    "Bk",
    "Br",
    "Db",
    "Lb",
    "Mb",
    "Dg",
    "Lg",
    "Gy",
    "Hp",
    "Mg",
    "Mv",
    "Or",
    "Pp",
    "Rd",
    "Wh",
    "Yl"
  )
  
  leg_cols <-   c("ULIn",
                  "LLIn",
                  "URIn",
                  "LRIn",
                  "ULOut",
                  "LLOut",
                  "UROut",
                  "LROut")
  
  leg_cols <- leg_cols[leg_cols %in% colnames(input.data)]
  
  status_info <- input.data %>%
    dplyr::rowwise() %>%
    dplyr::mutate_at("TagType1", ~ ifelse(TagStatus1  %in% c("Removed", "Lost"), TagType1 <-
                                            NA
                                          , .)) %>%
    dplyr::mutate_at("TagType2", ~ ifelse(TagStatus2  %in% c("Removed", "Lost"), TagType2 <-
                                            NA
                                          , .)) %>%
    dplyr::select(tidyselect::any_of(
      c(
        "OccurenceID",
        "BandNumberOut",
        "SpeciesCode",
        "BBLStatus",
        leg_cols[grepl("Out", leg_cols)],
        "TagType1",
        "TagType2",
        "Blood",
        "BBLStatus"
      )
    )) %>%
    dplyr::mutate_at(dplyr::vars(c(
      leg_cols[grepl("Out", leg_cols)], "TagType1", "TagType2", "BBLStatus"
    )), ~ as.character(.) %>% dplyr::na_if(., "")) %>%
    dplyr::mutate_at(
      leg_cols[grepl("Out", leg_cols)],
      ~ stringi::stri_replace_all_fixed(., my_colors, "01A", vectorize_all = FALSE)
    ) %>%
    dplyr::mutate(NumMarkers = sum(!unique(dplyr::c_across(
      c(leg_cols[grepl("Out", leg_cols)], "TagType1", "TagType2")
    ))  %in% c("X", NA))) %>%
    dplyr::mutate_at(
      "BBLStatus",
      ~ dplyr::case_when(
        NumMarkers >= 2 &
          Blood != "Y" ~ "325",
        NumMarkers >= 1 &
          Blood == "Y" ~ "319",
        NumMarkers == 0 &
          Blood == "Y" ~ "318",
        NumMarkers == 0 &
          Blood != "Y" ~ "300",
        NumMarkers == 1 &
          dplyr::if_any(c(leg_cols[grepl("Out", leg_cols)], "TagType1", "TagType2"), ~
                          . == "69") &
          Blood != "Y" ~ "369",
        NumMarkers == 1 &
          dplyr::if_any(c(leg_cols[grepl("Out", leg_cols)], "TagType1", "TagType2"), ~
                          grepl("^81", .)) &
          Blood != "Y" ~ "381",
        NumMarkers == 1 &
          dplyr::if_any(c(leg_cols[grepl("Out", leg_cols)], "TagType1", "TagType2"), ~
                          grepl("^80", .)) &
          Blood != "Y" ~ "380",
        NumMarkers == 1 &
          dplyr::if_any(c(leg_cols[grepl("Out", leg_cols)], "TagType1", "TagType2"), ~
                          grepl("^90", .)) &
          Blood != "Y" ~ "390",
        NumMarkers == 1 &
          dplyr::if_any(c(leg_cols[grepl("Out", leg_cols)], "TagType1", "TagType2"), ~
                          grepl("^01", .)) &
          Blood != "Y" ~ "301",
        TRUE ~ BBLStatus
      )
    ) %>%
    dplyr::rename(`temp:BBLStatus` = BBLStatus)
  
  input.data <- input.data %>%
    dplyr::mutate(across(where(is.character), ~ dplyr::na_if(.x, ""))) %>%
    dplyr::left_join(
      .,
      status_info[, c("OccurenceID", "temp:BBLStatus")],
      by = c("OccurenceID" = "OccurenceID"),
      keep = FALSE,
      relationship = "one-to-one"
    ) %>%
    dplyr::mutate_at("BBLStatus", ~ as.character(.)) %>%
    dplyr::mutate_at("BBLStatus",
                     ~ dplyr::case_when(is.na(.) ~ `temp:BBLStatus`, TRUE ~ BBLStatus))
  
  invalid.index <- which(input.data$BBLStatus != input.data$`temp:BBLStatus`)
  
  
  if (length(invalid.index) != 0) {
    for (a in invalid.index) {
      message(cat(
        paste0(
          "The status for the following individual appears invalid.\nSpecies: ",
          input.data$SpeciesCode[a],
          "\nBand number: ",
          input.data$BandNumberOut[a],
          "\nStatus (reported): ",
          input.data$BBLStatus[a],
          " [",
          bander_portal_lookups$description[grepl(stringr::str_sub(input.data$BBLStatus[a], 1, 1),
                                                  bander_portal_lookups$code) == TRUE &
                                              bander_portal_lookups$field == "Bird Status"],
          " ",
          bander_portal_lookups$description[grepl(stringr::str_sub(input.data$BBLStatus[a], 2, 3),
                                                  bander_portal_lookups$code) == TRUE &
                                              bander_portal_lookups$field == "Bird Status Extra Info"],
          "]",
          "\nStatus (detected): ",
          input.data$`temp:BBLStatus`[a],
          " [",
          bander_portal_lookups$description[grepl(
            stringr::str_sub(input.data$`temp:BBLStatus`[a], 1, 1),
            bander_portal_lookups$code
          ) == TRUE &
            bander_portal_lookups$field == "Bird Status"],
          " ",
          bander_portal_lookups$description[grepl(
            stringr::str_sub(input.data$`temp:BBLStatus`[a], 2, 3),
            bander_portal_lookups$code
          ) == TRUE &
            bander_portal_lookups$field == "Bird Status Extra Info"],
          "]",
          "\nBlood: ",
          input.data$Blood[a],
          "\nULOut: ",
          input.data$ULOut[a],
          "\nLLOut: ",
          input.data$LLOut[a],
          "\nUROut: ",
          input.data$UROut[a],
          "\nLROut: ",
          input.data$LROut[a],
          "\nTagStatus1: ",
          input.data$TagStatus1[a],
          "\nTagStatus2: ",
          input.data$TagStatus2[a]
        )
      ))
      
      status.choice <- utils::menu(
        choice = c("update as detected", "manually edit", "leave as is"),
        graphics = TRUE,
        title = cat("\nHow would you like to proceed?")
      )
      
      if (status.choice == 1) {
        input.data$BBLStatus[a] <- input.data$`temp:BBLStatus`[a]
        
        
      }
      if (status.choice == 2) {
        message(
          "\nIMPROTANT: you must click 'syncronise' (arrow loop icon) and then 'Done' to enact changes."
        )
        
        invalid.entries <- input.data %>%
          dplyr::ungroup() %>%
          dplyr::filter(dplyr::row_number() == a)
        
        invalid.entries.edit <-
          DataEditR::data_edit(
            invalid.entries,
            col_readonly = colnames(invalid.entries)[!colnames(invalid.entries) %in% "BBLStatus"],
            viewer = "pane"
          ) %>%
          dplyr::mutate_at("BBLStatus", as.character)
        
        input.data[match(invalid.entries.edit$OccurenceID,
                         input.data$OccurenceID), "BBLStatus"] <-
          invalid.entries.edit$BBLStatus
      }
      
    }
  }
  
  
  input.data <-
    input.data %>%
    dplyr::select(-tidyselect::starts_with("temp:"))
  
  raw.data.list[[x]] <- input.data
  
}
  
```

### Export

```{r}
# Provide directory path for the authoritative data file
download.path <- "downloads/data"

# Provide a directory path to write processed files
processed.data.path <- "data-processed/to_archive/"

# Do you want to merge the data with an authoritative file?
auth.merge <- TRUE

# Provide versioning format (when merging with an authoritative file)
vers.suffix <- "_v"

```

#### Data

```{r}

if (auth.merge == TRUE) {
  # Import authoritative data file
  auth.data <- importer(
    path = download.path,
    multiple = FALSE,
    graphics = TRUE,
    column.format = list(
      c(column = date.col, format = date.format),
      c(column = time.col, format = time.format)
    )
  )
  
  # Rejoin filtered data
  raw.data.list <-
    split(c(raw.data.list, other.data.list), c(names(raw.data.list), names(other.data.list))) %>%
    purrr::map(dplyr::bind_rows)
  
  # Select files to merge
  if(length(raw.data.list)>1) {
    raw.data.choice <- utils::select.list(
      c(names(raw.data.list)),
      multiple = TRUE,
      graphics = TRUE,
      title = cat(paste0(
        "\nSelect one or more data files to merge.\n"
      ))
    )
    
    raw.data.list <- raw.data.list %>% purrr::keep(names(raw.data.list) %in% raw.data.choice)
  }
  
  add.data <- dplyr::bind_rows(raw.data.list)
  
  
  # Create new columns
  new.cols <- add.data %>%
    dplyr::mutate_if(is.character, ~ dplyr::na_if(., c(""))) %>%
    purrr::discard(., ~ all(is.na(.))) %>%
    colnames(.)[!colnames(.) %in% colnames(auth.data)] %>%
    colnames()
  
  if (length(new.cols) != 0) {
    col.choice <- utils::select.list(
      c(new.cols),
      multiple = TRUE,
      graphics = TRUE,
      title = cat(
        paste0(
          "\nOne or more columns do not exist in the authoritative data file.\nSelect which if any to add\nNote, unselected columns will not be archived.\n"
        )
      )
    )
    if (length(col.choice) != 0) {
      for (a in col.choice) {
        loc.choice <- utils::select.list(
          c(colnames(auth.data)), 
          multiple = FALSE, 
          graphics = TRUE,
          title = cat(
          paste0(
            "\nInsert '",
            a,
            "' before which existing column?\nEnter 0 to add after the last column.\n"
          )
        ))
        if (loc.choice != 0) {
          before <- loc.choice
        } else {
          before <- NULL
        }
        auth.data <- auth.data %>% tibble::add_column(!!a := NA, .before = before)
      }
    }
  }
  
  # Merge data
  new.data <- add.data %>%
    merge(auth.data, all = TRUE) %>%
    dplyr::select(colnames(auth.data))
  
} else if (auth.merge == FALSE) {
  # Rejoin filtered data
  raw.data.list <-
    split(c(raw.data.list, other.data.list), c(names(raw.data.list), names(other.data.list))) %>%
    purrr::map(dplyr::bind_rows)
  
  #!!!# Evaluate changes
  test<- summary(arsenal::comparedf(
    dplyr::mutate(new.data, across(
      where(is.character), ~ dplyr::na_if(.x, "")
    )),
    dplyr::mutate(auth.data, across(
      where(is.character), ~ dplyr::na_if(.x, "")
    ))
  ))
  
  # Select files to merge
  raw.data.choice <- utils::select.list(
    c(names(raw.data.list)),
    multiple = TRUE,
    graphics = TRUE,
    title = cat(paste0(
      "\nSelect one or more data files to combine.\n"
    ))
  )
  
  
  raw.data.list <- raw.data.list %>% purrr::keep(names(raw.data.list) %in% raw.data.choice)
  
  new.data <- dplyr::bind_rows(raw.data.list)
  
}

```

```{r}

  # Versioning
  # detect.vers <- suppressWarnings(as.numeric(ifelse(
  #   grepl(paste0("\\", vers.suffix), auth.data.name, ignore.case = TRUE),
  #   sub(
  #     paste0(".*\\", vers.suffix),
  #     "",
  #     auth.data.name,
  #     ignore.case = TRUE
  #   ),
  #   ""
  # )))
  
  # if (is.na(detect.vers) == FALSE) {
  #   message(cat(
  #     paste0(
  #       "Provide a version number for the new authoritative data file.\nThe detected current version is '",
  #       detect.vers,
  #       "'."
  #     )
  #   ))
  # } else {
    message(cat(
      paste0(
        "Provide a version number for the new authoritative data file.\nA current version was not detected."
      )
    ))
  # }
  
  new.vers <-
    noquote(as.character(readline(prompt = )))
  
  if (new.vers != "") {
    new.vers <- paste0(vers.suffix, new.vers)
  }
  
  message(cat(paste0("\nProvide a name for the data file.\n")))
  
  new.data.name <- noquote(as.character(readline(prompt =)))
  
  new.data.name <-  sub(" ", "_", new.data.name)
  
  new.data.name<- gsub('^[[:punct:]]|[[:punct:]]$', '', new.data.name)

  new.data.name <- paste0(new.data.name, new.vers)
  


write.csv(
  new.data,
  paste0(processed.data.path, new.data.name, ".csv"),
  na = "",
  row.names = FALSE
)

```

#### Dictionary

```{r}

if(class(ref.dictionary) == "list") {
  export.choice <- utils::select.list(
    c("tabular dictionary", "mdjson dictionary", "mdEditor file"),
    multiple = TRUE,
    graphics = TRUE,
    title = cat(paste0(
      "\nWhich format(s) would you like to export?\n"
    ))
  )
  
  if ("tabular dictionary" %in% export.choice) {
    message(cat(
      paste0(
        "\nProvide a name for the mdjson dictionary.\nThe mdjson record name is '",
        input.dxnry[["data"]][[1]][["meta"]][["title"]],
        "'.\n"
      )
    ))
    
    dxnry.name <- noquote(as.character(readline(prompt =)))
    
    if (dxnry.name == "") {
      dxnry.name <- input.dxnry[["data"]][[1]][["meta"]][["title"]]
    }
    
    dxnry.name <-  sub(" ", "_", dxnry.name)
    
    dxnry.name <- strsplit(gsub("[[:punct:]]$", "", dxnry.name), " +")[[1]]
    
    table.dxnry <- mdJSONdictio::build.table(input.dxnry)
    
    
    write.csv(
      table.dxnry,
      paste0(
        processed.data.path,
        dxnry.name,
        "_",
        format(
          as.POSIXct(input.dxnry[["data"]][[1]][["attributes"]][["date-updated"]], tz = "UTC", "%Y-%m-%dT%H:%M:%OS"),
          '%Y%m%d-%H%M%S'
        ),
        ".csv"
      ),
      na = "",
      row.names = FALSE
    )
    
  }
  
  if ("mdjson dictionary" %in% export.choice) {
    if (!"tabular dictionary" %in% export.choice) {
      message(cat(
        paste0(
          "\nProvide a name for the mdjson dictionary.\nThe mdjson record name is '",
          input.dxnry[["data"]][[1]][["meta"]][["title"]],
          "'.\n"
        )
      ))
      
      dxnry.name <- noquote(as.character(readline(prompt = )))
      
      if (dxnry.name == "") {
        dxnry.name <- input.dxnry[["data"]][[1]][["meta"]][["title"]]
      }
      
      dxnry.name <-  sub(" ", "_", dxnry.name)
      
      dxnry.name <- strsplit(gsub("[[:punct:]]$", "", dxnry.name), " +")[[1]]
      
    }
    
    new.dxnry <- rjson::toJSON(x = input.dxnry)
    
    write(
      x = new.dxnry,
      file = paste0(
        processed.data.path,
        dxnry.name,
        "_mdeditor-",
        format(
          as.POSIXct(input.dxnry[["data"]][[1]][["attributes"]][["date-updated"]], tz = "UTC", "%Y-%m-%dT%H:%M:%OS"),
          '%Y%m%d-%H%M%S'
        ),
        ".json"
      )
    )
    
  }
  
  if ("mdEditor file" %in% export.choice) {
    message(cat(
      paste0(
        "\nProvide a name for the mdjson file."
        # \nThe current file name is '",
        # gsub("\\_mdeditor.*", "", dictionary.choice, ignore.case = TRUE) ,
        # "'.\n"
      )
    ))
    
    file.name <- noquote(as.character(readline(prompt = )))
    
    # if (file.name == "") {
    #   file.name <-  gsub("\\_mdeditor.*", "", dictionary.choice, ignore.case = TRUE)
    # }
    
    file.name <-  sub(" ", "_", file.name)
    file.name <- gsub('^[[:punct:]]|[[:punct:]]$', '', file.name)
    
    new.dxnry <- rjson::toJSON(x = ref.dictionary)
    
    write(
      x = new.dxnry,
      file = paste0(
        processed.data.path,
        file.name,
        "_mdeditor-",
        format(
          as.POSIXct(ref.dictionary[["data"]][[record.num]][["attributes"]][["date-updated"]], tz = "UTC", "%Y-%m-%dT%H:%M:%OS"),
          '%Y%m%d-%H%M%S'
        ),
        ".json"
      )
    )
    
  }
  
}

```

# Archive in the Regional Data Repository

```{r}
project = "mbmlb_004_Eielson"
path = "C:/Users/hvincelette/OneDrive - DOI/Documents/Data_management/landbirds_project_maintenance/Archived_Projects/Eielson"

```

```{r}
#
commit <-
  FWSAkRDRtools::commit.files(project = project,
               local.folder = path,
               recursive = TRUE)


```

# Publish to Science Base

# Capture data transformations

#### Provide inputs

```{r}
# Provide a directory path for data file(s) to import
data.path <- "data-processed/to_archive/"

# Provide the relevant information (if applicable)
missing_values<- c(NA)
species.col <- "SpeciesCode"
date.col <- "Date"
date.format <-  "%m/%d/%Y"
time.col <- "CaptureTime_2400"
time.format <- "%H:%M"

```

#### Import data

```{r}

data.list <- importer(
  path = data.path,
  multiple = TRUE,
  graphics = TRUE,
  column.format = list(
    c(column = date.col, format = date.format),
    c(column = time.col, format = time.format)
  )
)

```

#### Filter data

```{r}

# Select data to transform
filter.option <- utils::select.list(
  c("by species", "by date range"),
  multiple = FALSE,
  graphics = TRUE,
  title = cat(
    paste0("\nHow do you want to filter the imported files?\n"),
    paste0(names(data.list), sep = "\n"),
    paste0("\nEnter 0 to review all data.\n")
  )
)

if (filter.option == "by species") {
  sp.vectors <- sapply(sapply(data.list, "[", species.col), unique)
  
  sp.overlap <- Reduce(intersect, sp.vectors)
  
  if (length(sp.overlap) == 0) {
    message(cat("A common species was not detected across imported files."))
  } else if (length(sp.overlap) > 0) {
    message(cat(
      paste0("The following species were detected in all imported files.\n"),
      paste0(sp.overlap, sp = "\n")
    ))
  }
  
  filter.choice <- utils::select.list(
    c(unique(unlist(sp.vectors))),
    multiple = TRUE,
    graphics = TRUE,
    title = cat(
      paste0(
        "\nSelect one or more species.\nUnselected species will be excluded from subsequent data processing.\n"
      )
    )
  )
  
  if (sum(sapply(temp, nrow)) == 0) {
    warning("Filter did not return results and will not be applied.")
    
  } else {
    data.list <- data.list %>%
      purrr::map(.f = ~ .x %>%
                   dplyr::filter(.data[[species.col]] %in% filter.choice))
  }
  
  
}


if (filter.option == "by date range") {
  date.ranges <- data.list %>% purrr::map(
    .f = ~ .x %>%
      dplyr::select(all_of(date.col)) %>%
      dplyr::filter(.data[[date.col]] != "") %>%
      dplyr::mutate({{date.col}} := as.Date(.data[[date.col]], format = date.format))
    
    %>%
      append(list(
        "Min" = min(.[[date.col]]), "Max" = max(.[[date.col]])
      ))
  )
  
  date.message <- c()
  
  for (a in 1:length(date.ranges)) {
    date.message[a] <-
      paste0(names(date.ranges[a]),
             ": ",
             date.ranges[[a]][["Min"]],
             " - ",
             date.ranges[[a]][["Max"]])
  }
  
  message(cat(
    paste0("The following date ranges were detected the imported files.\n"),
    paste0(date.message, sp = "\n")
  ))
  
  message(cat(
    paste0(
      "Provide start and end dates formatted as %Y-%m-%d (e.g., 2022-1-1).\nMissing and out of range dates will be excluded from subsequent data processing.\n"
    )
  ))
  
  
  start.date <- NA
  
  while (is.na(start.date)) {
    start.date.entry <- as.character(readline(prompt = "Start: "))
    
    if (is.na(as.Date(start.date.entry, format = "%Y-%m-%d"))) {
      message("\nError: Date incorrectly formatted.\n")
      
    } else {
      start.date <- as.Date(start.date.entry, format = "%Y-%m-%d")
    }
    
  }
  
  end.date <- NA
  
  while (is.na(end.date)) {
    end.date.entry <- as.character(readline(prompt = "End: "))
    
    if (is.na(as.Date(end.date.entry, format = "%Y-%m-%d"))) {
      message("\nError: Date incorrectly formatted.\n")
      
    } else {
      end.date <- as.Date(end.date.entry, format = "%Y-%m-%d")
    }
    
  }
  
  temp <- data.list %>%
    purrr::map(.f = ~ .x %>%
                 dplyr::mutate({{date.col}} := as.Date(.data[[date.col]], format = date.format))
               
               %>%
                 dplyr::filter(
                   data.table::between(.data[[date.col]], start.date, end.date, NAbounds =
                                         FALSE)
                 ))
  
  if (sum(sapply(temp, nrow)) == 0) {
    warning("Filter did not return results and will not be applied.")
    
  } else {
    data.list <- temp
    
    }
}

# Remove empty dataframes
data.list <- data.list %>%
  purrr::discard(., ~ nrow(.) == 0) %>%
  purrr::map(.f = ~ .x %>%
               dplyr::mutate({{date.col}} := format(.data[[date.col]], format = date.format)))

```

### Biological samples inventory

```{r}

# Provide directory path for the authoritative data file
download.path <- "downloads/"

# Provide a directory path to write processed files
processed.data.path <- "data-processed/to_archive/"

# Do you want to merge the data with an authoritative file?
auth.merge <- TRUE

# Provide versioning format (when merging with authoritative file)
vers.suffix <- "_v"

```

```{r}

# Prep data
cap.data <- plyr::ldply(data.list)

cap.data %>%
  tibble::as_tibble() %>%
  dplyr::filter(if_all(c("Date", "Year"), is.na))

cap.data <- cap.data %>%
  dplyr::mutate_at("BandNumberOut", ~ as.character(.)) %>%
  dplyr::mutate_if(is.character, ~ dplyr::na_if(., "")) %>%
  dplyr::mutate_if(is.character, ~ dplyr::na_if(., "NULL")) %>%
  dplyr::mutate_if(is.character, ~ dplyr::na_if(., missing_values)) %>%
  tidyr::drop_na(BandNumberOut, Year)


# Load reference tables
ref_data <-
  lapply(list.files(path = "reference/biosample_inventory/", full.names = TRUE),
         utils::read.csv, na.strings = "")

names(ref_data) <-
  gsub(
    list.files(path = "reference/biosample_inventory/"),
    pattern = ".csv$",
    replacement = ""
  )

list2env(ref_data, globalenv())


# Import authoritative data file
if (auth.merge == TRUE) {
  auth.data <- importer(
    path = download.path,
    multiple = FALSE,
    graphics = TRUE,
    column.format = list(
      c(column = date.col, format = date.format),
      c(column = time.col, format = time.format)
    )
  )
    

  auth.data <- auth.data %>%
    dplyr::mutate_at("BandNumber", ~ as.character(.)) %>%
    dplyr::mutate_if(is.character, ~ dplyr::na_if(., "")) %>%
    dplyr::mutate_if(is.character, ~ dplyr::na_if(., "NULL")) %>%
    dplyr::mutate_if(is.character, ~ dplyr::na_if(., missing_values))
  
  # Remove inventoried data
  cap.data <- cap.data %>%
    dplyr::anti_join(auth.data,
                     by = dplyr::join_by("BandNumberOut" == "BandNumber", "Year" == "YearCollected"))
  
}


```

#### Recode data

```{r}

# Correct colnames
cap.data <- cap.data %>%
  dplyr::rename_at(dplyr::vars(matches("Out$")), ~ stringr::str_remove(., "Out$")) %>%
  plyr::rename(
    replace = c(
      "Year" = "YearCollected",
      "Month" = "MonthCollected",
      "Day" = "DayCollected",
      "BanderID" = "Collector"
    )
  ) %>%
  tidyr::unite("Feather",
               c("FeatherType", "FeatherType1")[c("FeatherType", "FeatherType1") %in% names(cap.data)],
               remove = TRUE,
               na.rm = TRUE) %>%
  dplyr::mutate_at("Feather", ~ dplyr::na_if(., ""))

# Pivot sample columns

# Blood
  # c("BloodGenetics",
  # "BloodIsotope",
  # "BloodHg")
  
sample.cols <- c(
  "Feather",
  "AISwabO",
  "AISwabC",
  "Fecal",
  "Nails",
  "BloodFilter",
  "BloodCapTube",
  "BloodSmear"
)

cap.data <- cap.data %>%
  tidyr::pivot_longer(
    cols = tidyselect::any_of(sample.cols),
    names_to = "SampleType",
    values_to = "SampleTaken"
  ) %>%
  dplyr::mutate("FeatherType1" = ifelse(SampleType == "Feather", SampleTaken, NA)) %>%
  dplyr::mutate("SampleTaken" = ifelse(
    SampleType == "Feather" &
      is.na(FeatherType1) == FALSE,
    "Y",
    SampleTaken
  )) %>%
  dplyr::filter(SampleTaken %in% c("Y"))


```

#### Translate data

```{r}

# Auto fill sample details
cap.data<- cap.data %>%
  dplyr::left_join(.,sample_info,
                   by = "SampleType",
                   keep = FALSE,
                   relationship = "many-to-one")


# Add missing info
## OccurenceID

cap.data %>%
  dplyr::filter_at("SpeciesCode", is.na)

cap.data %>%
  dplyr::filter(if_all(c("StateProvince", "City", "Area"), is.na))


cap.data<- cap.data %>%
  dplyr::mutate(
    "StorageMethod" = "frozen",
    "StorageLocation" = "Alaska MBM",
    "SampleStatus" = NA,
    "Contact" = NA,
    "Notes" = NA
  )


```

#### Export data

```{r}

if(auth.merge==TRUE) {
  # Merge data
  new.data <- auth.data %>%
    merge(cap.data, all = TRUE) %>%
    dplyr::select(colnames(auth.data))

  # Versioning
  detect.vers <- suppressWarnings(as.numeric(ifelse(
    grepl(paste0("\\", vers.suffix), auth.data.name, ignore.case = TRUE),
    sub(
      paste0(".*\\", vers.suffix),
      "",
      auth.data.name,
      ignore.case = TRUE
    ),
    ""
  )))
  
  if (is.na(detect.vers) == FALSE) {
    message(cat(
      paste0(
        "Provide a version number for the new authoritative data file.\nThe detected current version is '",
        detect.vers,
        "'."
      )
    ))
  } else {
    message(cat(
      paste0(
        "Provide a version number for the new authoritative data file.\nA current version was not detected."
      )
    ))
  }
  
  new.vers <-
    noquote(as.character(readline(prompt = )))
  
  if (new.vers != "") {
    new.vers <- paste0(vers.suffix, new.vers)
  }
  
  new.data.name <- paste0(sub(
    paste0("\\", vers.suffix, ".*"),
    "",
    auth.data.name,
    ignore.case = TRUE
  ), new.vers)
  
} else if (auth.merge == FALSE) {
  message(cat(paste0("\nProvide a name for the data file.\n")))
  
  new.data.name <- noquote(as.character(readline(prompt =)))
  
  new.data.name <-  sub(" ", "_", data.name)
  
  new.data.name <- strsplit(gsub("[[:punct:]]$", "", data.name), " +")[[1]]
  
}


write.csv(
  new.data,
  paste0(processed.data.path, new.data.name, ".csv"),
  na = "",
  row.names = FALSE
)


```

#### Update sample status

```{r}
# Provide the directory path to the sex results
file.path = "C:/Users/hvincelette/OneDrive - DOI/Documents/Fieldwork_wrap/2023/Sexing_Analysis/"
```

```{r}
# Import results
sex_results <- importer(path = file.path,
                        multiple = FALSE,
                        graphics = TRUE)


# Correct results
sex_results<- sex_results %>%
  dplyr::mutate_at("BandNumber", as.character) %>%
  tidyr::drop_na("OccurenceID") %>%
  dplyr::filter(Sex %in% c("F","M"))


# Check for duplicate results (creates duplicate rows in left_join)
sex_results[c(
  anyDuplicated(sex_results$BandNumber, fromLast = TRUE),
  anyDuplicated(sex_results$BandNumber, fromLast = FALSE)
), ]


# Join results
new.data <- new.data %>%
  dplyr::mutate(".id" = 1:nrow(.)) %>%
  dplyr::left_join(
    .,
    dplyr::distinct(sex_results[, c("SpeciesCode",
                                    "BandNumber",
                                    "YearCollected",
                                    "SampleTissue",
                                    "SampleForm",
                                    "Sex")]),
    by = c(
      "SpeciesCode",
      "BandNumber",
      "YearCollected",
      "SampleTissue",
      "SampleForm"
    ),
    keep = FALSE,
    relationship = "many-to-many"
  ) %>%
  dplyr::mutate_at(
    "SampleStatus",
    ~ dplyr::case_when(Sex %in% c("F", "M") ~ "analysis complete", TRUE ~ SampleStatus)
  ) %>%
  dplyr::select(-"Sex")


# # If status is out of sync..
# unanalyzed.index <- which(
#   new.data$Contact == "Sarah Sonsthagen" &
#     is.na(new.data$Sex) &
#     new.data$SpeciesCode != "REKN"
# ) 
# 
# new.data <- new.data %>%
#   dplyr::mutate_at("StorageMethod",
#             ~ dplyr::case_when(dplyr::row_number() %in% unanalyzed.index ~ "frozen", TRUE ~ .)) %>%
#   dplyr::mutate_at("StorageLocation",
#             ~ dplyr::case_when(dplyr::row_number() %in% unanalyzed.index ~ "Alaska MBM", TRUE ~ .)) %>%
#   dplyr::mutate_at("SampleStatus",
#             ~ dplyr::case_when(dplyr::row_number() %in% unanalyzed.index ~ NA, TRUE ~ .)) %>%
#    dplyr::mutate_at("Contact",
#             ~ dplyr::case_when(dplyr::row_number() %in% unanalyzed.index ~ NA, TRUE ~ .)) %>%
#     dplyr::select(-c("Sex",".id"))

  
write.csv(
  new.data,
  paste0(processed.data.path, new.data.name, ".csv"),
  na = "",
  row.names = FALSE
)

```

### Bander Portal

```{r}
# Provide a directory path to write processed files
processed.data.path <- "data-processed/bander_portal/"

```

```{r}

cap.data <- plyr::ldply(data.list)

# Reference tables
Master_Contacts <-
  read.csv(
    "~/Data_management/landbirds_project_maintenance/Reformatted_data/Master_Contacts.csv",
    na.strings = ""
  )

ref_data <-
  lapply(list.files(path = "reference/bander_portal/", full.names = TRUE),
         readxl::read_excel)

names(ref_data) <-
  gsub(
    list.files(path = "reference/bander_portal/"),
    pattern = ".xlsx$",
    replacement = ""
  )

list2env(ref_data, globalenv())


# Standardize missing values
cap.data <- cap.data %>%
  dplyr::mutate_if(is.character,  ~ dplyr::na_if(., "")) %>%
  dplyr::mutate_if(is.character,  ~ dplyr::na_if(., "NULL")) %>%
  dplyr::mutate_if(is.character,  ~ dplyr::na_if(., missing_values))

# Add missing columns
miss_cols <- setdiff(c(
  na.omit(capture_cols$my_col[!grepl("color", capture_cols$my_col)]),
  "TagType1",
  "TagType2",
  "TagStatus1",
  "TagStatus2"
),
colnames(cap.data))

cap.data[miss_cols] <-
  as.character(NA)

```

#### Recode data

```{r}

# How Obtained
recap.info <- cap.data %>%
  dplyr::filter(BBLDisposition == "R") %>%
  dplyr::select(SpeciesCode, BBLDisposition, BBLCondition) %>%
  dplyr::distinct()

BBL_ref <-
  bander_portal_lookups %>%
  dplyr::filter(field == "How Obtained" &
                  status != "Discontinued") %>%
  dplyr::pull(code) %>%
  noquote() %>%
  unlist()

typical.num <- which(BBL_ref == "66")
  
  if (nrow(recap.info) != 0) {
    for (a in 1:nrow(recap.info)) {
      
    condition_desc <- bander_portal_lookups %>%
      dplyr::filter(field == "Bird/Band Condition" &
                      code == recap.info$BBLCondition[a]) %>%
      dplyr::pull(description) %>%
      unlist()
    
    obtained.choice <- 0
    while (obtained.choice == 0) {
      obtained.choice <-
        utils::menu(c(
          bander_portal_lookups %>%
            dplyr::filter(field == "How Obtained" &
                            status != "Discontinued") %>%
            dplyr::pull(description)
        ),
        title = cat(
          paste0(
            "\nThe bird/band condition of one or more recaptured ",
            recap.info$SpeciesCode[a] ,
            " was '",
            condition_desc,
            "'.\nHow were these birds obtained?\nSelect ",
            typical.num,
            " for typical recaptures (i.e., bird/band condition 'ALIVE - RELEASED/LEFT ON BIRD')\n"
          )
        ))
    }
    
    cap.data <- cap.data %>%
      dplyr::rowwise() %>%
      dplyr::mutate_at(
        "HowObtained",
        ~ ifelse(
          BBLDisposition == "R" &
            SpeciesCode == recap.info$SpeciesCode[a] &
            BBLCondition == recap.info$BBLCondition[a],
          cap.data$HowObtained[a] <- BBL_ref[obtained.choice],
          
          .
        )
      )
  }
}
  

# Locations
loc.cols <- c("City", "Area")
loc.cols <- loc.cols[loc.cols %in% colnames(cap.data)]

cap.data <- cap.data %>%
  dplyr::mutate_at(
    dplyr::vars(which(names(.) %in% loc.cols)),
    ~ as.character(.) %>% dplyr::recode(
      .,
      "Eielson" = "EIEL",
      "Delta Junction" = "DELTA",
      "Eareckson" = "EARECK",
      "Beluga" = "BELUGA",
      "King Salmon" = "KING",
      "Nome" = "NOME",
      "Shemya" = "EARECK",
      "Anchorage" = "ANC",
      "Chicken" = "CHICKEN",
      "Fort Yukon" = "FORTYUK"
    )
  ) %>%
  tidyr::unite(
    LocationID,
    all_of(loc.cols),
    sep = "",
    remove = TRUE,
    na.rm = TRUE
  )

BBL_locations <- c(
  "EIEL",
  "DELTA",
  "EARECK",
  "BELUGA",
  "KING",
  "NOME",
  "EARECK",
  "ANC",
  "JBER",
  "OTTER",
  "CHICKEN",
  "FORTYUK"
)

loc.num <- which(!cap.data$LocationID %in%
                   BBL_locations)

for(a in loc.num) {
  loc.choice <-
    utils::menu(c(BBL_locations, "Enter a new location"),
                title = cat(
                  paste0(
                    "\n'",
                    cap.data$LocationID[a],
                    "' is not a listed BBL location ID.\nChoose an existing or new ID.\n"
                  )
                ))
  
  if (loc.choice == length(BBL_locations) + 1) {
    message(cat(
      paste0(
        "\nProvide a new location ID.\nThis will need to be added as a banding location in Bander Potal."
      )
    ))
    
    new_loc <-
      noquote(as.character(readline(prompt =)))
    
    
  } else {
    new_loc <- BBL_locations[loc.choice]
  }
  
  if (new_loc != "") {
    if (is.na(cap.data$LocationID[a])) {
      cap.data <-  cap.data %>%
        dplyr::mutate_at("LocationID",
                         ~ replace(.,
                                   is.na(.), new_loc))
    } else {
      cap.data <-  cap.data %>%
        dplyr::mutate_at("LocationID",
                         ~ replace(.,
                                   . == cap.data$LocationID[a], new_loc))
    }
  }
}

# Capture method
cap.data <- cap.data %>%
  dplyr::mutate_at(
    "CaptureMethod",
    ~ as.character(.) %>%
      dplyr::recode(
        .,
        "MN" = "Mist net",
        "WT" = "Walk-in trap",
        "HA" = "Hand capture",
        "BN" = "Bow net",
        "CN" = "Cannon net",
        "WN" = "Whoosh net",
        "NG" = "Net guns"
      )
  )

# Bander/Scribe ID
cap.data <- cap.data %>%
  dplyr::mutate_at(
    c("BanderID", "ScribeID"),
    ~ as.character(.) %>% plyr::mapvalues(
      .,
      from = Master_Contacts$FWSCode_FMLast,
      to = Master_Contacts$FWSCode_FML,
      warn_missing = FALSE
    )
  ) %>%
  dplyr::mutate_at(
    c("BanderID", "ScribeID"),
    ~ plyr::mapvalues(
      .,
      from = Master_Contacts$FWSCode_FLast,
      to = Master_Contacts$FWSCode_FL,
      warn_missing = FALSE
    )
  )

# Sex
cap.data <- cap.data %>%
  dplyr::mutate_at("FieldSex", ~ as.character(.)) %>%
  dplyr::rowwise() %>%
  dplyr::mutate_at("FieldSex",
                   ~ ifelse(
                     BBLDisposition == "R",
                     dplyr::recode(., "M" = "6", "F" = "7", "U" = "0"),
                     dplyr::recode(., "M" = "4", "F" = "5", "U" = "0")
                   ))


# Molt
cap.data <- cap.data %>%
  dplyr::mutate_at(
    "BodyMolt",
    ~ as.character(.) %>% dplyr::recode(
      .,
      "0" = "N",
      "1" = "Y",
      "2" = "Y",
      "3" = "Y",
      "4" = "Y"
    )
  ) %>%
  dplyr::mutate_at(
    "FlightFeatherMolt",
    ~ as.character(.) %>% dplyr::recode(
      .,
      "N" = "N",
      "S" = "Y",
      "A" = "Y",
      "J" = "Y"
    )
  )


# Leg attachments
leg_cols <-   c(
  "ULIn",
  "LLIn",
  "URIn",
  "LRIn",
  "ULOut",
  "LLOut",
  "UROut",
  "LROut"
)

leg_cols <- leg_cols[leg_cols %in% colnames(cap.data)]

color_cols<- paste0(leg_cols,"_colorband")


# test <- cap.data %>%
#   dplyr::filter_at(leg_cols, dplyr::any_vars(stringr::str_detect(., stringr::regex("-", ignore_case =
#                                                                                      TRUE))))

my_colors <-
  c(
    "Bk",
    "Br",
    "Db",
    "Lb",
    "Mb",
    "Dg",
    "Lg",
    "Gy",
    "Hp",
    "Mg",
    "Mv",
    "Or",
    "Pp",
    "Rd",
    "Wh",
    "Yl"
  )

BBL_colors <-
  c(
    "Black",
    "Brown",
    "Dark Blue",
    "Light Blue",
    "Blue",
    "Dark Green",
    "Light Green",
    "Gray",
    "Hot Pink",
    "Magenta",
    "Mauve",
    "Orange",
    "Purple",
    "Red",
    "White",
    "Yellow"
  )

cap.data <-  cap.data %>%
  dplyr::mutate_if(is.logical, as.factor) %>%
  dplyr::rowwise() %>%
  dplyr::mutate_at(
    c(leg_cols, "FlagColorIn", "FlagColorOut"),
    ~ stringi::stri_replace_all_fixed(.,
                                      my_colors,
                                      BBL_colors,
                                      vectorize_all = FALSE)) %>%
  dplyr::mutate_at(leg_cols,
                   ~ stringi::stri_replace_all_fixed(.,
                                                     "-",
                                                     "/",
                                                     vectorize_all = FALSE)) %>%
  dplyr::mutate(across(.cols = all_of(leg_cols),
                       .names = "{paste0(color_cols[!grepl('Flag', color_cols)])}")) %>%
  dplyr::mutate_at(
    leg_cols,
    ~ stringi::stri_replace_all_fixed(.,
                                      BBL_colors,
                                      "01A",
                                      vectorize_all = FALSE)) %>%
  dplyr::mutate_at(leg_cols,
                   ~ stringi::stri_replace_all_fixed(.,
                                                     c("X", "Flag"),
                                                     c("00", "69"),
                                                     vectorize_all = FALSE)) %>%
  dplyr::mutate_at(
    color_cols,
    ~ ifelse(
      grepl("/", .),
      stringi::stri_replace_all_fixed(.,
                                      c("X", "Geo", "Flag"),
                                      "", vectorize_all = FALSE),
      stringi::stri_replace_all_fixed(.,
                                      c("X", "Geo", "Flag"),
                                      NA, vectorize_all = FALSE)
    )) %>%
  dplyr::mutate_at(color_cols,
                   ~ ifelse(
                     grepl("/$", .),
                     stringi::stri_replace_all_fixed(.,
                                                     "/",
                                                     "", vectorize_all = FALSE),
                     .
     
              ))


## Flag code colors by species - new/retained/replaced
flag_info <- cap.data %>%
  dplyr::filter(dplyr::if_any(c(leg_cols),
                              ~ stringr::str_detect(., "69")))  %>%
  dplyr::rowwise() %>%
  dplyr::mutate(
    FlagCodeChange = ifelse(FlagCodeIn != FlagCodeOut, 1, 0),
    FlagIn = ifelse(is.na(FlagCodeIn) == FALSE, 1, 0),
    FlagOut = ifelse(is.na(FlagCodeOut) == FALSE, 1, 0)
  ) %>%
  dplyr::select(any_of(c(
    "SpeciesCode", "FlagIn", "FlagOut", "FlagCodeChange"
  ))) %>%
  dplyr::distinct() %>%
  dplyr::mutate_all(as.character) %>%
  dplyr::mutate_at("FlagCodeChange", ~ replace(., is.na(.), 0))

BBL_ref <- bander_portal_lookups$code[bander_portal_lookups$field ==
                                        "Marker color"]
BBL_ref<- BBL_ref[!grepl("/", BBL_ref)]

cap.data[, "FlagCodeIn_color"] <- NA
cap.data[, "FlagCodeOut_color"] <- NA

if(nrow(flag_info)!=0) {
  for (b in 1:nrow(flag_info)) {
    flag_status <-
      ifelse(flag_info$FlagIn[b] == 0,
             "received a new",
             "were recaptured with a")
    
    flag_status <-
      ifelse(flag_info$FlagCodeChange[b] == 0,
             flag_status,
             "received a replacement")
    
    flag_status.2 <-
      ifelse(flag_info$FlagCodeChange[b] == 0, "", "original ")
    
    
    flag.choice <- 0
    while (flag.choice == 0) {
      flag.choice <-
        utils::menu(c(BBL_ref), title = cat(
          paste0(
            "\nOne or more ",
            flag_info$SpeciesCode[b] ,
            " ",
            flag_status,
            " flag.\nWhat is the ",
            flag_status.2 ,
            "flag code color?\n"
          )
        ))
    }
    
    if (flag_info$FlagIn[b] == 0) {
      ### new
      cap.data$FlagCodeOut_color[cap.data$SpeciesCode == flag_info$SpeciesCode[b] &
                                   is.na(cap.data$FlagCodeIn) &
                                   !is.na(cap.data$FlagCodeOut)] <-
        BBL_ref[flag.choice]
    } else {
      if (flag_info$FlagCodeChange[b] == 1) {
        flag.choice.2 <- 0
        while (flag.choice.2 == 0) {
          flag.choice.2 <-
            utils::menu(c(BBL_ref), title = cat(paste0(
              "\nWhat is the new flag code color?\n"
            )))
        }
        
        ### replaced
        cap.data$FlagCodeIn_color[cap.data$SpeciesCode == flag_info$SpeciesCode[b] &
                                    !is.na(cap.data$FlagCodeIn) &
                                    cap.data$FlagCodeIn != cap.data$FlagCodeOut &
                                    !is.na(cap.data$FlagCodeOut)] <-
          BBL_ref[flag.choice]
        
        cap.data$FlagCodeOut_color[cap.data$SpeciesCode == flag_info$SpeciesCode[b] &
                                     !is.na(cap.data$FlagCodeIn) &
                                     cap.data$FlagCodeIn != cap.data$FlagCodeOut &
                                     !is.na(cap.data$FlagCodeOut)] <-
          BBL_ref[flag.choice.2]
        
      } else {
        ### retained
        cap.data$FlagCodeIn_color[cap.data$SpeciesCode == flag_info$SpeciesCode[b] &
                                    !is.na(cap.data$FlagCodeIn) &
                                    cap.data$FlagCodeIn == cap.data$FlagCodeOut] <-
          BBL_ref[flag.choice]
        
        cap.data$FlagCodeOut_color[cap.data$SpeciesCode == flag_info$SpeciesCode[b] &
                                     !is.na(cap.data$FlagCodeIn) &
                                     cap.data$FlagCodeIn == cap.data$FlagCodeOut] <-
          BBL_ref[flag.choice]
      }
      
    }
  }
}


# Tags
tag_types <- c("Geo", "Radio", "GPS", "PTT")
tag_info <- cap.data %>%
  dplyr::filter_at(c("TagType1", "TagType2"),
                   dplyr::any_vars(stringr::str_detect(
                     ., stringr::regex(paste(tag_types, collapse = '|'), ignore_case =
                                         TRUE)
                   ))) %>%
  tidyr::pivot_longer(cols = starts_with("TagType"),
                      names_to = "TagType.col",
                      values_to = "TagType") %>%
  tidyr::pivot_longer(
    cols = starts_with("TagStatus"),
    names_to = "TagStatus.col",
    values_to = "TagStatus"
  ) %>%
  dplyr::filter(readr::parse_number(TagStatus.col) == readr::parse_number(TagType.col)) %>%
  dplyr::filter(!is.na(TagType)) %>%
  dplyr::select(SpeciesCode,
                TagType,
                TagStatus) %>%
  dplyr::distinct()

if(nrow(tag_info)!=0){
  for(a in 1:nrow(tag_info)) {
    BBL_tags <-
      unlist(strsplit(capture_recode$BBL_code[capture_recode$my_code == tag_info$TagType[a]], ","))
    
    BBL_ref <- bander_portal_lookups %>%
      dplyr::filter(field == "Marker Type", code %in% BBL_tags) %>%
      dplyr::pull(description) %>%
      stringr::str_extract_all("(?<=\\().+?(?=\\))") %>%
      unlist()
    
    tag.choice <- 0
    while (tag.choice == 0) {
      tag.choice <-
        utils::menu(c(BBL_ref), title = cat(
          paste0(
            "\nA ",
            tolower(tag_info$TagType[a]),
            " tag was ",
            tolower(tag_info$TagStatus[a]),
            " on one or more ",
            tag_info$SpeciesCode[a],
            ".\nHow was this marker attached?\n"
          )
        ))
    }
    
    cap.data <- cap.data %>%
      dplyr::rowwise() %>%
      dplyr::mutate_at(
        c(leg_cols, "TagType1"),
        ~ ifelse(
          !is.na(TagType1) &
            SpeciesCode == tag_info$SpeciesCode[a] &
            TagStatus1 == tag_info$TagStatus[a],
          stringi::stri_replace_all_regex(., tag_info$TagType[a], BBL_tags[tag.choice], vectorize_all = FALSE),
          
          .
        )
      ) %>%
      dplyr::mutate_at(
        c(leg_cols, "TagType2"),
        ~ ifelse(
          !is.na(TagType2) &
            SpeciesCode == tag_info$SpeciesCode[a] &
            TagStatus2 == tag_info$TagStatus[a],
          stringi::stri_replace_all_regex(., tag_info$TagType[a], BBL_tags[tag.choice], vectorize_all = FALSE),
          .
        )
      )
  }
}


```

#### Quality checks

```{r}

# Species measurements
if(class(bander_portal_species_measurements$bandSize)!= "list") {
  bander_portal_species_measurements <-
    bander_portal_species_measurements %>%
    dplyr::mutate_if(is.numeric, ~ dplyr::na_if(., 0)) %>%
    dplyr::rowwise() %>%
    dplyr::mutate_at("bandSize", ~ ifelse(grepl(", ", .), as.list(strsplit(., ", ")), as.list(.)))
}


mydata_col <- c("Mass_g", "Wing_mm", "Tail_mm")
BBL_col <-
  c(
    "femaleMass_min_g",
    "femaleMass_max_g",
    "maleMass_min_g",
    "maleMass_max_g",
    "femaleWingChord_min_mm",
    "femaleWingChord_max_mm",
    "maleWingChord_min_mm",
    "maleWingChord_max_mm",
    "femaleTailLength_min_mm",
    "femaleTailLength_max_mm",
    "maleTailLength_min_mm",
    "maleTailLength_max_mm"
  )

for (x in unique(cap.data$SpeciesCode)) {
  species.num <-
    which(bander_portal_species_measurements$alphaCode == x)
  if (length(species.num) == 0) {
    warning(paste0("SpeciesCode '", x, "' is invalid."))
  } else {
    sp_mydata <- cap.data %>%
      dplyr::filter(SpeciesCode == x) %>%
      dplyr::rowwise() %>%
      dplyr::mutate_at("FieldSex",
                       ~ dplyr::recode(
                         .,
                         "4" = "M",
                         "5" = "F",
                         "6" = "M",
                         "7" = "F" ,
                         "0" = "U"
                       ))
    
    sp_BBLdata <- bander_portal_species_measurements %>%
      dplyr::filter(alphaCode == x)
    
    ### Band size
    if (length(setdiff(unique(na.omit(
      sp_mydata$BandSize
    )), unlist(na.omit(
      sp_BBLdata$bandSize
    ))))
    != 0) {
      warning(
        paste0(
          "One or more ",
          sp_BBLdata$commonName,
          " has a band size inconsistent with BBL recommendations."
        )
      )
    }
    
    ## Mass/Wing chord/Tail
    for (a in c("Mass", "Wing", "Tail")) {
      for (b in  c("F", "M", "U")) {
        for (c in c("min", "max")) {
          mydata_col.match <- mydata_col[grep(a, mydata_col)]
          
          if (b == "U") {
            BBL_col.match <-
              BBL_col[grepl(a, BBL_col) & grepl(c, BBL_col)]
            sex_code <- c("U", NA)
            sex <- "unknown sex"
            
          } else {
            BBL_col.match <-
              BBL_col[grepl(a, BBL_col) &
                        grepl(paste0("^", b), BBL_col, ignore.case = TRUE) &
                        grepl(c, BBL_col)]
            sex_code <-  b
            sex <-
              grep(
                paste0("^", b),
                c("female", "male"),
                ignore.case = TRUE,
                value = TRUE
              )
            
          }
          fun <- match.fun(c)
          sign <- c("<", ">")[grep(c, c("min", "max"))]
          comparative <-
            c("smaller", "greater")[grep(c, c("min", "max"))]
          measurement <-
            c("mass", "wing chord", "tail")[grep(a, c("Mass", "Wing", "Tail"))]
          
          
          if (is.na(sum(sp_BBLdata[which(colnames(sp_BBLdata) %in% BBL_col.match)])) == FALSE) {
            if (length(na.omit(sp_mydata[[mydata_col.match]][sp_mydata$FieldSex %in% sex_code])) !=
                0) {
              if (get(sign)(fun(sp_BBLdata[which(colnames(sp_BBLdata) %in% BBL_col.match)]), fun(na.omit(sp_mydata[[mydata_col.match]][sp_mydata$FieldSex %in% sex_code]))) ==
                  FALSE)
              {
                {
                  warning(
                    paste0(
                      "One or more ",
                      sp_BBLdata$commonName,
                      " (",
                      sex,
                      ") have a ",
                      comparative,
                      " ",
                      measurement,
                      " than expected for the species (",
                      c,
                      ": ",
                      fun(sp_BBLdata[which(colnames(sp_BBLdata) %in% BBL_col.match)]),
                      " ",
                      sub('.*\\_', '', BBL_col.match[1]),
                      ")"
                    )
                  )
                }
              }
            }
          }
        }
      }
    }
  }
}


# lookup tables

lookups <- c("Age",
             "Sex",
             "How Aged",
             "How Sexed",
             "Marker Type",
             "Marker color")

for (a in lookups) {
  b <- which(capture_cols$BBL_col[capture_cols$BBL_template=="band_aux"] == a)
  for (c in b) {
    invalid.num <-
      which(!na.omit(cap.data[[capture_cols$my_col[c]]]) %in% bander_portal_lookups$code[bander_portal_lookups$field == a])
    
    invalid.index <-
      setdiff(1:nrow(cap.data), which(is.na(cap.data[[capture_cols$my_col[c]]])))[invalid.num]
    
    entry.skip <- c()
    
    if (length(invalid.index) != 0) {
      for (d in invalid.index) {
        if (d %in% entry.skip) {
          next
        } else{
          entry.choice <-
            utils::menu(c("replace these entries",
                          "do nothing"),
                        title = cat(
                          paste0(
                            "\n'",
                            cap.data[[capture_cols$my_col[c]]][d],
                            "' (",
                            capture_cols$my_col[c],
                            ") is not listed as a BBL ",
                            a,
                            "\nHow would you like to proceed?"
                          )
                        ))
          
          descriptions <- bander_portal_lookups$description[bander_portal_lookups$field == a]
          if (all(is.na(descriptions))) {
            descriptions <- ""
          } else {
            paste0(" - ", descriptions)
          }
          
          if (entry.choice == 1) {
            replace.choice <-
              utils::menu(c(
                paste0(
                  bander_portal_lookups$code[bander_portal_lookups$field == a],
                  descriptions
                )
              ),
              title = cat(paste0(
                "\nChoose a replacement value."
              )))
            
            if (replace.choice != 0) {
              entry.skip <- c(entry.skip,
                              which(cap.data[[capture_cols$my_col[c]]] == cap.data[[capture_cols$my_col[c]]][d]))
              
              cap.data[[capture_cols$my_col[c]]] <-
                replace(cap.data[[capture_cols$my_col[c]]],
                        cap.data[[capture_cols$my_col[c]]] %in% cap.data[[capture_cols$my_col[c]]][d],
                        bander_portal_lookups$code[bander_portal_lookups$field == a][replace.choice])
            }
            
          } else if (entry.choice == 2) {
            entry.skip <- c(entry.skip, which(cap.data[[capture_cols$my_col[c]]] == cap.data[[capture_cols$my_col[c]]][d]))
          }
        }
      }
    }
  }
}

# Bird status
## Status codes
status_info <- cap.data %>%
  dplyr::rowwise() %>%
  dplyr::mutate_at("TagType1", ~ ifelse(TagStatus1 %in% c("Removed", "Lost"), TagType1 <-
                                          NA
                                        , .)) %>%
  dplyr::mutate_at("TagType2", ~ ifelse(TagStatus2  %in% c("Removed", "Lost"), TagType2 <-
                                          NA
                                        , .)) %>%
  dplyr::select(
    "OccurenceID",
    "BandNumberOut",
    "SpeciesCode",
    "BBLStatus",
    leg_cols[grepl("Out", leg_cols)],
    "TagType1",
    "TagType2",
    "Blood",
    "BBLStatus"
  ) %>%
  dplyr::mutate_at(dplyr::vars(c(
    leg_cols[grepl("Out", leg_cols)], "TagType1", "TagType2", "BBLStatus"
  )), ~as.character(.) %>% dplyr::na_if(.,"")) %>%
  dplyr::mutate(NumMarkers = sum(!unique(dplyr::c_across(
    c(leg_cols[grepl("Out", leg_cols)], "TagType1", "TagType2")
  ))  %in% c("X", NA))) %>%
  dplyr::mutate_at(
    "BBLStatus",
    ~ dplyr::case_when(
      NumMarkers >= 2 &
        Blood != "Y" ~ "325",
      NumMarkers >= 1 &
        Blood == "Y" ~ "319",
      NumMarkers == 0 &
        Blood == "Y" ~ "318",
      NumMarkers == 0 &
        Blood != "Y" ~ "300",
      NumMarkers == 1 &
        dplyr::if_any(c(leg_cols[grepl("Out", leg_cols)], "TagType1", "TagType2"), ~
                        . == "69") &
        Blood != "Y" ~ "369",
      NumMarkers == 1 &
        dplyr::if_any(c(leg_cols[grepl("Out", leg_cols)], "TagType1", "TagType2"), ~
                        grepl("^81", .)) &
        Blood != "Y" ~ "381",
      NumMarkers == 1 &
        dplyr::if_any(c(leg_cols[grepl("Out", leg_cols)], "TagType1", "TagType2"), ~
                        grepl("^80", .)) &
        Blood != "Y" ~ "380",
      NumMarkers == 1 &
        dplyr::if_any(c(leg_cols[grepl("Out", leg_cols)], "TagType1", "TagType2"), ~
                        grepl("^90", .)) &
        Blood != "Y" ~ "390",
      NumMarkers == 1 &
        dplyr::if_any(c(leg_cols[grepl("Out", leg_cols)], "TagType1", "TagType2"), ~
                        grepl("^01", .)) &
        Blood != "Y" ~ "301",
      TRUE ~ BBLStatus
    )
  ) %>%
  dplyr::rename(`temp:BBLStatus` = BBLStatus)

cap.data <- cap.data %>%
  dplyr::left_join(
    .,
    status_info[, c("OccurenceID", "temp:BBLStatus")],
    by = c("OccurenceID" = "OccurenceID"),
    keep = FALSE,
    relationship = "one-to-one"
  ) %>%
  dplyr::mutate_at("BBLStatus", ~ as.character(.))

invalid.index <- which(cap.data$BBLStatus != cap.data$`temp:BBLStatus`)


if (length(invalid.index) != 0) {
  for (a in invalid.index) {
    message(cat(
      paste0(
        "The status for the following individual appears invalid.\nSpecies: ",
        cap.data$SpeciesCode[a],
        "\nBand number: ",
        cap.data$BandNumberOut[a],
        "\nStatus (reported): ",
        cap.data$BBLStatus[a],
        " [",
        bander_portal_lookups$description[grepl(stringr::str_sub(cap.data$BBLStatus[a], 1, 1),
                                                bander_portal_lookups$code) == TRUE &
                                            bander_portal_lookups$field == "Bird Status"],
        " ",
        bander_portal_lookups$description[grepl(stringr::str_sub(cap.data$BBLStatus[a], 2, 3),
                                                bander_portal_lookups$code) == TRUE &
                                            bander_portal_lookups$field == "Bird Status Extra Info"],
        "]",
        "\nStatus (detected): ",
        cap.data$`temp:BBLStatus`[a],
        " [",
        bander_portal_lookups$description[grepl(
          stringr::str_sub(cap.data$`temp:BBLStatus`[a], 1, 1),
          bander_portal_lookups$code
        ) == TRUE &
          bander_portal_lookups$field == "Bird Status"],
        " ",
        bander_portal_lookups$description[grepl(
          stringr::str_sub(cap.data$`temp:BBLStatus`[a], 2, 3),
          bander_portal_lookups$code
        ) == TRUE &
          bander_portal_lookups$field == "Bird Status Extra Info"],
        "]",
        "\nBlood: ",
        cap.data$Blood[a],
        "\nULOut: ",
        cap.data$ULOut[a],
        "\nLLOut: ",
        cap.data$LLOut[a],
        "\nUROut: ",
        cap.data$UROut[a],
        "\nLROut: ",
        cap.data$LROut[a],
        "\nTagStatus1: ",
        cap.data$TagStatus1[a],
        "\nTagStatus2: ",
        cap.data$TagStatus2[a]
      )
    ))
    
    status.choice <- utils::menu(
      choice = c("update as detected", "manually edit", "leave as is"),
      graphics = TRUE,
      title = cat("\nHow would you like to proceed?")
    )
    
    if (status.choice == 1) {
      cap.data$BBLStatus[a] <- cap.data$`temp:BBLStatus`[a]
      
      
    }
    if (status.choice == 2) {
      message(
        "\nIMPROTANT: you must click 'syncronise' (arrow loop icon) and then 'Done' to enact changes."
      )
      
      invalid.entries <- cap.data %>%
        dplyr::ungroup() %>%
        dplyr::filter(dplyr::row_number() == a)
      
      invalid.entries.edit <-
        DataEditR::data_edit(
          invalid.entries,
          col_readonly = colnames(invalid.entries)[!colnames(invalid.entries) %in% "BBLStatus"],
          viewer = "pane"
        ) %>%
        dplyr::mutate_at("BBLStatus", as.character)
      
      cap.data[match(invalid.entries.edit$OccurenceID,
                     cap.data$OccurenceID), "BBLStatus"] <-
        invalid.entries.edit$BBLStatus
    }
    
  }
}


cap.data <-
  cap.data %>%
  dplyr::select(-tidyselect::starts_with("temp:"))


# Other
## Band number
for(a in c("In", "Out")) {
  invalid_code.num <-
    which(!c(nchar(na.omit(
      gsub("-", "", cap.data[[paste0("BandNumber", a)]])
    ))) %in% c(8, 9))
  
  dir <-
    c("incoming", "outgoing")[grep(a, c("BandNumberIn", "BandNumberOut"))]
  
  if (length(invalid_code.num) != 0) {
    warning(
      paste0(
        "The following ",
        dir,
        " band numbers appear invalid. BBL issued federal metal bands should have 8 or 9 digits.\n"
      ),
      paste0(na.omit(cap.data[[paste0("BandNumber", a)]])[invalid_code.num], sep = "\n")
    )
  }
}

invalid_bands <- cap.data %>%
  dplyr::group_by(BandNumberOut, BBLDisposition) %>%
  dplyr::filter(BBLDisposition == "1",
                dplyr::n() > 1)

for(a in unique(invalid_bands$BandNumberOut)) {
  unique(invalid_bands$SpeciesCode[invalid_bands$BandNumberOut == a])
  warning(
    paste0(
      "Band number ",
      a,
      " is duplicated for the following species/flag codes.\n"
    ),
    paste0(
      paste0(
        invalid_bands$SpeciesCode[invalid_bands$BandNumberOut == a],
        " (",
        invalid_bands$FlagCodeOut[invalid_bands$BandNumberOut == a],
        ")"
      ),
      sep = "\n"
    )
  )
}

## Date
if(NA %in% cap.data$Day |
   NA %in% cap.data$Month |
   NA %in% cap.data$Year) {
  if ("Date" %in% colnames(cap.data)) {
    if (!NA %in% suppressWarnings(chron::chron(cap.data$Date, format = gsub("%", "",date.format)))) {
      
      date.choice <-
        utils::menu(c("please do", "absolutely not"),
                    title = cat(
                      paste0(
                        "\nThe variables Day, Month, and/or Year are incomplete. Would you like to parse this information from Date?\n"
                      )
                    ))
    
      if(date.choice==1) {
        cap.data <-
          cap.data %>%
          dplyr::mutate(
            Date = as.Date(Date, date.format),
            Day = lubridate::day(Date),
            Month = lubridate::month(Date),
            Year = lubridate::year(Date)
          )
      }
    } else {
      warning("The variables Day, Month, and/or Year are incomplete.")
    }
  } else {
    warning("The variables Day, Month, and/or Year are incomplete.")
  }
}

## Time
time.num <-
  which(is.na(lubridate::hm(cap.data$CaptureTime_2400, quiet = TRUE)) |
          grepl("[A-Za-z]", cap.data$CaptureTime_2400) == TRUE)

if (length(time.num) != 0) {
  warning(
    paste0(
      "One or more time entries (CaptureTime_2400) are invalid. Time must be formatted as 24:00.\n"
    ),
    paste0(unique(cap.data$CaptureTime_2400[time.num]), sep = "\n")
  )
}

              
## Bander ID/Scribe ID
for (a in c("BanderID", "ScribeID")) {
  invalid_code.num <- which(nchar(unique(cap.data[[a]])) > 3)
  
  for (b in invalid_code.num) {
    message(cat(
      paste0(
        "'",
        cap.data[[a]][b],
        "' (",
        a,
        ") appears invalid.\nProvide a replacement value. Press Enter to retain the current value."
      )
    ))
    
    new_code <-
      noquote(as.character(readline(prompt =)))
    
    if (new_code != "") {
      cap.data <-  cap.data %>%
        dplyr::mutate_at(c("BanderID", "ScribeID"),
                         ~ replace(., . == cap.data[[a]][b], new_code))
    }
  }
}

```

#### Translate data

```{r}

# Partition by disposition
cap.data_aux <-
  cap.data %>% dplyr::filter(dplyr::if_any(leg_cols, ~ !. %in% c(NA, "00")))

cap.data_no_aux <-
  cap.data %>% dplyr::filter(dplyr::if_all(leg_cols, ~ . %in% c(NA, "00")))

cap.data_aux <- split(cap.data_aux, cap.data_aux$BBLDisposition)
cap.data_no_aux <- split(cap.data_no_aux, cap.data_no_aux$BBLDisposition)

#!!# Improve process for no data
#!!# 8 just reports band number in band_no_aux
for (a in c("cap.data_aux", "cap.data_no_aux")) {
  for (b in c("1", "R")) {
    if (!b %in% names(eval(as.name(a)))) {
      temp <-  list(rbind(tibble::as_tibble(
        matrix(
          ncol = length(cap.data),
          nrow = 1,
          dimnames = list(NULL, colnames(cap.data))
        ), check.names = FALSE
      )))
      
      names(temp) <- b
      
      temp<- c(eval(as.name(a)),temp)
      
      assign(a, temp)
      
    }
  }
}


# Merge 1, 5, & 8 dispositions
cap.data_no_aux$'1' <-
  tibble::as_tibble(data.table::rbindlist(cap.data_no_aux[c('1', '5', '8')]))
cap.data_no_aux <- suppressWarnings(within(cap.data_no_aux, rm('5','8')))
cap.data_aux$'1' <-
  tibble::as_tibble(data.table::rbindlist(cap.data_aux[c('1', '5', '8')]))
cap.data_aux <- suppressWarnings(within(cap.data_aux, rm('5', '8')))

## Correct variable classes
cap.data_aux <- lapply(cap.data_aux, function(x) {
  x %>%
    dplyr::mutate_if(is.character, ~ dplyr::na_if(., ''))
})
cap.data_no_aux <- lapply(cap.data_no_aux, function(x) {
  x %>%
    dplyr::mutate_if(is.character, ~ dplyr::na_if(., ''))
})


# Complete templates
# Create empty dataframes
template_filenames <-
  list.files("templates/bander_portal",
             pattern = ".xlsx",
             full.names = TRUE)
template_colnames <-
  lapply(template_filenames, function(x)
    suppressWarnings(colnames(readxl::read_excel(x))))
names(template_colnames) <- basename(template_filenames)


band_no_aux_col <-
  unlist(template_colnames[names(template_colnames)
                    [grepl("band", names(template_colnames), ignore.case = TRUE) &
                     grepl("no", names(template_colnames), ignore.case = TRUE)]])
band_no_aux <-
  data.frame(matrix(
    ncol = length(band_no_aux_col),
    nrow = nrow(cap.data_no_aux$'1'),
    dimnames = list(NULL, band_no_aux_col)
  ),
  check.names = FALSE)

recap_no_aux_col <-
  unlist(template_colnames[names(template_colnames)
                    [grepl("recap", names(template_colnames), ignore.case = TRUE) &
                     grepl("no", names(template_colnames), ignore.case = TRUE)]])
recap_no_aux <-
  data.frame(matrix(
    ncol = length(recap_no_aux_col),
    nrow = nrow(cap.data_no_aux$'R'),
    dimnames = list(NULL, recap_no_aux_col)
  ),
  check.names = FALSE)

band_aux_col <-
  unlist(template_colnames[names(template_colnames)
                    [grepl("band", names(template_colnames), ignore.case = TRUE) &
                     grepl("aux", names(template_colnames), ignore.case = TRUE)]])
band_aux <-
  data.frame(matrix(
    ncol = length(band_aux_col),
    nrow = nrow(cap.data_aux$'1'),
    dimnames = list(NULL, band_aux_col)
  ),
  check.names = FALSE)

recap_aux_col <-
  unlist(template_colnames[names(template_colnames)
                    [grepl("recap", names(template_colnames), ignore.case = TRUE) &
                     grepl("aux", names(template_colnames), ignore.case = TRUE)]])
recap_aux <-
  data.frame(matrix(
    ncol = length(recap_aux_col),
    nrow = nrow(cap.data_aux$'R'),
    dimnames = list(NULL, recap_aux_col)
  ),
  check.names = FALSE)

band_no_aux_ref<- capture_cols %>% dplyr::filter(BBL_template=="band_no_aux")
recap_no_aux_ref<- capture_cols %>% dplyr::filter(BBL_template=="recap_no_aux")
band_aux_ref<- capture_cols %>% dplyr::filter(BBL_template=="band_aux")
recap_aux_ref<- capture_cols %>% dplyr::filter(BBL_template=="recap_aux")



## Fill banding - no aux
auto.match.num <-
  which(band_no_aux_ref$BBL_col %in% colnames(band_no_aux) &
          !is.na(band_no_aux_ref$my_col))


for (a in na.omit(auto.match.num)) {
  band_no_aux[[band_no_aux_ref$BBL_col[a]]] <-
    cap.data_no_aux$'1'[[band_no_aux_ref$my_col[a]]]
}

band_no_aux[["Replaced Band Number"]] <-
  unlist(cap.data_no_aux$'1'["BandNumberIn"])

banded_legs <- c()
banded_legs <-
  c(banded_legs, sapply(1:nrow(cap.data_no_aux$'1'), function(x)
    substring(colnames(cap.data_no_aux$'1'[leg_cols])[cap.data_no_aux$'1'[x, leg_cols] %in% "00"][1], 2, 2)))

band_no_aux[["Banded Leg"]]<- banded_legs

## Fill recaptures - no aux
auto.match.num <-
  which(recap_no_aux_ref$BBL_col %in% colnames(recap_no_aux) &
          !is.na(recap_no_aux_ref$my_col))


for (a in na.omit(auto.match.num)) {
  recap_no_aux[[recap_no_aux_ref$BBL_col[a]]] <-
    cap.data_no_aux$'R'[[recap_no_aux_ref$my_col[a]]]
}

banded_legs <- c()
banded_legs <-
  c(banded_legs, sapply(1:nrow(cap.data_no_aux$'R'), function(x)
    substring(colnames(cap.data_no_aux$'R'[leg_cols])[cap.data_no_aux$'R'[x, leg_cols] %in% "00"][1], 2, 2)))

recap_no_aux[["Banded Leg"]]<- banded_legs


## Fill banding - aux
auto.match.num <-
  which(band_aux_ref$BBL_col %in% colnames(band_aux) &
          !is.na(band_aux_ref$my_col))

manual.match.num <-
  which(!band_aux_ref$BBL_col %in% colnames(band_aux))


for (a in na.omit(auto.match.num)) {
  band_aux[[band_aux_ref$BBL_col[a]]] <-
    cap.data_aux$'1'[[band_aux_ref$my_col[a]]]
}


for (a in 1:nrow(cap.data_aux$'1')) {
  n <- 0
  marker_string <- c()
  for (b in c(leg_cols[grepl("Out",leg_cols)])) {
    if (!is.na(cap.data_aux$'1'[a, b])) {
      for (c in unlist(strsplit(paste0(cap.data_aux$'1'[a, b]), "/"))) {
        n <- n + 1
         marker_cols <-
            colnames(band_aux)[grepl(n, colnames(band_aux))]
        
        band_aux[[marker_cols[1]]][a] <-
          c
        
        code_col <-
          "FlagCodeOut"[sapply("69", function(x)
            grepl(x, c))]
        if (length(code_col) != 0) {
          band_aux[[marker_cols[2]]][a] <-
            cap.data_aux$'1'[[code_col]][a]
        }
        
        color_col <-
          c("FlagColorOut",
            paste0(b, "_colorband"))[sapply(c("69", "01A"), function(x)
              grepl(x, c))]
        if (length(color_col) != 0) {
          band_aux[[marker_cols[3]]][a] <-
            cap.data_aux$'1'[[color_col]][a]
        }
        
        code_color_col <-
          "FlagCodeOut_color"[sapply("69", function(x)
            grepl(x, c))]
        if (length(code_color_col) != 0) {
          band_aux[[marker_cols[4]]][a] <-
            cap.data_aux$'1'[[code_color_col]][a]
        }
        
        band_aux[[marker_cols[5]]][a] <-
          c("L", "R")[sapply(c(".{1}L", ".{1}R"), function(x)
            grepl(x, b))]
        
        band_aux[[marker_cols[6]]][a] <-
          paste0("L", c("B", "A")[sapply(c("^L", "^U"), function(x)
            grepl(x, b))])
        
        if (cap.data_aux$'1'[a, b] == "00") {
          band_aux[["Banded Leg"]][a] <-
            c("L", "R")[sapply(c(".{1}L", ".{1}R"), function(x)
              grepl(x, b))]
        }
      }
    }
    
  }
  
  for (b in c("TagType1", "TagType2")) {
    if (!is.na(cap.data_aux$'1'[a, b])) {
      n <- n + 1
      
      status <-
        cap.data_aux$'1'[a, paste0("TagStatus", c("1", "2")[sapply(c("1", "2"), function(x)
          grepl(x, b))])]
      
      if (!TRUE %in% sapply(leg_cols, function(x)
        grepl(cap.data_aux$'1'[a, b], cap.data_aux$'1'[a, x])) &
        status != "Removed") {
  
        band_aux[[colnames(band_aux)[grepl(n, colnames(band_aux))][1]]][a] <-
          unlist(cap.data_aux$'1'[a, b])
        
      }
    }
  }
  
  end.num <-
    c(27, 33, 39, 45, 51, 57)[sapply(c(1, 2, 3, 4, 5, 6), function(x)
      grepl(x, n))]
  
  marker_string <-
    tidyr::replace_na(unlist(band_aux[a, 22:end.num]), "")
  
  for (b in
       names(marker_string)[grepl("Side", names(marker_string))]) {
    marker_string <-
      append(marker_string, c("", ""), after = which(names(marker_string) == b))
  }
  
  band_aux[['All Marker Info At Release']][a] <-
    paste0(paste0(marker_string, collapse = "|"), "|")
}

band_aux[["Replaced Band Number"]]<-  unlist(cap.data_aux$'1'["BandNumberIn"])


## Fill recaptures - aux
auto.match.num <-
  which(recap_aux_ref$BBL_col %in% colnames(recap_aux) &
          !is.na(recap_aux_ref$my_col))

for (a in na.omit(auto.match.num)) {
  recap_aux[[recap_aux_ref$BBL_col[a]]] <-
    cap.data_aux$'R'[[recap_aux_ref$my_col[a]]]
}


for (a in 1:nrow(cap.data_aux$'R')) {
  in.num <- 0
  out.num <- 0
  
  for (b in leg_cols) {
    if (!is.na(cap.data_aux$'R'[a, b])) {
      for (c in unlist(strsplit(paste0(cap.data_aux$'R'[a, b]), "/"))) {
        if (grepl("In", b)) {
          in.num <- in.num + 1
          n <- in.num
          marker_cols <-
            colnames(recap_aux)[grepl(n, colnames(recap_aux)) &
                                  grepl("capture", colnames(recap_aux))]
          
        } else if (grepl("Out", b)) {
          out.num <- out.num + 1
          n <- out.num
          marker_cols <-
            colnames(recap_aux)[grepl(n, colnames(recap_aux)) &
                                  grepl("release", colnames(recap_aux))]
        }
        
        recap_aux[[marker_cols[1]]][a] <-
          c
        
        dir <-  c("In", "Out")[sapply(c("In", "Out"), function(x)
          grepl(x, b))]
        
        code_col <-
          paste0("FlagCode", dir)[sapply(c("69"), function(x)
            grepl(x, c))]
        if (length(code_col) != 0) {
          recap_aux[[marker_cols[2]]][a] <-
            cap.data_aux$'R'[[code_col]][a]
        }
        
        color_col <-
          c(paste0("FlagColor", dir), paste0(b, "_colorband"))[sapply(c("69", "01A"), function(x)
            grepl(x, c))]
        if (length(color_col) != 0) {
          recap_aux[[marker_cols[3]]][a] <-
            cap.data_aux$'R'[[color_col]][a]
        }
        
        code_color_col <-
          paste0("FlagCode", dir, "_color")[sapply(c("69"), function(x)
            grepl(x, c))]
        if (length(code_color_col) != 0) {
          recap_aux[[marker_cols[4]]][a] <-
            cap.data_aux$'R'[[code_color_col]][a]
        }
        
        recap_aux[[marker_cols[5]]][a] <-
          c("L", "R")[sapply(c(".{1}L", ".{1}R"), function(x)
            grepl(x, b))]
        
        recap_aux[[marker_cols[6]]][a] <-
          paste0("L", c("B", "A")[sapply(c("^L", "^U"), function(x)
            grepl(x, b))])
        
        if (cap.data_aux$'R'[a, b] == "00") {
          recap_aux[["Banded Leg"]][a] <-
            c("L", "R")[sapply(c(".{1}L", ".{1}R"), function(x)
              grepl(x, b))]
        }
        
      }
    }
  }
  
  for (b in c("TagType1", "TagType2")) {
    if (!is.na(cap.data_aux$'R'[a, b]) &
        !TRUE %in% sapply(leg_cols, function(x)
          grepl(cap.data_aux$'R'[a, b], cap.data_aux$'R'[a, x]))) {
      status <-
        cap.data_aux$'R'[a, paste0("TagStatus", c("1", "2")[sapply(c("1", "2"), function(x)
          grepl(x, b))])]
      
      dir <-
        list("Out", "In",
          c("In", "Out"),
          c("In", "Out"))[c("Deployed",
                            "Removed",
                            "Replaced",
                            "Retained") %in% status]
      
      for (c in unlist(dir)) {
        direction  <-
          c("capture", "release")[sapply(c("In", "Out"), function(x)
            grepl(x, c))]
        
        if (grepl("In", c)) {
          in.num <- in.num + 1
          n <- in.num
        } else if (grepl("Out", c)) {
          out.num <- out.num + 1
          n <- out.num
        }
        
        recap_aux[[colnames(recap_aux)[grepl(n, colnames(recap_aux)) &
                                         grepl(direction, colnames(recap_aux))][1]]][a] <-
          unlist(cap.data_aux$'R'[a, b])
        
      }
    }
  }
  
  capture.end.num <-
    c(29, 35, 41, 47, 53, 59)[sapply(c(1, 2, 3, 4, 5, 6), function(x)
      grepl(x, in.num))]

    marker_string <-
      tidyr::replace_na(unlist(recap_aux[a, 24:capture.end.num]), "")
  
  for (b in
       names(marker_string)[grepl("Side", names(marker_string))]) {
    marker_string <-
      append(marker_string, c("", ""), after = which(names(marker_string) == b))
  }
  
    if (paste0(paste0(marker_string, collapse = "|"), "|") == "00||||R|||LB|") {
      recap_aux[a, colnames(recap_aux)[grepl("1", colnames(recap_aux)) &
                                         grepl("capture", colnames(recap_aux))]] <- NA
      
    } else {
      recap_aux[['All Marker Info At Capture']][a] <-
        paste0(paste0(marker_string, collapse = "|"), "|")
    }
  
  
  release.end.num <-
    c(65, 71, 77, 83, 89, 95)[sapply(c(1, 2, 3, 4, 5, 6), function(x)
      grepl(x, in.num))]
  
  marker_string <-
    tidyr::replace_na(unlist(recap_aux[a, 60:release.end.num]), "")
  
  for (b in
       names(marker_string)[grepl("Side", names(marker_string))]) {
    marker_string <-
      append(marker_string, c("", ""), after = which(names(marker_string) == b))
  }
  
  recap_aux[['All Marker Info At Release']][a] <-
    paste0(paste0(marker_string, collapse = "|"), "|")
  
}

```

#### Export files

```{r}

message(cat(
  paste0(
    "The following data files have been processed and are ready for Bander Portal.\n"
  ),
  paste0(names(data.list), sep = "\n"),
  paste0("\n\nProvide a prefix for the newly written files.\n")
))

prefix <-
  gsub(" ", "_", noquote(as.character(readline(prompt =))))

if (!TRUE %in% sapply(c("_", ".", "-"), function(x)
  endsWith(prefix, x))) {
 prefix<- paste0(prefix,"_")
}

for(a in c("band_no_aux", "band_aux", "recap_no_aux", "recap_aux")) {
  if (rowSums(is.na(get(a)))[1] != ncol(get(a))) {
    openxlsx::write.xlsx(get(a),
                         paste0(
                           processed.data.path,
                           prefix,
                           a,
                           "_",
                           format(Sys.time(), '%Y%m%d-%H%M%S'),
                           ".xlsx"
                         ))
    
  }
}

message(cat(
  paste0(
    "The files were sucessfully written to '" ,
    processed.data.path,
    "'"
  )
))

utils::browseURL(processed.data.path)

```

# Survey data transformations

### ebird

```{r}
# survey.to.ebird
library("auk")

# ebird
# usfwslandbirds
# fwslandbirds
# API key: e2m7tt6sfo4n


```

# Rename photos

### Provide inputs

```{r}
photo.file.path <- "C:/Users/hvincelette/OneDrive - DOI/Documents/Fieldwork_wrap/2024/Eielson/photos/ALMS_points/CFGesmundo/"

ref.file.path<- "data-processed/to_archive/"

date.col<- "Date"
date.format <- "%m/%d/%Y"
time.col<- "StartTime"
time.format<- "%H:%M"
obs.col<- "ObserverID"
ref.col<- "UnitID"
prefix<- "EIEL"
missing.photo<- c("221","181")

variation <- c("")

```

### Import data

```{r}
# Import ref data file(s)
ref.file <- importer(
  path = ref.file.path,
  multiple = FALSE,
  graphics = TRUE,
  column.format = list(
    c(column = date.col, format = date.format),
    c(column = time.col, format = time.format)
  )
)

```

### Filter data

```{r}

ref.file <- ref.file %>%
  dplyr::mutate({{date.col}} := lubridate::as_date(.data[[date.col]], format = date.format)) %>%
  dplyr::mutate({{time.col}} := lubridate::as_datetime(.data[[time.col]], format = time.format)) %>%
  dplyr::mutate("temp:Year" := format(.data[[date.col]], "%Y"))

year.options <-
  ref.file %>%
  dplyr::select("temp:Year") %>%
  dplyr::filter("temp:Year" != "") %>%
  na.omit() %>%
  purrr::pluck("temp:Year") %>%
  unique() %>%
  sort()

year.choice <- utils::select.list(
  year.options,
  graphics = TRUE,
  multiple = FALSE,
  title = cat(
    paste0("\nSelect a year.\nEnter 0 to proceed with all years.\n")
  )
)
 
if(year.choice !=""){
   ref.file <- ref.file %>% dplyr::filter(`temp:Year` == year.choice)
}

obs.choice <- utils::select.list(
  c(unique(na.omit(ref.file[[obs.col]]))),
  multiple = FALSE,
  graphics = TRUE,
  title = cat(
    paste0(
      "\nSelect an observer.\nNote, photos taken by each observer must be processed individually due to the possibility of concurrent timestamps.\n"
    )
  )
)

if (length(obs.choice) != 0) {
  ref.file <- ref.file %>% dplyr::filter(.[[obs.col]] %in% obs.choice)
}

```

### Formulate names

```{r}

ref.file <- ref.file %>%
  dplyr::select(c({{date.col}}, {{time.col}}, {{ref.col}})) %>%
  dplyr::distinct() %>%
  dplyr::arrange(.[[date.col]], .[[time.col]]) 

ref.order <- ref.file %>%
  purrr::pluck(ref.col) %>%
  sub(paste0(prefix, "_"), "", .) %>%
  subset(!(. %in% missing.photo)) %>%
  rep(., each=length(variation))

variation.vector <- rep(variation, length(fs::dir_ls(path = photo.file.path, type = "file")) / 4)

date.taken <- fs::dir_ls(path = photo.file.path, type = "file") %>%
  exifr::read_exif(tags =
                     c("DateTimeOriginal")) %>%
  dplyr::arrange(DateTimeOriginal) %>%
  dplyr::mutate_at("SourceFile", basename) %>%
  dplyr::mutate(Ext= tools::file_ext(SourceFile)) %>%
  dplyr::mutate_at(
    "DateTimeOriginal",
    ~ lubridate::as_date(., format = "%Y:%m:%d %H:%M:%S") %>% format(., "%Y%m%d")
  )


if (length(ref.order) != nrow(date.taken)) {
  temp <- ref.file %>%
    dplyr::group_by(.[[date.col]]) %>%
    dplyr::summarize(RefTotal = dplyr::n())
  
  message(
    "Images and reference points are not of equal length.\nCheck missing data are accurately reported before continuing.\n"
    
  )
  
  message(huxtable::print_screen(
    huxtable::hux(
      date.taken %>%
        dplyr::group_by(DateTimeOriginal) %>%
        dplyr::summarize(PhotoTotal = dplyr::n() / length(variation)) %>%
        dplyr::mutate(ReferenceTotal = temp$RefTotal)
    ),
    colnames = FALSE
  ))
  
}

if (any(is.na(date.taken$DateTimeOriginal)) == TRUE) {
  browseURL(photo.file.path)
  
  for (a in which(is.na(date.taken$DateTimeOriginal))) {
    pass <- 0
    while (pass == 0) {
      message(cat(
        paste0(
          "\nThe image ",
          date.taken$SourceFile[a],
          " is missing a capture date.\nProvide a date formatted as %Y%m%d (e.g., 20240630).\n"
        )
      ))
      
      new.date <- as.character(readline(prompt = ""))
      
      if (nchar(new.date) != 8 |
          new.date == "" |
          is.na(suppressWarnings(as.numeric(new.date)))) {
        message("Incorrect date format.\n")
      } else {
        date.taken$DateTimeOriginal[date.taken$SourceFile == date.taken$SourceFile[a]] <- new.date
        pass <- 1
      }
    }
  }
}

date.taken <- date.taken %>%
  dplyr::mutate(
    "FileName" = paste0(
      prefix,
      "_",
      ref.order,
      # "_",
      # variation.vector,
      "_",
      date.taken$"DateTimeOriginal",
      ".",
      date.taken$"Ext"
    )
  )

file.rename(
  paste0(photo.file.path, date.taken$SourceFile),
  paste0(photo.file.path, date.taken$FileName)
)

```

# Tracking data

### Movebank import

Verify the stored login credentials are correct, and add/remove credentials as needed.

```{r warning=FALSE, include=FALSE, results=FALSE}

# Check existing login credentials
keyring::key_list()

# Add login credentials
move2::movebank_store_credentials(username = "mbmlandbirds",
                                  password = "",
                                  key_name = "movebank")

# Remove login credentials
move2::movebank_remove_credentials(key_name = getOption("move2_movebank_key_name"))


```

*Optional: If you want to import Movebank data from a secondary account, paste the associated service name listed in the key list (e.g., "myOtherAccount") and run the following code. Note the account with the service name "movebank" is the default, so you only need to run this code if you want to access a different account. Restart R or execute* `options("move2_movebank_key_name" = "movebank")` *to* *return to the default account. See* `vignette("movebank", package="move2")` *for more details.*

```{r}

keyring::key_list()

options("move2_movebank_key_name" = "movebank")

```

Run the following code to import location data from one or more studies into R. Single study imports have the option of selecting a time range of location data, whereas multiple study imports include all location data. By default, the location and reference data are both imported

```{r warning=FALSE, include=FALSE, results=FALSE}
          
# Import Movebank location and/or reference data
df.loc <- get.move.data(
  location = TRUE,
  reference = FALSE,
  remove_movebank_outliers <- FALSE,
  omit_derived_data <- FALSE
)

Movebank_dictionary <- readxl::read_excel("reference/movebank/Migratory_Tracking_data_dictionary.xlsx", na = "")

load("reference/movebank/animal_id_key.RData")

aou_code <- read.csv("reference/IBP-AOS-LIST23.csv", na = "")

# Unlist dataframes (if necessary)
if(any(sapply(import, function(x)
  any(class(x) == "list")))==TRUE) {
  import.dfs <- rrapply::rrapply(
    import,
    classes = "data.frame",
    how = "flatten",
    options = list(namesep = "-",
                   simplify = FALSE)
  )
} else if (inherits(import, "data.frame") == TRUE) {
  import.dfs <- list(import)
  
} else{
  import.dfs <- import
}


```

### Transformations

```{r}

# Extract coordinates
for(x in 1:length(import.dfs)) {
  loc.cols <-
    colnames(dplyr::select(import.dfs[[x]], matches("location")))
  for (a in loc.cols) {
    if ("sfc_POINT" %in% class(import.dfs[[x]][[a]])) {
      names <- names(unlist(import.dfs[[x]][[a]][1]))
      names <- sapply(names, function(x)
        sub(pattern = "^argos_", replacement = "argos:", x), USE.NAMES = FALSE)
      suppressWarnings(
        import.dfs[[x]] <-  import.dfs[[x]] %>%
          tibble::add_column(
            !!names[1] := sf::st_coordinates(import.dfs[[x]][[a]])[, 1],
            !!names[2] := sf::st_coordinates(import.dfs[[x]][[a]])[, 2],
            .before = paste0(a)
          ) %>%
          dplyr::select(!a)
      )
      
    }
  }
}

# Update field names
import.dfs <- import.dfs %>%
  purrr::map(
    .f = ~ .x %>%
      setNames(., gsub(
        pattern = "_", replacement = "-", names(.)
      )) %>%
      setNames(., sub(
        pattern = "^argos-", replacement = "argos:", names(.)
      )) %>%
      data.table::setnames(
        .,
        old = c(
          "individual-id",
          "nick-name",
          "weight",
          "individual-comments",
          "individual-number-of-deployments"
        ),
        new = c(
          "animal-id",
          "nickname",
          "tag-mass",
          "animal-comments",
          "animal-number-of-deployments"
        )
        ,
        skip_absent = TRUE
      )
  )


movebank_attributes<- Movebank_dictionary$codeName[Movebank_dictionary$domainItem_value=="dataField"]

all_colnames <- import.dfs %>%
  purrr::map_df(~ names(.x) %>%
                  tibble::enframe(name = NULL)) %>% dplyr::pull(value)


new_colnames <-
  sapply(all_colnames[!all_colnames %in% movebank_attributes], function(x) {
    grep(paste0("-", x),
         movebank_attributes,
         value = TRUE)
  }) %>% purrr::compact(.)


if (length(new_colnames) != 0) {
  import.dfs <- import.dfs %>%
    purrr::map(.f = ~ data.table::setnames(
      .x,
      old = names(new_colnames),
      new = unlist(new_colnames),
      skip_absent = TRUE
    ))
}

  
# Add missing columns
req_cols <- c("animal-ring-id", "animal-marker-id")

import.dfs <- import.dfs %>%
  purrr::map(.f = ~ .x %>%
               dplyr::mutate(!!!setNames(
                 rep(NA, length(setdiff(
                   req_cols, colnames(.x)
                 ))), setdiff(req_cols, colnames(.x))
               )) %>% dplyr::relocate(
                 c("animal-ring-id", "animal-marker-id", "tag-local-identifier"),
                 .after = "animal-id"
               ))


import.dfs[grep("reference-data$", names(import.dfs))] <-
  import.dfs[grep("reference-data$", names(import.dfs))] %>%
  purrr::map(
    .f = ~ .x %>%
      dplyr::mutate(
        "deploy-date" = lubridate::as_date(.$"deploy-on-timestamp"),
        .after = "deploy-on-timestamp"
      ) %>%
      dplyr::mutate("deploy-year" = as.character(
        lubridate::year(.$"deploy-on-timestamp"),
        .after = "deploy-date"
      ))
  )


# Link banding data

## Import banding files
# rdr_folders <-
#   dir(path = "//ifw7ro-file.fws.doi.net/datamgt/mbm",
#       pattern = "mbmlb")
# 
# 
# banding_files<- lapply(rdr_folders, function(x) {
#     FWSAkRDRtools::read.tables(
#       pattern = "banding.*\\.csv$",
#       project = x,
#       subfolder.path = "data/final_data",
#       recursive = FALSE,
#       all = TRUE
#     )
# })
# 
# save(banding_files, file = "reference/movebank/banding_files.RData")

all_band_data <-
  banding_files %>%
  purrr::discard(is.null) %>%
  plyr::ldply(.) %>%
  dplyr::mutate_at(dplyr::vars("BandNumberOut", "Year"), as.character)


# Join missing band numbers
# na_animal_ring <- import.dfs$`reference-data` %>%
#   dplyr::filter(is.na(`animal-ring-id`)) %>%
#   dplyr::inner_join(
#     .,
#     aou_code[, c("SCINAME", "SPEC")],
#     by = c("individual-taxon-canonical-name" = "SCINAME"),
#     keep = FALSE,
#     relationship = "many-to-many"
#   ) %>%
#   dplyr::left_join(
#     .,
#     all_band_data[, c("Year", "BandNumberOut", "SpeciesCode", "FlagCodeOut")],
#     by = c(
#       "SPEC" = "SpeciesCode",
#       "deploy-year"  = "Year",
#       "individual-local-identifier" = "FlagCodeOut"
#     ),
#     keep = FALSE,
#     relationship = "many-to-many"
#   ) %>%
#   dplyr::mutate("animal-ring-id" = `BandNumberOut`,
#                 "animal-marker-id" = `individual-local-identifier`) %>%
#   dplyr::select(
#     "deploy-year",
#     "individual-taxon-canonical-name",
#     "animal-id",
#     "animal-ring-id",
#     "animal-marker-id"
#   ) %>%
#   dplyr::distinct() %>%
#   dplyr::mutate_at("animal-ring-id",
#                    ~ as.character(.)) %>%
#   dplyr::mutate_at(
#     "animal-ring-id",
#     ~ dplyr::case_when(
#       `animal-marker-id` == "CL/EX" ~ "123233886",
#       `animal-marker-id` == "04A" ~ "129210985",
#       `animal-marker-id` == "E4" ~ "137213263",
#       `animal-marker-id` == "E3" ~ "137213260",
#       TRUE ~ `animal-ring-id`
#     )
#   )
# 
# 
# animal_id_key <- import.dfs$`reference-data` %>%
#   dplyr::filter(!is.na(`animal-ring-id`)) %>%
#   dplyr::select(
#     "deploy-year",
#     "individual-taxon-canonical-name",
#     "animal-id",
#     "animal-ring-id",
#     "animal-marker-id"
#   ) %>%
#   dplyr::distinct() %>%
#   rbind(na_animal_ring) %>%
#   dplyr::mutate_at("animal-ring-id", ~ gsub(pattern = "-", replacement = "", .)) %>%
#   suppressWarnings(dplyr::mutate_at(
#     # if marker id occurs in numeric ring id
#     "animal-marker-id",
#     ~ dplyr::case_when(
#       is.na(as.numeric(`animal-ring-id`)) ~ `animal-ring-id`,
#       TRUE ~ `animal-marker-id`
#     )
#   )) %>%
#   suppressWarnings(dplyr::mutate_at("animal-ring-id", ~ ifelse(is.na(
#     as.numeric(`animal-ring-id`)
#   ), NA, `animal-ring-id`))) %>%
#   dplyr::mutate_all(as.character) %>%
#   dplyr::inner_join(
#     .,
#     aou_code[, c("SCINAME", "SPEC")],
#     by = c("individual-taxon-canonical-name" = "SCINAME"),
#     keep = FALSE,
#     relationship = "many-to-many"
#   ) %>%
#   dplyr::left_join(
#     .,
#     all_band_data[, c("Year", "BandNumberOut", "SpeciesCode", "FlagCodeOut")],
#     by = c(
#       "SPEC" = "SpeciesCode",
#       "deploy-year"  = "Year",
#       "animal-ring-id" = "BandNumberOut"
#     ),
#     keep = FALSE,
#     relationship = "many-to-many"
#   ) %>%
#   dplyr::mutate_at(
#     "animal-marker-id",
#     ~ dplyr::case_when(is.na(.) ~ `FlagCodeOut`, TRUE ~ `animal-marker-id`)
#   )
# 
# marker_id_check <- animal_id_key %>%
#   dplyr::filter(`animal-marker-id` != `FlagCodeOut`) %>%
#   dplyr::filter(!`animal-marker-id` %in% unlist(sapply(`FlagCodeOut`, function(x)
#     grep(x, `animal-marker-id`, value = TRUE)))) %>%
#   dplyr::mutate("animal-marker-id" = FlagCodeOut) %>%
#   dplyr::select(-"FlagCodeOut")
# 
# 
# animal_id_key <- animal_id_key %>%
#   dplyr::select(-"FlagCodeOut") %>%
#   dplyr::rows_update(
#     marker_id_check,
#     by = c(
#       "deploy-year",
#       "individual-taxon-canonical-name",
#       "animal-id",
#       "animal-ring-id"
#     )
#   )
# 
# save(animal_id_key, file = "reference/movebank/animal_id_key.RData")

# Check
dup_ring <- animal_id_key %>%
  dplyr::distinct(
    `deploy-year`,
    `animal-id`,
    `animal-ring-id`,
    `individual-taxon-canonical-name`,
    `animal-marker-id`
  ) %>%
  dplyr::filter(!is.na(`animal-ring-id`)) %>%
  subset(duplicated(`animal-ring-id`) |
           duplicated(`animal-ring-id`, fromLast = TRUE)) %>%
  data.table::setorder(`animal-ring-id`, na.last = TRUE)


animal_id_key <- animal_id_key %>%
  setNames(paste0("temp:", names(.)))
  

suppressWarnings(
  import.dfs<- 
    import.dfs %>%
    purrr::map(
      .f = ~ .x %>%
        dplyr::mutate("id" = seq_along(1:nrow(.x))) %>%
        dplyr::mutate_at(
          "animal-ring-id",
          ~ gsub(pattern = "-",
                 replacement = "", .) %>%
            as.numeric(.) # Only run this if all band numbers should be numeric
        ) %>%
        dplyr::mutate_at(dplyr::vars(
          c("animal-id",
            "animal-ring-id",
            "animal-marker-id")
        ), as.character) %>%
        dplyr::left_join(
          .,
          animal_id_key[, c(
            "temp:animal-id",
            "temp:animal-ring-id",
            "temp:animal-marker-id",
            "temp:deploy-year"
          )],
          by = c("animal-id" = "temp:animal-id"),
          keep = FALSE,
          relationship = "many-to-many" # different years
        ) %>%
        dplyr::mutate_at(
          "animal-ring-id",
          ~ dplyr::case_when(is.na(.) ~ `temp:animal-ring-id`, TRUE ~ `animal-ring-id`)
        ) %>%
        dplyr::mutate_at(
          "animal-marker-id",
          ~ dplyr::case_when(is.na(.) ~ `temp:animal-marker-id`, TRUE ~ `animal-marker-id`)
        ) %>%
        dplyr::select(-c(tidyselect::any_of(
          c(
            "deployment-local-identifier",
            "animal-nickname",
            "individual-local-identifier",
            "deploy-on-person",
            "deploy-off-person",
            grep(
              "-comments",
              colnames(.),
              value = TRUE,
              ignore.case = TRUE
            )
          )
        ), tidyselect::starts_with("temp:"))) %>% 
    dplyr::distinct()
    )
)

# Check
dup_id <-
  import.dfs %>%
  purrr::map(
    .f = ~ .x %>%
      dplyr::select(
        `id`,
        `animal-id`,
        `animal-ring-id`,
        `individual-taxon-canonical-name`,
        `animal-marker-id`
      ) %>%
      subset(
        duplicated(`id`) |
          duplicated(`id`, fromLast = TRUE)
      )
  )


## Prep band info
suppressWarnings(
  band_info <-
    all_band_data %>%
    dplyr::select(
      c(
        "SpeciesCode",
        "Year",
        "Date",
        "Latitude_dec",
        "Longitude_dec",
        "BandNumberOut",
        "Mass_g",
        "Age",
        "CHDSex",
        "FieldSex",
        "FlagCodeOut"
        # , grep(
        #   "tag",
        #   colnames(all_band_data),
        #   value = TRUE,
        #   ignore.case = TRUE
        # )
      )
    ) %>%
    dplyr::mutate_at("Date",~format(as.Date(., format = "%m/%d/%Y"), "%Y-%m-%d")) %>%
    dplyr::mutate_at("Mass_g",~units::set_units(., "g", mode = "standard")) %>%
    dplyr::mutate_at(c("FieldSex", "CHDSex"), ~tolower(.) %>%
                     dplyr::na_if(., "u")) %>%
    dplyr::mutate_at("Age", ~dplyr::recode(., HY = "juvenile",
                                          L = "juvenile",
                                          AHY = "adult",
                                          SY = "adult",
                                          ASY = "adult",
                                          TY = "adult",
                                          ATY = "adult") %>%
                                          dplyr::na_if(.,"U")
                                          ) %>%
    
    dplyr::mutate_at("Year",  ~ ifelse(
      is.na(.) & !is.na("Date"),
      format(as.Date(all_band_data$Date, format = "%d/%m/%Y"), "%Y"),
      .
    )) %>%
    dplyr::mutate_at(dplyr::vars(c("BandNumberOut","Year")), as.character) %>%
    dplyr::rowwise() %>%
    dplyr::mutate_at("FlagCodeOut", ~ ifelse(
      !is.na(as.numeric(.)),
      format(as.numeric(.),
             scientific = FALSE),
      .
    )) %>%
    setNames(paste0("temp:", names(.)))
)


## Join sex results (usually in data postprocessing)
band_info <- list(band_info) %>%
  purrr::map(
    .f = ~ .x %>%
      dplyr::mutate(!!!setNames(
        rep(NA, length(setdiff(
          "temp:CHDSex", colnames(.x)
        ))), setdiff("temp:CHDSex", colnames(.x))
      )) %>%
      dplyr::relocate("temp:CHDSex", .before = "temp:FieldSex") %>%
      dplyr::left_join(
        .,
        dplyr::distinct(sex_results[, c("BandNumber", "Sex")]),
        by = c("temp:BandNumberOut" = "BandNumber"),
        keep = FALSE,
        relationship = "many-to-many"
      ) %>%
      dplyr::mutate_at("temp:CHDSex", ~ dplyr::case_when(is.na(.) ~ Sex, TRUE ~ `temp:CHDSex`)) %>%
      dplyr::select(-"Sex")
  ) %>% `[[`(1)


## Mutate/fill missing values in reference dfs
req_cols <-
  c(
    "animal-life-stage",
    "animal-sex",
    "animal-mass",
    "tag-manufacturer-name",
    "tag-model",
    "capture-longitude",
    "capture-latitude"
  )

import.dfs[grep("reference-data$", names(import.dfs))] <-
  import.dfs[grep("reference-data$", names(import.dfs))] %>%
  purrr::map(
    .f = ~ .x %>%
      dplyr::mutate("id" = seq_along(1:nrow(.x))) %>%
      dplyr::relocate("id") %>%
      dplyr::mutate(!!!setNames(
        rep(NA, length(setdiff(
          req_cols, colnames(.x)
        ))), setdiff(req_cols, colnames(.x))
      )) %>%
      dplyr::left_join(
        .,
        dplyr::distinct(band_info[,!(names(band_info) == c("deploy-date"))]),
        by = c("animal-ring-id" = "temp:BandNumberOut",
               "deploy-year" = "temp:Year"),
        keep = FALSE,
        relationship = "many-to-many"
      ) %>%
      dplyr::mutate_at(
        "animal-marker-id",
        ~ dplyr::case_when(is.na(.) ~ `temp:FlagCodeOut`, TRUE ~ `animal-marker-id`)
      ) %>%
      dplyr::mutate_at(
        "animal-life-stage",
        ~ dplyr::case_when(is.na(.) ~ `temp:Age`, TRUE ~ `animal-life-stage`) %>%
          dplyr::recode(
            .,
            HY = "juvenile",
            L = "juvenile",
            AHY = "adult",
            SY = "adult",
            ASY = "adult",
            TY = "adult",
            ATY = "adult",
            "brood rearing" = "adult",
            "brood reaing" = "adult"
          ) %>% tolower(.)
        
      ) %>%
      dplyr::mutate_at(
        "animal-sex",
        ~ dplyr::case_when(
          is.na(.) ~ dplyr::case_when(is.na(`temp:CHDSex`) ~ `temp:FieldSex`,
                                      TRUE ~ `temp:CHDSex`),
          TRUE ~ `animal-sex`
        )
      ) %>%
      dplyr::mutate_at(
        "animal-mass",
        ~ dplyr::case_when(is.na(.) ~ `temp:Mass_g`, TRUE ~ `animal-mass`)
      ) %>%
      dplyr::mutate_at(
        "capture-longitude",
        ~ dplyr::case_when(is.na(.) ~ `temp:Longitude_dec`, TRUE ~ `capture-longitude`)
      ) %>%
      dplyr::mutate_at(
        "capture-latitude",
        ~ dplyr::case_when(is.na(.) ~ `temp:Latitude_dec`, TRUE ~ `capture-latitude`)
      ) %>%
      dplyr::mutate("id.2" = seq_along(1:nrow(.))) %>%
       magrittr::extract(-c(purrr::pluck(dplyr::filter(subset(.,duplicated(id)|duplicated(id, fromLast = TRUE)),`deploy-date`!= `temp:Date`),"id.2"))) %>%
      dplyr::select(-c(starts_with("temp:"), "id.2"))
      
    )
    
# Check
dup_ring_data <-
  import.dfs[grep("reference-data$", names(import.dfs))] %>%
  purrr::map(
    .f = ~ .x %>%
      dplyr::distinct(
        `deploy-year`,
        `animal-id`,
        `animal-ring-id`,
        `individual-taxon-canonical-name`,
        `animal-marker-id`
      ) %>%
      dplyr::filter(!is.na(`animal-ring-id`)) %>%
      subset(
        duplicated(`animal-ring-id`) |
          duplicated(`animal-ring-id`, fromLast = TRUE)
      ) %>%
      data.table::setorder(`animal-ring-id`, na.last = TRUE)
  )

dup_ring_data
dup_ring_key


import.dfs[grep("location-data$", names(import.dfs))] <-
  import.dfs[grep("location-data$", names(import.dfs))] %>%
  purrr::map(.f = ~ .x %>%
               dplyr::filter_at(dplyr::vars(c(
                 grep("lat", colnames(.), value = TRUE),
                 grep("long", colnames(.), value = TRUE)
               )),
               dplyr::any_vars(!is.na(.))))


# Check
invalid_timestamps <-
  import.dfs[grep("location-data$", names(import.dfs))] %>%
  purrr::map(
    .f = ~ .x %>%
      dplyr::mutate("year" = lubridate::year(.$"timestamp"),
                    .after = "timestamp") %>%
      dplyr::mutate("month" = lubridate::month(.$"timestamp"),
                    .after = "year") %>%
      # dplyr::mutate_at(c("year", "month"), as.numeric) %>%
      dplyr::filter((`year` > lubridate::year(Sys.Date())) |
                      (
                        `year` == lubridate::year(Sys.Date()) &
                          `month` > lubridate::month(Sys.Date())
                      )
      )
  )


# Check
lapply(import.dfs, function(x)
  colnames(x)[!colnames(x) %in% movebank_attributes])

# Save dataframes as separate objects 
export<- lapply(import.dfs, function(x)
  split(x, x$`individual-taxon-canonical-name`))

if(any(sapply(export, function(x)
  any(class(x) == "list")))==TRUE) {
  export.dfs <- rrapply::rrapply(
    export,
    classes = "data.frame",
    how = "flatten",
    options = list(namesep = "-",
                   simplify = FALSE)
  )}
# else if (inherits(import, "data.frame") == TRUE) {
#   import.dfs <- list(import)
#   
# } else{
#   import.dfs <- import
# }

export.dfs <-
  export.dfs %>%
  purrr::map(.f = ~ .x %>%
               dplyr::select_if( ~ !(all(is.na(
                 .
               )))))


ref_num <- sapply(1:length(export.dfs), function(x) {
  if (grepl("location-data", names(export.dfs[x]))) {
    # max(as.numeric(export.dfs[[x]]$`event-id`))
    format(max(export.dfs[[x]]$`timestamp`), '%Y%m%d-%H%M%S')
    
  } else {
    format(Sys.time(), '%Y%m%d-%H%M%S')
  }
})

names(export.dfs) <-
  as.data.frame(list(names(export.dfs)), col.names = "original_name") %>%
  tidyr::separate(1,
                  into = c("type", "sp"),
                  sep = "(?<=location-data-)|(?<=reference-data-)") %>%
  dplyr::mutate_at("type", ~ gsub("(^-)|(-$)", "", .)) %>%
  dplyr::mutate("new_name" = paste0(sp, "-", type, "-", ref_num)) %>%
  dplyr::mutate_at("new_name", ~ gsub(" ", "_", .)) %>%
  purrr::pluck("new_name")



dest.path = "data-processed/movebank/"

for (i in seq_along(export.dfs)) {
  write.csv(export.dfs[[i]],
            paste0(dest.path, names(export.dfs[i]), ".csv"),
            na = "",
            row.names = FALSE, quote = FALSE)
  # assign(names(export.dfs[i]),as.data.frame(export.dfs[[i]]))
}
  

# Get updated movebank vocabulary
move.vocab <-
  move2::movebank_get_vocabulary(xml = "http://vocab.nerc.ac.uk/collection/MVB/current/",
                                 return_type = "list",
                                 omit_deprecated = FALSE)

move.vocab <- rrapply::rrapply(
  move.vocab,
  how = "bind",
  options = list(
    simplify = FALSE,
    coldepth = 3,
    namecols = TRUE
  )
) %>%
  tidyr::pivot_wider(
    names_from = "L2",
    values_from = "1",
    values_fill = NA,
    values_fn = unique
  ) %>%
  tidyr::unnest(colnames(.)) %>%
  dplyr::rename("codeName" = "L1") %>%
  dplyr::mutate_at(c("codeName", "altLabel"), ~ stringr::str_replace_all(., c("argos " = "argos:", " " = "-"))) %>%
  dplyr::mutate(
    "units" =
      stringr::str_match(definition, "Units: \\s*(.*?)\\s*;") %>% .[, 2] %>%
      replace(.,  . %in% c(
        "none", "not defined", "unitless", "not specified"
      ), NA) %>%
      gsub(".*none \\((.*)\\).*", "\\1", .)
  ) %>% dplyr::select("codeName", "definition", "units")

write.csv(move.vocab, "move.vocab.csv", na = "", row.names = FALSE)


#!# remove trailing " " & ";"

# move.vocab <- move.vocab %>%
#   dplyr::mutate_at("definition", ~
#                      stringr::str_replace_all(
#                        .,
#                        c(
#                          "Units: \\s*(.*?)\\s*;" = "",
#                          "Values are chosen from a controlled list: \\s*(.*?)\\s*\\." = "",
#                          "Allowed values are \\s*(.*?)\\s*\\." = ""
#                        )
#                      ))


```
