---
title: "Landbird Data Processing Workflow"
name: Hannah
date: "`r format(Sys.time(), '%B %Y')`"
output:
  html_notebook:
    toc: yes
    number_sections: yes
  html_document:
    toc: yes
    df_print: paged
address: Alaska Regional Office 1011 E Tudor Rd, Anchorage, AK
email: hannah_vincelette@fws.gov
github: "https://github.com/hdvincelette/landbird-dp-workflow"
surname: Vincelette
position: Wildlife Biologist
editor_options:
  chunk_output_type: console
---

# Overview

The purpose of this document is to provide a reproducible workflow for processing data collected annually by the Landbird section of USFWS Alaska Migratory Bird Management. Data processing is just one component of data management. For complete guidance, view the [Alaska Region Interim Data Management User Guide.](https://ak-region-dst.gitbook.io/alaska-region-interim-data-management-user-guide/)

#### What is an R Notebook?

["An R Notebook is an R Markdown document with chunks that can be executed independently and interactively, with output visible immediately beneath the input." - R Markdown: The Definitive Guide](https://bookdown.org/yihui/rmarkdown/notebook.html)

#### How to use this guide

This document is meant to be interactive, with text prompts and "code chunks" which may require input before being executed. The following workflow was created for tabular data collected annually for projects archived on the Regional Data Repository. While effort is made to generalize code to all actively collected data, further edits may be required to process data in an standardized format. In this case, it is recommended users duplicate and rename the "landbird-dp-workflow.Rmd" before making changes.

# Workspace setup

### Load packages

Run the following installation code.

```{r include=FALSE, results=TRUE}

# Clear environment
rm(list=ls())

# Install CRAN packages
if (!require("pacman"))
  install.packages("pacman")
pacman::p_load("devtools", "move", "auk", "magrittr", "shiny","bslib")

# Load project
setwd(dirname(rstudioapi::documentPath()))
rstudioapi::filesPaneNavigate(getwd())
devtools::load_all()

theme.selection<- shinythemes::shinytheme("superhero")

# Install GitHub packages
# # Uninstall packages (if re-installing)
# remove.packages("FWSAkRDRtools")
# remove.packages("mdJSONdictio")
# .rs.restartR()

# Set GitHub download file method (optional)
# options(download.file.method = "wininet")
# 
# ## Install FWSAkRDRtools from GitHub
# devtools::install_github("hdvincelette/FWSAkRDRtools")
# 
# ## Install mdJSONdictio from GitHub
# devtools::install_github("hdvincelette/mdJSONdictio")

current.dir<- dirname(rstudioapi::documentPath())


```

If the installations fail, refer to [Install mdJSONdictio R package](https://hdvincelette.github.io/mdJSONdictio/articles/03_Setup_mdJSONdictio.html#install-mdjsondictio-r-package) for troubleshooting tips.

### Create subfolders

Run the following installation code to install necessary subfolders

```{r}

fs::dir_create(names(which(!fs::dir_exists(
  c(
    "downloads/dictionaries",
    "downloads/data",
    "downloads/templates",
    "data-raw",
    "data-processed/bander_portal",
    "data-processed/movebank",
    "data-processed/to_archive",
    "data-processed/data_packages/"
  )
))))

```

# Data preprocessing

### Download authoritative files

Open the [Migratory Bird Management Regional Data Repository](file://ifw7ro-file.fws.doi.net/datamgt/mbm/) and navigate to the project folder. If you are working from a remote location, you must be connected to the VPN (Virtual Private Network) to access this folder.

Paste the names of the project folder, data dictionary(ies), and data entry template(s).

```{r include=FALSE}

# Provide the relevent information
RDR.keyword <- "Eielson"

```

Now run the following code chunk.

```{r include=FALSE, results=FALSE}


if(dir.exists(paste0("//ifw7ro-file.fws.doi.net/datamgt/mbm/"))==FALSE) {
  stop("Check the VPN connection and try again.")
}

project.folder <- dir(
  path = paste0("//ifw7ro-file.fws.doi.net/datamgt/mbm/"),
  pattern = paste0("^mbmlb.*", RDR.keyword),
  full.names = FALSE,
  recursive = FALSE,
  ignore.case = TRUE
)

if(length(project.folder)>1) {
  folder.choice <- utils::menu(
    c(project.folder),
    graphics = TRUE,
    title = cat(
      paste0(
        "\nMore than one project folder was found.\nSelect one to access.\n"
      )
    )
  )
  project.folder<- project.folder[folder.choice]
}

 
select.download.app(
  from.dir = paste0(
    "//ifw7ro-file.fws.doi.net/datamgt/mbm/",
    project.folder,
    "/data"
  ),
  to.dir = paste0(
    current.dir,
    "/downloads/data"
  ),
  ext = ".*",
  multiple = TRUE,
  type = "data"
)

for (y in 1:nrow(selected.files)) {
  downloader::download(
    url =
      paste0(
        "File://",
        gsub("^[[:punct:][:space:]]+", "", selected.files$path[y]),
        "/",
        selected.files$filename[1]
      ),
    destfile = file.path(selected.dir, selected.files$filename[y]),
    method = "auto",
    mode = "wb",
    quiet = FALSE
  )
}

select.download.app(
  from.dir = paste0("//ifw7ro-file.fws.doi.net/datamgt/mbm/", project.folder,"/metadata"),
  to.dir = paste0(
    current.dir,
    "/downloads/dictionaries"
  ),
  ext = ".*",
  multiple = TRUE,
  type = "dictionary"
)

for (y in 1:nrow(selected.files)) {
  downloader::download(
    url =
      paste0(
        "File://",
        gsub("^[[:punct:][:space:]]+", "", selected.files$path[y]),
        "/",
        selected.files$filename[1]
      ),
    destfile = file.path(selected.dir, selected.files$filename[y]),
    method = "auto",
    mode = "wb",
    quiet = FALSE
  )
}



```

View and verify the downloaded data dictionary(ies) and data entry template(s) (the default location is the "downloads" sub-folder of this R project).

*Optional: You can skip the above steps and download everything manually from the [Migratory Bird Management Regional Data Repository](file://ifw7ro-file.fws.doi.net/datamgt/mbm/) project folder. Data dictionaries and data entry templates should be available in the "metadata" sub-folder. Save a copy of data dictionary(ies) and data entry template(s) in an accessible location (i.e. the "downloads" sub-folder of this R project)*

### Enter data

Copy and save each data entry template in a preferred location for data entry (this can be in the "data-raw" sub-folder of this R project). Use a unique and descriptive name - refer to [Best Practices in Naming Conventions](https://ak-region-dst.gitbook.io/alaska-region-interim-data-management-user-guide/alaska-data-management-101/file-organization-and-best-practices/best-practices-in-naming-conventions). It's okay to save this file in an Excel file format to preserve validation rules.

Enter scanned or hard copy data according to the respective data dictionary(ies). Review entered data to check for errors. The Excel Filter function (Home\>Editing\>Filter) is especially useful for viewing unique values and value ranges.

Save the entered data file(s) in the "data-raw" sub-folder of this R project. Convert the file(s) to comma-separated values (CSV) format, if necessary. Make sure the correct sheet is selected when converting if more than one sheet is used (i.e., validation rules).

### Meld data

```{r}

# import data files and dictionary
select.import.app(dir = paste0(dirname(rstudioapi::documentPath()),"/data-raw"), 
                  ext = c("txt", "csv", "xls*"),
                  multiple = TRUE,
                  type = "data")

raw.data.list <- importer(
  path = selected.files$path[1],
  files = selected.files$filename
)

ref.data.list<- raw.data.list

data.filenames<- tools::file_path_sans_ext(selected.files$filename)

select.import.app(dir = paste0(dirname(rstudioapi::documentPath()),
                               "/downloads/dictionaries"),
                  ext = c("json","csv", "xls*"), 
                  multiple = FALSE,
                  type = "dictionary")

ref.dictionary <- importer(path = selected.files$path[1],
                            files = selected.files$filename)[[1]]

dxnry.filename<- tools::file_path_sans_ext(selected.files$filename)


# merge files
if(length(raw.data.list)>1) {
  # make data frame selections
  merge.df.app(raw.data.list)
  
  # execute forced rbind
  if (is.null(selected.df) == FALSE) {
    raw.data.list <-  c(list(
      merged_data = data.table::rbindlist(raw.data.list[names(raw.data.list) %in% selected.df], 
                                          fill = TRUE, 
                                          idcol = TRUE
                                          )),
    raw.data.list[!names(raw.data.list) %in% selected.df])
  }
  
  raw.data.list <- raw.data.list %>%
    purrr::map(
      .f = ~ .x %>%
        dplyr::mutate_if(is.character, ~ dplyr::na_if(., "")) %>%
        dplyr::mutate_if(is.character, ~ dplyr::na_if(., "NULL")) %>%
        dplyr::mutate_if(is.character, ~ dplyr::na_if(., missing_values))
    )
}

# map colnames to dictionary
assign("input.dxnry", ref.dictionary)
if (class(ref.dictionary) == "list") {
  input.dxnry <- mdJSONdictio::extract.mdJSON(x = input.dxnry,
                                              record.type = "dictionaries",
                                              multiple = FALSE)
  
  record.num <-  which(sapply(ref.dictionary[["data"]], "[", "id") == input.dxnry[["data"]][[1]][["id"]])
  
  
  dxnry.colnames <- unlist(sapply(rjson::fromJSON(input.dxnry[["data"]][[1]][["attributes"]][["json"]])[["dataDictionary"]][["entity"]][[1]][["attribute"]], "[", "codeName"))
} else {
  dxnry.colnames <- unique(input.dxnry$codeName)
}
   
for (a in 1:length(raw.data.list)) {
  assign("input.data", raw.data.list[[a]])
  
  invalid.cols <- colnames(input.data)[!colnames(input.data) %in% dxnry.colnames]
  
  
  if (length(invalid.cols) != 0) {
    for (invalid.col in invalid.cols) {
      # allow repeated use of colnames
      # unused.cols <- as.vector(dxnry.colnames[!dxnry.colnames %in% colnames(input.data)])
        unused.cols <- as.vector(dxnry.colnames)
      
      correct.colname.app(input.data, invalid.col, unused.cols)
      
      
      if (decision == "remove") {
        input.data <- input.data %>% dplyr::select(., -invalid.col)
        
      } else if (decision == "replace") {
        input.data <- input.data %>% dplyr::rename(., !!col.choice := !!invalid.col)
        # allow repeated use of colnames
        # unused.cols <- unused.cols[!unused.cols %in% col.choice]
        
      } else if (decision == "rename") {
        input.data <- input.data %>% dplyr::rename(., !!col.choice := !!invalid.col)
      }
    }
  }
  

  raw.data.list[[a]] <- input.data
  
}

names(warning.list) <- names(raw.data.list)


# converge repeat colnames

# fix colnames
# raw.data.list<- lapply(raw.data.list, function(df) {
#   names(df) <- stringr::str_replace(names(df), "[1-9]$", "")
#   df
# })

# fix values
raw.data.list <- lapply(raw.data.list, function(df) {
  df %>%
    (function(x) {
      names(x) <- make.unique(names(x), sep = ".")
      x
    }) %>%
    dplyr::mutate(dplyr::across(
      .cols = dplyr::where(is.character),
      .fns = ~ dplyr::na_if(., "")
    )) %>%
    dplyr::select(dplyr::where(~ !all(is.na(.))))
})

# convert number columns to numeric type
raw.data.list <- lapply(raw.data.list, function(df) {
  numeric_cols <- sapply(names(df), function(col) {
    col_clean <- na.omit(as.character(df[[col]]))
    col_clean <- col_clean[nzchar(col_clean)]
    if (length(col_clean) == 0)
      return(FALSE)
    all(!is.na(suppressWarnings(as.numeric(col_clean))))
  
  })
  
  for (col in names(df)[numeric_cols]) {
    df[[col]] <- suppressWarnings(as.numeric(as.character(df[[col]])))
  }
  
  df
})


raw.data.list <- lapply(raw.data.list, function(df) {
  unique_names <- unique(sub("(?<!^)\\.[^.]+$", "", names(df), perl = TRUE))
  result <- data.frame(matrix(NA, nrow = nrow(df), ncol = 0))
  
  for (name in unique_names) {
    cols <- names(df)[which(sub("(?<!^)\\.[^.]+$", 
                                "", 
                                names(df), perl = TRUE) == name)]
    
    if (length(cols) > 1) {
      temp_df <- df %>% dplyr::select(all_of(cols))
      
      if (any(rowSums(!is.na(temp_df)) > 1)) {
        print(cols)
        
        result[cols] <- lapply(cols, function(c)
          df[[c]])
        
      } else {
        col_types <- sapply(temp_df, class)
        if (any(col_types %in% c("character"))) {
          temp_df <- dplyr::mutate_all(temp_df, as.character)
        }
        
        
        result[[name]] <- dplyr::coalesce(!!!temp_df)
      }
      
    } else {
      result[[name]] <- df[[name]]
    }
  }
  result
})


raw.data.list <- lapply(raw.data.list, function(df) {
  df %>%
    dplyr::select(dplyr::any_of(unused.cols), dplyr::everything())
})
  

melded_df<- raw.data.list[[1]]

write.table(
  melded_df,
  paste0("data-processed/", "RUBL_melded_v1", ".txt"),
  na = "",
  row.names = FALSE,
  quote = FALSE,
  sep = "\t"
)

  
```

# Data quality checks

Before beginning, make sure all entered data and data dictionary(ies) are saved in the sub-folders of this R project ("data-raw" and "downloads", respectively).

### Provide inputs

```{r}

# Provide the relevant information (if applicable)
missing_values<- c(NA)
species.col <- "SpeciesCode"
species.col <- "Species"
date.col <- "Date"
date.format <-  "%m/%d/%Y"
time.col <- "CaptureTime_2400"
time.col <- "Time"
time.format <- "%H:%M"
na.if<- c("","-","—",".","NA")

```

### Import

```{r}


select.import.app(dir = paste0(dirname(rstudioapi::documentPath()),"/data-raw"), 
                  ext = c("txt", "csv", "xls*"),
                  multiple = TRUE,
                  type = "data")

raw.data.list <- importer(
  path = selected.files$path[1],
  files = selected.files$filename,
  column.format = list(
    c(column = date.col, format = date.format),
    c(column = time.col, format = time.format)
  )
)

ref.data.list<- raw.data.list

data.filenames<- tools::file_path_sans_ext(selected.files$filename)

select.import.app(dir = paste0(dirname(rstudioapi::documentPath()),
                               "/downloads/dictionaries"),
                  ext = c("json","csv", "xls*"), 
                  multiple = FALSE,
                  type = "dictionary")

ref.dictionary <- importer(path = selected.files$path[1],
                            files = selected.files$filename)[[1]]

dxnry.filename<- tools::file_path_sans_ext(selected.files$filename)

```

### Filter data

```{r}

aou_code <- read.csv(list.files("reference/IBPAOU", full.names = TRUE)
                     , na = "")

# Remove NA rows
raw.data.list <- raw.data.list %>%
  purrr::map(.f = ~ .x %>%
               dplyr::mutate(dplyr::across(
                 where(is.character), ~ ifelse(. %in% na.if, NA, .)
               )) %>%
               dplyr::filter(!dplyr::if_all(colnames(.), is.na)))

## Translate species names/codes (if neccessary)
# 
# raw.data.list <- raw.data.list %>%
#   purrr::map(
#     .f = ~ .x %>%
#       dplyr::rename_with(., ~gsub('\\.', '', .x)) %>%
#       dplyr::inner_join(
#         .,
#         aou_code[, c("COMMONNAME", "SPEC")],
#         by = c("CommonName" = "COMMONNAME"),
#         keep = FALSE,
#         relationship = "many-to-many"
#       ) %>%
#       dplyr::mutate("SpeciesCode" = `SPEC`) %>%
#       dplyr::select(-"SPEC")
#   )
# 
# # Check species names/codes
# raw.data.list %>%
#   purrr::map(.f = ~ .x %>%
#                dplyr::filter(!.data[[species.col]] %in% aou_code$SPEC))


if(any(c(species.col, date.col) %in% unlist(purrr::map(raw.data.list, colnames)))) {
  # Make filter selections
  if (species.col %in% unlist(purrr::map(raw.data.list, colnames))) {
    species <- unique(purrr::map(raw.data.list, ~ .x %>%
                                   purrr::pluck(species.col)) %>%
                        unlist(.))
  } else {
    species<- character(0)
  }
  
  
  if (date.col %in% unlist(purrr::map(raw.data.list, colnames))) {
    date.ranges <- raw.data.list %>% purrr::map(
      .f = ~ .x %>%
        dplyr::mutate("{{date.col}}" := ifelse(!date.col %in% names(.), "", {{date.col}})) %>%
        dplyr::select(any_of(date.col)) %>%
        dplyr::filter(.data[[date.col]] != "") %>%
        dplyr::mutate_at(dplyr::vars(tidyselect::any_of(date.col)), as.Date, format = date.format) %>%
        # dplyr::mutate({{date.col}} := as.Date(.data[[date.col]], format = date.format)) %>%
        append(list(
          "Min" = min(.[[date.col]]), "Max" = max(.[[date.col]])
        ))
    )
    
  
    dates.df <- raw.data.list %>% purrr::map(
      .f = ~ .x %>%
        dplyr::mutate({{date.col}} := ifelse(!date.col %in% names(.), "", .data[[date.col]])) %>%
        dplyr::select(all_of(date.col)) %>%
        dplyr::filter(.data[[date.col]] != "") %>%
        dplyr::mutate({{date.col}} := as.Date(.data[[date.col]], format = date.format))
    ) %>%
      dplyr::bind_rows(., .id = "id")
    
    min.date <- min(do.call("c", lapply(date.ranges, "[[", "Min")))
    max.date <- max(do.call("c", lapply(date.ranges, "[[", "Max")))
    
  } else {
    min.date <- as.Date("0001-01-01")
    max.date <- as.Date("0001-01-01")
  }
  
  


  filter.df.app(raw.data.list, species.col, date.col, species, min.date, max.date)


# Apply filters
  id.var <- dplyr::last(make.names(c(unique(
    unlist(purrr::map(raw.data.list, colnames))
  ), ".id"), unique = TRUE))
  
  
  raw.data.list <- raw.data.list %>%
    purrr::map(.f = ~ .x %>%
                 tibble::rowid_to_column(var = id.var))
  
  other.data.list <- raw.data.list
  
  if (length(species) != 0) {
    raw.data.list <- raw.data.list %>%
      purrr::map(.f = ~ .x %>%
                   dplyr::filter(if_any(
                     matches(species.col), ~ .data[[species.col]] %in% selected.species
                   )))
  }
  
  
  if (min.date != "0001-01-01" |
      max.date != "0001-01-01") {
    raw.data.list <- raw.data.list %>%
      purrr::map(
        .f = ~ .x %>%
          dplyr::mutate_at(dplyr::vars(tidyselect::any_of(date.col)), as.Date, format = date.format) %>%
          dplyr::filter(if_any(
            matches(date.col),
            ~ data.table::between(.data[[date.col]], min(selected.dates), max(selected.dates), NAbounds =
                                    FALSE)
          ))
      )
  }
  
  rm.row <-  purrr::map(raw.data.list, id.var)
  
  other.data.list <- lapply(seq_along(rm.row), function(i) {
    if (rlang::is_empty(rm.row[[i]]) == FALSE) {
      other.data.list[[i]] %>%
        dplyr::slice(-rm.row[[i]])
    } else {
      other.data.list[[i]]
    }
  }) %>%
    purrr::set_names(names(raw.data.list))
  
  raw.data.list <-
    purrr::map(raw.data.list, ~ (.x %>%
                                   dplyr::select(-tidyselect::any_of(id.var))))
  
  other.data.list <- 
    purrr::map(other.data.list, ~ (.x %>% 
                                     dplyr::select(-tidyselect::any_of(id.var))))
   

# Remove empty dataframes
      raw.data.list <- raw.data.list %>%
        purrr::discard(., ~ nrow(.) == 0) %>%
        purrr::map(.f = ~ .x %>%
                           dplyr::mutate_at(dplyr::vars(tidyselect::any_of(date.col)), format, format = date.format))
                     
      
      other.data.list <- other.data.list %>%
        purrr::discard(., ~ nrow(.) == 0) %>%
         purrr::map(.f = ~ .x %>%
                           dplyr::mutate_at(dplyr::vars(tidyselect::any_of(date.col)), format, format = date.format))
  
}
```

### Merge data

```{r}

if(length(raw.data.list)>1) {
  # Make data frame selections
  merge.df.app(raw.data.list)
  
  # Execute forced rbind
  if (is.null(selected.df) == FALSE) {
    raw.data.list <-  c(list(
      merged_data = data.table::rbindlist(raw.data.list[names(raw.data.list) %in% selected.df], 
                                          fill = TRUE, 
                                          idcol = TRUE
                                          )),
    raw.data.list[!names(raw.data.list) %in% selected.df])
  }
  
  raw.data.list <- raw.data.list %>%
    purrr::map(
      .f = ~ .x %>%
        dplyr::mutate_if(is.character, ~ dplyr::na_if(., "")) %>%
        dplyr::mutate_if(is.character, ~ dplyr::na_if(., "NULL")) %>%
        dplyr::mutate_if(is.character, ~ dplyr::na_if(., missing_values))
    )
}

```

### Validate data

```{r}

# Validate data file(s) against the dictionary
warning.list <- list()
assign("input.dxnry", ref.dictionary)
if (class(ref.dictionary) == "list") {
  input.dxnry <- mdJSONdictio::extract.mdJSON(x = input.dxnry,
                                              record.type = "dictionaries",
                                              multiple = FALSE)
  
  record.num <-  which(sapply(ref.dictionary[["data"]], "[", "id") == input.dxnry[["data"]][[1]][["id"]])
  
  
  dxnry.colnames <- unlist(sapply(rjson::fromJSON(input.dxnry[["data"]][[1]][["attributes"]][["json"]])[["dataDictionary"]][["entity"]][[1]][["attribute"]], "[", "codeName"))
} else {
  dxnry.colnames <- unique(input.dxnry$codeName)
}
   
for (a in 1:length(raw.data.list)) {
  assign("input.data", raw.data.list[[a]])
  
  # Check colnames
  invalid.cols <- colnames(input.data)[!colnames(input.data) %in% dxnry.colnames]
  
  
  if (length(invalid.cols) != 0) {
    for (invalid.col in invalid.cols) {
      unused.cols <- as.vector(dxnry.colnames[!dxnry.colnames %in% colnames(input.data)])
      
      correct.colname.app(input.data, invalid.col, unused.cols)
      
      
      if (decision == "remove") {
        input.data <- input.data %>% dplyr::select(., -invalid.col)
        
      } else if (decision == "replace") {
        input.data <- input.data %>% dplyr::rename(., !!col.choice := !!invalid.col)
        unused.cols <- unused.cols[!unused.cols %in% col.choice]
        
      } else if (decision == "rename") {
        input.data <- input.data %>% dplyr::rename(., !!col.choice := !!invalid.col)
      }
    }
  }
  
  # Create warning table
  if (class(ref.dictionary) == "list") {
    warning.list[[a]] <-
      mdJSONdictio::validate.mdJSON(x = input.dxnry, y = input.data)
  } else {
    warning.list[[a]] <-
      mdJSONdictio::validate.table(x = input.dxnry, y = input.data)
  }
  
  raw.data.list[[a]] <- input.data
  
}

names(warning.list) <- names(raw.data.list)


```

### Correct data

```{r}

for (a in 1:length(raw.data.list)) {
  assign("input.data", raw.data.list[[a]])
  
  if (class(ref.dictionary) == "list") {
    ref.dxnry <- mdJSONdictio::build.table(input.dxnry)
  } else {
    assign("ref.dxnry", input.dxnry)
  }
  
  warn.table <- warning.list[[a]]
  
  # Correct entry errors
  domainItem_value.table <-
    warn.table %>% dplyr::filter(Category == "domainItem_value")
  
  if (nrow(domainItem_value.table) != 0) {
    for (b in 1:nrow(domainItem_value.table)) {
      invalid.values <-
        scan(
          text = noquote(gsub(
            ".*: ", "", domainItem_value.table$Message[b]
          )),
          what = "",
          quiet = TRUE,
          sep = ",",
          strip.white = TRUE
        )
      
      
      variable <- domainItem_value.table$Variable[b]
      
      choose.value.app(variable, invalid.values, ref.dxnry)
      
      #!!# temporary patch
      if ("" %in% selected.values) {
        selected.values <- ifelse(selected.values == "", "NA", selected.values)
      }
      
      for (invalid.value in selected.values) {
        correct.value.app(variable, invalid.value, input.data, ref.dxnry)
        
        if (decision %in% c("correct", "replace")) {
          class <- class(input.data[[variable]])
          
          input.data <- input.data %>%
            dplyr::mutate_at(
              variable,
              ~ as.character(.) %>% dplyr::recode(., !!invalid.value := value.choice)
            )
          
          # if (numeric == TRUE &
          #     is.numeric(input.data[[variable]]) == TRUE) {
          #   input.data <- input.data %>%
          #     dplyr::mutate_at(variable, ~ as.numeric(.))
          #   
          # }
          # 
          # if (is.na(match.fun(paste0("as.", class))(value.choice))
          #   == TRUE) {
          #     
          #   }
          
        }
        
        
      }
    }
  }
  
      
  #     test<- summary(arsenal::comparedf(
  #   dplyr::mutate(df, across(
  #     where(is.character), ~ dplyr::na_if(.x, "")
  #   )),
  #   dplyr::mutate(input.data, across(
  #     where(is.character), ~ dplyr::na_if(.x, "")
  #   ))
  # ))
      
  
  raw.data.list[[a]] <- input.data
}

raw.data.list <- purrr::map(raw.data.list, ~ (.x %>% dplyr::select(-tidyselect::any_of(id.var))))


```

### Correct dictionary

```{r}

req.attribute.fields <-
      c("codeName", "allowNull", "dataType", "definition", "domainId")

for (a in 1:length(warning.list)) {
  warn.table <- warning.list[[a]]
  
  
  # codeName
  codeName.warn <-
    warn.table %>% dplyr::filter(Category == "codeName")
  
  if (nrow(codeName.warn) != 0) {
 
    add.codeName.app(
      data = as.data.frame(raw.data.list[[a]]),
      filename = names(raw.data.list[a]),
      missing.variables = codeName.warn$Variable
    )
    
    if (length(codeName.choice)!=0) {
      for (b in 1:length(codeName.choice)) {
        if (class(ref.dictionary) == "list") {
          input.dxnry <-
            mdJSONdictio::modify.mdJSON(
              x = input.dxnry,
              how = "add_attribute",
              attribute_codeName = codeName.choice[b],
              attribute_units = NULL,
              attribute_unitsResolution = NULL,
              attribute_isCaseSensitive = NULL,
              attribute_missingValue = NULL,
              attribute_minValue = NULL,
              attribute_maxValue = NULL,
              attribute_fieldWidth = NULL,
              quiet = FALSE
            )
        } else {
          input.dxnry <-
            mdJSONdictio::modify.table(
              x = input.dxnry,
              how = "add_attribute",
              attribute_codeName = codeName.choice[b],
              attribute_units = NULL,
              attribute_unitsResolution = NULL,
              attribute_isCaseSensitive = NULL,
              attribute_missingValue = NULL,
              attribute_minValue = NULL,
              attribute_maxValue = NULL,
              attribute_fieldWidth = NULL,
              quiet = FALSE
            )
        }
      }
    }
  }
  

  # domainItem_value
  domainItem_value.table <-
    warn.table %>% dplyr::filter(Category == "domainItem_value")
  
  if (class(ref.dictionary) == "list") {
    ref.dxnry <- mdJSONdictio::build.table(input.dxnry)
  } else {
    assign("ref.dxnry", input.dxnry)
  }
  
  if (nrow(domainItem_value.table) != 0) {
    for (b in 1:nrow(domainItem_value.table)) {
      domainItem.vector <-
        scan(
          text = noquote(gsub(
            ".*: ", "", domainItem_value.table$Message[b]
          )),
          what = "",
          quiet = TRUE,
          sep = ",",
          strip.white = TRUE
        )
      
      add.domainItem.app(
        domainItem_value.table$Variable[b],
        domainItem.vector,
        ref.dxnry)
      
      if (length(domainItem.choice)!=0) {
        for (c in 1:length(domainItem.choice)) {
          
          if (class(ref.dictionary) == "list") {
            input.dxnry <-
              mdJSONdictio::modify.mdJSON(
                x = input.dxnry,
                how = "add_domainItem",
                domain_codeName = domainItem_value.table$Variable[b],
                domainItem_name = domainItem.choice[c],
                domainItem_value = domainItem.choice[c],
                quiet = FALSE
              )
          } else {
            input.dxnry <-
              mdJSONdictio::modify.table(
                x = input.dxnry,
                how = "add_domainItem",
                domain_codeName = domainItem_value.table$Variable[b],
                domainItem_name = domainItem.choice[c],
                domainItem_value = domainItem.choice[c],
                quiet = FALSE
              )
          }
        }
        
      }
    }
  }
  
  # dataType
  
  dataType.table <-
    warn.table %>% dplyr::filter(Category == "dataType_RDatatype")
  
   dataType.vector <- get0("dataType.vector", envir = asNamespace("mdJSONdictio"))
   
   datatype.rules <- get0("datatype.rules", envir = asNamespace("mdJSONdictio"))
  
  if (nrow(dataType.table) != 0) {
    for (b in 1:nrow(dataType.table)) {
      detected.dataType <-
        gsub("[\\(\\)]",
             "",
             regmatches(
               dataType.table$Message[b],
               gregexpr("\\(.*?\\)", dataType.table$Message[b])
             )[[1]])[1]
      dxnry.dataType <-
        gsub("[\\(\\)]",
             "",
             regmatches(
               dataType.table$Message[b],
               gregexpr("\\(.*?\\)", dataType.table$Message[b])
             )[[1]])[2]
      
      correct.dataType.app(
        variable = dataType.table$Variable[b],
        dxnry.dataType = dxnry.dataType,
        detected.dataType = detected.dataType,
        data = raw.data.list[[a]],
        filename = names(raw.data.list[a])
      )
      
       if ("update" %in% dataType.choice) {
        input.dxnry <-
          mdJSONdictio::modify.mdJSON(
            x = input.dxnry,
            how = "update_attribute",
            attribute_codeName = dataType.table$Variable[b],
            attribute_allowNull = NULL,
            attribute_dataType = dataType.choice,
            attribute_definition = NULL,
            attribute_units = NULL,
            attribute_unitsResolution = NULL,
            attribute_isCaseSensitive = NULL,
            attribute_missingValue = NULL,
            attribute_minValue = NULL,
            attribute_maxValue = NULL,
            attribute_fieldWidth = NULL,
            domain_codeName = NULL,
            new.attribute_codeName = NULL,
            quiet = FALSE
          )
        
      }
    }
  }

}

ref.dictionary[["data"]][[record.num]] <- input.dxnry[["data"]][[1]]

  
```

### Final checks

```{r}
# continuous: box plots
# discrete: range of values (interactive table)
# plot lat/long
# Check flag code, flag color, and leg attachments for recaps

# Check/correct time cols

# All data
for (x in 1:length(raw.data.list)) {
  assign("input.data", raw.data.list[[x]])
  
  ## Duplicate rows

  
  # All capture data
  
  ## Unique values
  ### BBLDisposition = 1 & BandNumberOut
  ### Flag code & BandNumberOut
  ### Colorband combo & BandNumberOut

  ## Species v AOU codes
  
  ## Status codes
 
  ## "Out" leg_cols contains an X
  leg_cols <-   c("ULIn",
                  "LLIn",
                  "URIn",
                  "LRIn",
                  "ULOut",
                  "LLOut",
                  "UROut",
                  "LROut")
  
  leg_cols <- leg_cols[leg_cols %in% colnames(input.data)]
  
  which(rowSums(is.na(input.data[, leg_cols[grepl("Out", leg_cols)]])) == 4)
  
  # Recapture data
  ## "In" leg_cols contains an X
  which(rowSums(is.na(input.data[, leg_cols[grepl("In", leg_cols)]])) == 4 &
          input.data$BBLDisposition == "R")
  
  ## BandNumberIn == BandNumberOut
  which(is.na(input.data$BandNumberIn) &
          input.data$BBLDisposition == "R")
  
  raw.data.list[[x]] <- input.data
  
}

```

# Data postprocessing

### Add uuids

```{r}
raw.data.list <- raw.data.list %>%   purrr::map(
  .f = ~ .x %>%       dplyr::mutate(!!!setNames(
    rep(NA, length(setdiff(
      "OccurrenceID", colnames(.x)
    ))), setdiff("OccurrenceID", colnames(.x))
  )) %>% dplyr::relocate("OccurrenceID") %>%       dplyr::rowwise() %>%       dplyr::mutate_at(
    "OccurrenceID",
    ~ ifelse(is.na(.) == TRUE ||
               . == "", uuid::UUIDgenerate(use.time = FALSE), .)
  )
)    
```

### Join ancillary data

#### Location

```{r}
# Provide directory paths for the location data file(s)
ref.data.path <- "reference/ALMS/"

# Provide the relevent information
ref.point.prefix <- "EIEL_"
data.point.col <- "UnitID"
data.lat.col <- "Latitude_dec"
data.long.col <- "Longitude_dec"

```

```{r}

# Import reference data file(s)
ref.data.files <- importer(
  path = ref.data.path,
  multiple = TRUE,
  graphics = TRUE,
  pattern = c("*.gpx$", "*.kml$")
)

locations<- data.table::rbindlist(ref.data.files, fill = TRUE)

# Join location(s)
raw.data.list <- raw.data.list %>%
  purrr::map(
    .f = ~ .x %>%
      dplyr::left_join(
        .,
        locations,
        by = c(setNames("temp:Name", data.point.col)), 
               # "Year" = "temp:year"),
        keep = FALSE
      ) %>%
      dplyr::mutate_at(
        data.lat.col,
        ~ dplyr::case_when(is.na(.) ~ `temp:Latitude`, TRUE ~ !!as.name(data.lat.col))
      ) %>%
      dplyr::mutate_at(
        data.long.col,
        ~ dplyr::case_when(is.na(.) ~ `temp:Longitude`, TRUE ~ !!as.name(data.long.col))
      ) %>%
      dplyr::select(-tidyselect::starts_with("temp:"))
  )
    
```

#### Sex results

```{r}

selected.files<- fs::dir_ls(
  path =  "reference/analysis_results"
)

if(length(selected.files)==1) {
  sex_results <- importer(path = dirname(selected.files),
                        files = basename(selected.files))[[1]]
  
  
} else {
  select.import.app(
    dir =  paste0(current.dir, "reference/analysis_results"),
    ext = c("txt", "csv", "xls*"),
    multiple = FALSE,
    type = "data"
  )
  
  sex_results <- importer(path = selected.files$path[1], files = selected.files$filename)[[1]]
}

# filter for results
sex_results <- sex_results %>%
  dplyr::mutate_at("BandNumberOut", as.character) %>%
  # dplyr::filter(Sex %in% c("F", "M")) %>%
  dplyr::mutate_at("GenSex", ~ dplyr::recode(GenSex, "m" = "M", "f" = "F")) 


# Check for duplicate results (creates duplicate rows in left_join)
sex_results[c(
  anyDuplicated(sex_results$BandNumberOut, fromLast = TRUE),
  anyDuplicated(sex_results$BandNumberOut, fromLast = FALSE)
), ]


# Join sex results
raw.data.list <- raw.data.list %>%
  purrr::map(
    .f = ~ .x %>%
      dplyr::mutate_at("BandNumberOut", as.character) %>%
      dplyr::mutate(!!!setNames(
        rep(NA, length(setdiff(
          "CHDSex", colnames(.x)
        ))), setdiff("CHDSex", colnames(.x))
      )) %>%
      dplyr::relocate("CHDSex", .before = "FieldSex") %>%
      dplyr::mutate_at("CHDSex", ~  as.character(.) %>%
                         dplyr::na_if(., "")) %>% 
      dplyr::left_join(.,dplyr::distinct(sex_results[,c("BandNumberOut","GenSex")]), by = c("BandNumberOut" = "BandNumberOut"), keep = FALSE, relationship = "many-to-many") %>%
      dplyr::mutate_at("CHDSex", ~ dplyr::case_when(is.na(.) ~ GenSex, TRUE ~ CHDSex))
    # %>%
    #   dplyr::select(-"GenSex")
  )


# lapply(raw.data.list, function(df) df$CHDSex)
# lapply(temp, function(df) df$CHDSex)


```

### Fill missing info

#### Bird status

```{r}

ref_data <-
  lapply(list.files(path = "reference/bander_portal/", full.names = TRUE),
         readxl::read_excel)

names(ref_data) <-
  gsub(
    list.files(path = "reference/bander_portal/"),
    pattern = ".xlsx$",
    replacement = ""
  )

list2env(ref_data, globalenv())

for (x in 1:length(raw.data.list)) {
  assign("input.data", raw.data.list[[x]])
  
  # Add missing columns
  miss_cols <- setdiff(
    c(
      na.omit(capture_cols$my_col[!grepl("color", capture_cols$my_col)]),
      "TagType1",
      "TagType2",
      "TagStatus1",
      "TagStatus2"
    ),
    colnames(input.data)
  )
  
  input.data[miss_cols] <-
    as.character(NA)
  
  
  my_colors <- c(
    "Bk",
    "Br",
    "Db",
    "Lb",
    "Mb",
    "Dg",
    "Lg",
    "Gy",
    "Hp",
    "Mg",
    "Mv",
    "Or",
    "Pp",
    "Rd",
    "Wh",
    "Yl"
  )
  
  leg_cols <-   c("ULIn",
                  "LLIn",
                  "URIn",
                  "LRIn",
                  "ULOut",
                  "LLOut",
                  "UROut",
                  "LROut")
  
  leg_cols <- leg_cols[leg_cols %in% colnames(input.data)]
  
  status_info <- input.data %>%
    dplyr::rowwise() %>%
    dplyr::mutate_at("TagType1", ~ ifelse(TagStatus1  %in% c("Removed", "Lost"), TagType1 <-
                                            NA
                                          , .)) %>%
    dplyr::mutate_at("TagType2", ~ ifelse(TagStatus2  %in% c("Removed", "Lost"), TagType2 <-
                                            NA
                                          , .)) %>%
    dplyr::select(tidyselect::any_of(
      c(
        "OccurrenceID",
        "BandNumberOut",
        "SpeciesCode",
        "BBLStatus",
        leg_cols[grepl("Out", leg_cols)],
        "TagType1",
        "TagType2",
        "Blood",
        "BBLStatus"
      )
    )) %>%
    dplyr::mutate_at(dplyr::vars(c(
      leg_cols[grepl("Out", leg_cols)], "TagType1", "TagType2", "BBLStatus"
    )), ~ as.character(.) %>% dplyr::na_if(., "")) %>%
    dplyr::mutate_at(
      leg_cols[grepl("Out", leg_cols)],
      ~ stringi::stri_replace_all_fixed(., my_colors, "01A", vectorize_all = FALSE)
    ) %>%
    dplyr::mutate(NumMarkers = sum(!unique(dplyr::c_across(
      c(leg_cols[grepl("Out", leg_cols)], "TagType1", "TagType2")
    ))  %in% c("X", NA))) %>%
    dplyr::mutate_at(
      "BBLStatus",
      ~ dplyr::case_when(
        NumMarkers >= 2 &
          Blood != "Y" ~ "325",
        NumMarkers >= 1 &
          Blood == "Y" ~ "319",
        NumMarkers == 0 &
          Blood == "Y" ~ "318",
        NumMarkers == 0 &
          Blood != "Y" ~ "300",
        NumMarkers == 1 &
          dplyr::if_any(c(leg_cols[grepl("Out", leg_cols)], "TagType1", "TagType2"), ~
                          . == "69") &
          Blood != "Y" ~ "369",
        NumMarkers == 1 &
          dplyr::if_any(c(leg_cols[grepl("Out", leg_cols)], "TagType1", "TagType2"), ~
                          grepl("^81", .)) &
          Blood != "Y" ~ "381",
        NumMarkers == 1 &
          dplyr::if_any(c(leg_cols[grepl("Out", leg_cols)], "TagType1", "TagType2"), ~
                          grepl("^80", .)) &
          Blood != "Y" ~ "380",
        NumMarkers == 1 &
          dplyr::if_any(c(leg_cols[grepl("Out", leg_cols)], "TagType1", "TagType2"), ~
                          grepl("^90", .)) &
          Blood != "Y" ~ "390",
        NumMarkers == 1 &
          dplyr::if_any(c(leg_cols[grepl("Out", leg_cols)], "TagType1", "TagType2"), ~
                          grepl("^01", .)) &
          Blood != "Y" ~ "301",
        TRUE ~ BBLStatus
      )
    ) %>%
    dplyr::rename(`temp:BBLStatus` = BBLStatus)
  
  input.data <- input.data %>%
    dplyr::mutate(across(where(is.character), ~ dplyr::na_if(.x, ""))) %>%
    dplyr::left_join(
      .,
      status_info[, c("OccurrenceID", "temp:BBLStatus")],
      by = c("OccurrenceID" = "OccurrenceID"),
      keep = FALSE,
      relationship = "one-to-one"
    ) %>%
    dplyr::mutate_at("BBLStatus", ~ as.character(.)) %>%
    dplyr::mutate_at("BBLStatus",
                     ~ dplyr::case_when(is.na(.) ~ `temp:BBLStatus`, TRUE ~ BBLStatus))
  
  invalid.index <- which(input.data$BBLStatus != input.data$`temp:BBLStatus`)
  
  
  if (length(invalid.index) != 0) {
    for (a in invalid.index) {
      message(cat(
        paste0(
          "The status for the following individual appears invalid.\nSpecies: ",
          input.data$SpeciesCode[a],
          "\nBand number: ",
          input.data$BandNumberOut[a],
          "\nStatus (reported): ",
          input.data$BBLStatus[a],
          " [",
          bander_portal_lookups$description[grepl(stringr::str_sub(input.data$BBLStatus[a], 1, 1),
                                                  bander_portal_lookups$code) == TRUE &
                                              bander_portal_lookups$field == "Bird Status"],
          " ",
          bander_portal_lookups$description[grepl(stringr::str_sub(input.data$BBLStatus[a], 2, 3),
                                                  bander_portal_lookups$code) == TRUE &
                                              bander_portal_lookups$field == "Bird Status Extra Info"],
          "]",
          "\nStatus (detected): ",
          input.data$`temp:BBLStatus`[a],
          " [",
          bander_portal_lookups$description[grepl(
            stringr::str_sub(input.data$`temp:BBLStatus`[a], 1, 1),
            bander_portal_lookups$code
          ) == TRUE &
            bander_portal_lookups$field == "Bird Status"],
          " ",
          bander_portal_lookups$description[grepl(
            stringr::str_sub(input.data$`temp:BBLStatus`[a], 2, 3),
            bander_portal_lookups$code
          ) == TRUE &
            bander_portal_lookups$field == "Bird Status Extra Info"],
          "]",
          "\nBlood: ",
          input.data$Blood[a],
          "\nULOut: ",
          input.data$ULOut[a],
          "\nLLOut: ",
          input.data$LLOut[a],
          "\nUROut: ",
          input.data$UROut[a],
          "\nLROut: ",
          input.data$LROut[a],
          "\nTagStatus1: ",
          input.data$TagStatus1[a],
          "\nTagStatus2: ",
          input.data$TagStatus2[a]
        )
      ))
      
      status.choice <- utils::menu(
        choice = c("update as detected", "manually edit", "leave as is"),
        graphics = TRUE,
        title = cat("\nHow would you like to proceed?")
      )
      
      if (status.choice == 1) {
        input.data$BBLStatus[a] <- input.data$`temp:BBLStatus`[a]
        
        
      }
      if (status.choice == 2) {
        message(
          "\nIMPROTANT: you must click 'syncronise' (arrow loop icon) and then 'Done' to enact changes."
        )
        
        invalid.entries <- input.data %>%
          dplyr::ungroup() %>%
          dplyr::filter(dplyr::row_number() == a)
        
        invalid.entries.edit <-
          DataEditR::data_edit(
            invalid.entries,
            col_readonly = colnames(invalid.entries)[!colnames(invalid.entries) %in% "BBLStatus"],
            viewer = "pane"
          ) %>%
          dplyr::mutate_at("BBLStatus", as.character)
        
        input.data[match(invalid.entries.edit$OccurrenceID,
                         input.data$OccurrenceID), "BBLStatus"] <-
          invalid.entries.edit$BBLStatus
      }
      
    }
  }
  
  
  input.data <-
    input.data %>%
    dplyr::select(-tidyselect::starts_with("temp:"))
  
  raw.data.list[[x]] <- input.data
  
}
  
```

#### Other variables

```{r}

for(a in names(raw.data.list)){
  raw.data.list[[a]]<- raw.data.list[[a]] %>%
  DataEditR::data_edit()
}


raw.data.list %>% purrr::map(
  .f = ~ .x %>%
    dplyr::select(tidyselect::where(~ any(is.na(
      .
    ))))
)

```

### Export files

#### Provide inputs

```{r}

# Do you want to merge the data with an authoritative file?
auth.merge <- FALSE


```

#### Data

```{r}

# Rejoin filtered data
raw.data.list <-
  split(c(raw.data.list, other.data.list), c(names(raw.data.list), names(other.data.list))) %>%
  purrr::map(dplyr::bind_rows)


if (auth.merge == TRUE) {
  # Select files to merge
  if (length(raw.data.list) > 1) {
    # Make data frame selections
    merge.df.app(raw.data.list)
    
    # Execute forced rbind
    if (is.null(selected.df) == FALSE) {
      add.data <-  data.table::rbindlist(raw.data.list[names(raw.data.list) %in% selected.df], fill = TRUE) %>%
        as.data.frame(.) %>%
        dplyr::mutate_if(is.character, ~ dplyr::na_if(., "")) %>%
        dplyr::mutate_if(is.character, ~ dplyr::na_if(., "NULL")) %>%
        dplyr::mutate_if(is.character, ~ dplyr::na_if(., missing_values))
      
    }
  } else {
    add.data <-  data.table::rbindlist(raw.data.list, fill = TRUE)
  }
  
  
  # Import authoritative data file
  select.import.app(
    dir = paste0(current.dir, "/downloads/data"),
    ext = c("txt", "csv", "xls*"),
    multiple = FALSE
  )
  
  auth.data <- importer(path = selected.files$path[1], files = selected.files$filename)[[1]]
  
  auth.filename <- tools::file_path_sans_ext(selected.files$filename)
  
  # Add missing columns
  missing.cols <- add.data %>%
    dplyr::mutate_if(is.character, ~ dplyr::na_if(., c(""))) %>%
    purrr::discard(., ~ all(is.na(.))) %>%
    dplyr::select(colnames(.)[!colnames(.) %in% colnames(auth.data)]) %>%
    colnames()
  
  col.choice<- c()
  
  if (length(missing.cols) != 0) {
    add.column.app(missing.cols, add.data)
  }
  
  
  if (length(col.choice) != 0) {
    for (a in col.choice) {
      existing.cols <- colnames(auth.data)
      place.column.app(a, existing.cols)
      
      if (last == TRUE) {
        auth.data <- auth.data %>% tibble::add_column(!!a := NA, .before = NULL)
      } else if (loc.choice != "") {
        auth.data <- auth.data %>% tibble::add_column(!!a := NA, .before = loc.choice)
      }
      
    }
  }
  
  # Merge data
  new.data <- auth.data %>%
    merge(add.data, all = TRUE, sort = FALSE) %>%
    dplyr::select(colnames(auth.data))
  
  
  # Evaluate changes
  old <- auth.data %>%
    dplyr::mutate(across(where(is.character), ~ dplyr::na_if(.x, "")))
  
  new <- new.data %>%
    dplyr::mutate(across(where(is.character), ~ dplyr::na_if(.x, "")))
  
  # # alt stats
  # compare.stats<- dataCompareR::locateMismatches(old, new)
  # compare.stats<- dataCompareR::rCompare(old, new)
  # compare.stats<- summary(compare.stats)
  
  
  compare.stats <- arsenal::comparedf(new, old)
  compare.stats <- summary(compare.stats)
  
  comparison <- compareDF::compare_df(new, old, c("OccurrenceID"))
  
  compareDF::create_output_table(
    comparison,
    file_name = "data-processed/comparison/temp.xlsx",
    output_type = "xlsx",
    limit = max(nrow(new), nrow(old))
  )
  
  compare.df <- openxlsx::read.xlsx("data-processed/comparison/temp.xlsx")
  
  compare.dfs.app(new, old, compare.df, compare.stats)
  
  
  # Save files - versioning
  save.file.app(
    dir = paste0(
      current.dir,
      "/data-processed/to_archive"
    ),
    proposed.filename = paste0(auth.filename, " (1)"),
    ref.filename = auth.filename
  )
  
  write.csv(
    new.data,
    paste0(
      selected.dir,
      "\\",
      tools::file_path_sans_ext(new.name),
      ".csv"
    ),
    na = "",
    row.names = FALSE
  )
  
} else if (auth.merge == FALSE) {
  
  for (a in 1:length(raw.data.list)) {
    assign("old.data", ref.data.list[[a]])
    assign("new.data", raw.data.list[[a]])
    
    # Evaluate changes
    missing_cols <-   setdiff(names(new.data), names(old.data)) 
    #!# added because of an error; is this necessary?
    
    old <- old.data %>%
      dplyr::mutate(across(where(is.character), ~ dplyr::na_if(.x, ""))) %>%
      dplyr::bind_cols(setNames(replicate(length(missing_cols), NA, simplify = FALSE), missing_cols))
    
  
  new <- new.data %>%
    dplyr::mutate(across(where(is.character), ~ dplyr::na_if(.x, "")))
  
  # # alt stats
  # compare.stats<- dataCompareR::locateMismatches(old, new)
  # compare.stats<- dataCompareR::rCompare(old, new)
  # compare.stats<- summary(compare.stats)
  
  compare.stats <- arsenal::comparedf(new, old)
  compare.stats <- summary(compare.stats)
  
  suppressWarnings(comparison <- compareDF::compare_df(new, old, stop_on_error = FALSE))
  
  compareDF::create_output_table(
    comparison,
    file_name = "data-processed/comparison/temp.xlsx",
    output_type = "xlsx",
    limit = max(nrow(new), nrow(old))
  )
  
  compare.df <- openxlsx::read.xlsx("data-processed/comparison/temp.xlsx")
  
  compare.dfs.app(new, old, compare.df, compare.stats)
  
  
  # Save files - versioning
  save.file.app(
    dir = paste0(
      current.dir,
      "/data-processed/to_archive"
    ),
    proposed.filename = paste0(data.filenames[a], " (1)"),
    ref.filename = data.filenames[a]
  )
  
  selected.dir <- paste0(current.dir, "/data-processed/to_archive")
  
  write.csv(
    new.data,
    paste0(
      selected.dir,
      "\\",
      tools::file_path_sans_ext(new.name),
      ".csv"
    ),
    na = "",
    row.names = FALSE
  )
  
  }
}



```

```{r}

# Split into species-specific files
# unique(raw.data.list[[1]]$SpeciesCode)
# 
# aou_code <- read.csv(list.files("reference/IBPAOU", full.names = TRUE)) %>%
#   tibble::add_row(SPEC = "BALO", COMMONNAME = "band lost") %>%
#     tibble::add_row(SPEC = "BADE", COMMONNAME = "band destroyed")
#                      
# split.col<- "SpeciesCode"
# proj.prefix<- c("Eielson","JBER")
# data.desc<- "Banding"
# 
# split.dfs <- raw.data.list %>%
#   set_names(proj.prefix) %>%
#   purrr::map(.f = ~ .x %>%
#                dplyr::inner_join(
#                  .,
#                  aou_code[, c("COMMONNAME", "SPEC")],
#                  by = c("SpeciesCode" = "SPEC"),
#                  keep = FALSE,
#                  relationship = "many-to-many"
#                  ) %>%
#                split(.[["COMMONNAME"]]) %>%
#                purrr::map(., ~ .x %>%  dplyr::select(-"COMMONNAME"))) %>%
#   purrr::imap(., function(l, n) {
#     names(l) <- paste0(n, "_", gsub(" ", "_", names(l)))
#     return(l)
#     }) %>%
#   purrr::map(.f = ~ .x %>%
#       purrr::iwalk(., ~ write.csv(.x, 
#                                   file = paste0("data-processed/to_archive/",
#                                     .y, "_",data.desc,"_v1.csv"
#                                     ),
#                                    na = "",
#                                   row.names = FALSE)))
  

```

#### Dictionary

```{r}

 if ("list" %in% class(ref.dictionary)) {
   choose.format.app(
     formats =  c("CSV", "json (dictionary record)", "json (mdEditor file)"),
     filename = input.dxnry[["data"]][[1]][["meta"]][["title"]]
   )
   
   for (a in selected.format) {
     if (a == "CSV") {
       save.file.app(
         dir = paste0(
           dirname(rstudioapi::documentPath()),
           "/data-processed/to_archive"
         ),
         proposed.filename =  gsub(" ", "_", 
                              input.dxnry[["data"]][[1]][["meta"]][["title"]]), 
         ref.filename = input.dxnry[["data"]][[1]][["meta"]][["title"]],
         format = a
       )
       
        table.dxnry <- mdJSONdictio::build.table(input.dxnry)
    
        write.csv(
          table.dxnry,
          paste0(
            selected.dir, 
            "\\",
            new.name,
            "_",
            format(
              as.POSIXct(input.dxnry[["data"]][[1]][["attributes"]][["date-updated"]], tz = "UTC", "%Y-%m-%dT%H:%M:%OS"),
              '%Y%m%d-%H%M%S'
            ),
            ".csv"
          ),
          na = "",
          row.names = FALSE
        )
       
     } else if (a == "json (dictionary record)") {
       save.file.app(
         dir = paste0(
           dirname(rstudioapi::documentPath()),
           "/data-processed/to_archive"
         ),
         proposed.filename = gsub(" ", "_", 
                             input.dxnry[["data"]][[1]][["meta"]][["title"]]),
         ref.filename = input.dxnry[["data"]][[1]][["meta"]][["title"]],
         format = a
       )
       
       new.dxnry <- rjson::toJSON(x = input.dxnry)
       
       write(
         x = new.dxnry,
         file = paste0(
            selected.dir, 
            "\\",
            new.name,
           "_mdeditor-",
           format(
             as.POSIXct(input.dxnry[["data"]][[1]][["attributes"]][["date-updated"]], tz = "UTC", "%Y-%m-%dT%H:%M:%OS"),
             '%Y%m%d-%H%M%S'
           ),
           ".json"
         )
       )
       
     } else if (a == "json (mdEditor file)") {
       save.file.app(
         dir = paste0(
           dirname(rstudioapi::documentPath()),
           "/data-processed/to_archive"
         ),
         proposed.filename = gsub(
           "[[:punct:]]$",
           "",
           stringr::str_extract(dxnry.filename, "[^\\d]+")
         ),
         ref.filename = dxnry.filename,
         format = a
       )
       
       if (stringr::str_detect(new.name, 
                               stringr::regex('mdeditor', ignore_case = T)) 
           == FALSE) {
         new.name<- paste0(new.name, "_mdeditor")
       }
       
       new.dxnry <- rjson::toJSON(x = ref.dictionary)
       
       write(x = new.dxnry,
             file = paste0(
               selected.dir,
               "\\",
               new.name,
               "-",
               format(
                 as.POSIXct(ref.dictionary[["data"]][[record.num]][["attributes"]][["date-updated"]], tz = "UTC", "%Y-%m-%dT%H:%M:%OS"),
                 '%Y%m%d-%H%M%S'
               ),
               ".json"
             ))
       
     }
   }
 } else {
   choose.format.app(formats =  c("CSV", "json (dictionary record)"),
                     filename = dxnry.filename)
   
   for (a in selected.format) {
     if (a == "CSV") {
       save.file.app(
         dir = paste0(
           dirname(rstudioapi::documentPath()),
           "/data-processed/to_archive"
         ),
         proposed.filename = dxnry.filename,
         ref.filename = dxnry.filename,
         format = a
       )
    
        write.csv(
          input.dxnry,
          paste0(
            selected.dir, 
            "\\",
            new.name,
            "_",
            format(
              as.POSIXct(lubridate::now(), tz = "UTC", "%Y-%m-%dT%H:%M:%OS"),
              '%Y%m%d-%H%M%S'
            ),
            ".csv"
          ),
          na = "",
          row.names = FALSE
        )
       
     } else if (a == "json (dictionary record)") {
       save.file.app(
         dir = paste0(
           dirname(rstudioapi::documentPath()),
           "/data-processed/to_archive"
         ),
         proposed.filename = dxnry.filename,
         ref.filename = dxnry.filename,
         format = a
       )
       
       json.dxnry <- 
         mdJSONdictio::build.mdJSON(input.dxnry, 
                                    title = gsub("-", " ",
                                                 gsub("_", " ",
                                                      dxnry.filename)))
       new.dxnry <- rjson::toJSON(x = json.dxnry)
       
       write(
         x = new.dxnry,
         file = paste0(
            selected.dir, 
            "\\",
            new.name,
           "_mdeditor-",
           format(
             as.POSIXct(json.dxnry[["data"]][[1]][["attributes"]][["date-updated"]], tz = "UTC", "%Y-%m-%dT%H:%M:%OS"),
             '%Y%m%d-%H%M%S'
           ),
           ".json"
         )
       )
       
     }
   }
   
 }

     
 dxnry.name <-  sub(" ", "_", dxnry.name)
 dxnry.name <- strsplit(gsub("[[:punct:]]$", "", dxnry.filename), " +")[[1]]
 

```

# Archive in the Regional Data Repository

```{r}
project = "Yakutat"

```

```{r}


path<- rstudioapi::selectDirectory(path = current.dir)

project<- FWSAkRDRtools::find.projects(pattern = project)

commit <-
  FWSAkRDRtools::commit.files(project = project,
               local.path = path,
               recursive = TRUE)


```

# Create data packages

### Provide inputs

```{r}
# Provide a directory path for data file(s) to import
data.path <- "data-processed/to_archive/"

# Provide a directory path to write files
processed.data.path <- "data-processed/data_packages/Eielson_data_package_2024/"

# Provide versioning format, if applicable
vers.suffix <- "_v"

# Provide suffix for data files (i.e. year if filtering data)
data.suffix <- "_2024"

# Provide the relevant information (if applicable)
missing_values<- c(NA)
species.col <- "SpeciesCode"
date.col <- "Date"
date.format <-  "%m/%d/%Y"
time.col <- "CaptureTime_2400"
time.format <- "%H:%M"

```

### Import data

```{r}

data.list <- importer(
  path = data.path,
  multiple = TRUE,
  graphics = TRUE,
  column.format = list(
    c(column = date.col, format = date.format),
    c(column = time.col, format = time.format)
  )
)

metadata.list<- importer(
  path = data.path,
  multiple = TRUE,
  graphics = TRUE,
  column.format = list(
    c(column = date.col, format = date.format),
    c(column = time.col, format = time.format)
  )
)

```

### Filter data

```{r}

aou_code <- read.csv("reference/IBP-AOS-LIST23.csv", na = "")

## Translate species names/codes (if neccessary)
# 
# data.list <- data.list %>%
#   purrr::map(
#     .f = ~ .x %>%
#       dplyr::rename_with(., ~gsub('\\.', '', .x)) %>%
#       dplyr::inner_join(
#         .,
#         aou_code[, c("COMMONNAME", "SPEC")],
#         by = c("CommonName" = "COMMONNAME"),
#         keep = FALSE,
#         relationship = "many-to-many"
#       ) %>%
#       dplyr::mutate("SpeciesCode" = `SPEC`) %>%
#       dplyr::select(-"SPEC")
#   )

# Check species names/codes
data.list %>%
  purrr::map(.f = ~ .x %>%
               dplyr::filter(!.data[[species.col]] %in% aou_code$SPEC))


if(any(c(species.col, date.col) %in% unlist(purrr::map(data.list, colnames)))) {
  # Make filter selections
  if (species.col %in% unlist(purrr::map(data.list, colnames))) {
    species <- unique(unlist(sapply(data.list, "[", species.col)))
  } else {
    species<- character(0)
  }
  
  if (date.col %in% unlist(purrr::map(data.list, colnames))) {
    date.ranges <- data.list %>% purrr::map(
      .f = ~ .x %>%
        dplyr::select(all_of(date.col)) %>%
        dplyr::filter(.data[[date.col]] != "") %>%
        dplyr::mutate({{date.col}} := as.Date(.data[[date.col]], format = date.format))
      
      %>%
        append(list(
          "Min" = min(.[[date.col]]), "Max" = max(.[[date.col]])
        ))
    )
    
    dates.df <- data.list %>% purrr::map(
      .f = ~ .x %>%
        dplyr::select(all_of(date.col)) %>%
        dplyr::filter(.data[[date.col]] != "") %>%
        dplyr::mutate({{date.col}} := as.Date(.data[[date.col]], format = date.format))
    ) %>%
      dplyr::bind_rows(., .id = "id")
    
    min.date <- min(do.call("c", lapply(date.ranges, "[[", "Min")))
    max.date <- max(do.call("c", lapply(date.ranges, "[[", "Max")))
    
  } else {
    min.date <- as.Date("0001-01-01")
    max.date <- as.Date("0001-01-01")
  }


  filter.df.app(data.list, species.col, date.col, species, min.date, max.date)


# Apply filters
  id.var <- dplyr::last(make.names(c(unique(
    unlist(purrr::map(data.list, colnames))
  ), ".id"), unique = TRUE))
  
  data.list <- data.list %>%
    purrr::map(.f = ~ .x %>%
                 tibble::rowid_to_column(var = id.var))
  
  other.data.list <- data.list
  
  if (length(species) != 0) {
    data.list <- data.list %>%
      purrr::map(.f = ~ .x %>%
                   dplyr::filter(if_any(
                     matches(species.col), ~ .data[[species.col]] %in% selected.species
                   )))
  }
  
  
  if (min.date != "0001-01-01" |
      max.date != "0001-01-01") {
    data.list <- data.list %>%
      purrr::map(
        .f = ~ .x %>%
          dplyr::mutate_at(dplyr::vars(tidyselect::any_of(date.col)), as.Date, format = date.format) %>%
          dplyr::filter(if_any(
            matches(date.col),
            ~ data.table::between(.data[[date.col]], min(selected.dates), max(selected.dates), NAbounds =
                                    FALSE)
          ))
      )
  }
  
  rm.row <-  purrr::map(data.list, id.var)
  
  other.data.list <- lapply(seq_along(rm.row), function(i) {
    if (rlang::is_empty(rm.row[[i]]) == FALSE) {
      other.data.list[[i]] %>%
        dplyr::slice(-rm.row[[i]])
    } else {
      other.data.list[[i]]
    }
  }) %>%
    purrr::set_names(names(data.list))
  
    data.list <- purrr::map(data.list, ~ (.x %>% dplyr::select(-tidyselect::any_of(id.var))))
    
      other.data.list <- purrr::map(other.data.list, ~ (.x %>% dplyr::select(-tidyselect::any_of(id.var))))
   

# Remove empty dataframes
  data.list <- data.list %>%
    purrr::discard(., ~ nrow(.) == 0) %>%
    purrr::map(.f = ~ .x %>%
                 dplyr::mutate({{date.col}} := format(.data[[date.col]], format = date.format)))
  
  
  other.data.list <- other.data.list %>%
    purrr::discard(., ~ nrow(.) == 0) %>%
    purrr::map(.f = ~ .x %>%
                 dplyr::mutate({{date.col}} := format(.data[[date.col]], format = date.format)))
  
}

```

### Merge data

```{r}

if(length(data.list)>1) {
  # Make data frame selections
  merge.df.app(data.list)
  
  # Execute forced rbind
  if (is.null(selected.df) == FALSE) {
    data.list <-  c(list(merged_data = data.table::rbindlist(data.list[names(data.list) %in% selected.df], fill = TRUE)),
                        data.list[!names(data.list) %in% selected.df])
  }
  
  data.list <- data.list %>%
    purrr::map(
      .f = ~ .x %>%
        dplyr::mutate_if(is.character, ~ dplyr::na_if(., "")) %>%
        dplyr::mutate_if(is.character, ~ dplyr::na_if(., "NULL")) %>%
        dplyr::mutate_if(is.character, ~ dplyr::na_if(., missing_values))
    )
}

```

### Translate dictionaries

```{r}

dxnry.list <- list()

for (x in 1:length(metadata.list)) {
  if(is.data.frame(metadata.list[[x]])){
     dxnry.list[[x]] <- metadata.list[[x]]
     
  } else if (names(metadata.list[[x]]) == "data") {
    temp <-
      mdJSONdictio::extract.mdJSON(metadata.list[[x]], record.type = "dictionaries")
    
    names <- character(0)
    temp3 <- list()
    
    for (y in 1:length(temp[["data"]])) {
      temp2 <- list()
      
      temp2[["data"]][[1]] <- temp[["data"]][[y]]
      
      temp3[[y]] <- mdJSONdictio::build.table(temp2)
      
      names <- c(names, rjson::fromJSON(temp2[["data"]][[1]][["attributes"]][["json"]])[["dataDictionary"]][["citation"]][["title"]])
      
    }
    
    names(temp3) <- names
    
    dxnry.list[[x]] <- temp3
    
  }
}

names(dxnry.list)<- names(metadata.list)

```

### Export files

```{r}

fs::dir_create(names(which(!fs::dir_exists(processed.data.path))))
  
for(a in 1:length(data.list)) {
  write.csv(
    data.list[[a]],
    paste0(
      gsub("/$", "", processed.data.path),
      "/",
      gsub(paste0("(", vers.suffix, ").*"), "", names(data.list[a])),
      "_",
      gsub("^[^A-Za-z0-9]|[^A-Za-z0-9]$", "", data.suffix),
      ".csv"
    ),
    na = "",
    row.names = FALSE
  )
}

for(a in 1:length(dxnry.list)) {
  for(b in 1:length(dxnry.list[[a]])){
  write.csv(
    dxnry.list[[a]][[b]],
    paste0(
      gsub("/$", "", processed.data.path),
      "/",
      names(dxnry.list[[a]][b]),
      "_",
      format(Sys.time(), '%Y%m%d'),
      ".csv"
    ),
    na = "",
    row.names = FALSE
  )
  }
}




file.create(paste0(processed.data.path, "README.txt"), overwrite = TRUE)

sink(paste0(processed.data.path,"README.txt"))
cat(paste0("\nThis folder contains the following:\n\n"))

dirs <- list.dirs(processed.data.path,
                  recursive = TRUE,
                  full.names = FALSE)

for(a in 1:length(dirs)) {
  files <- basename(fs::dir_info(
    paste0(processed.data.path, dirs[a]),
    recurse = FALSE,
    type = "file"
  )$path)
  files<- files[!files =="README.txt"]
  
  if (length(files != 0)) {
   cat(paste0(dirs[a], "\n"), paste0(files, sep = "\n"))
    cat("\n")
  }
  
}

closeAllConnections()


```

# Publish to Science Base

# Capture data transformations

### Provide inputs

```{r}

# Provide the relevant information (if applicable)
missing_values<- c(NA)
species.col <- "SpeciesCode"
date.col <- "Date"
date.format <-  "%m/%d/%Y"
time.col <- "CaptureTime_2400"
time.format <- "%H:%M"
na.if<- c("","-","—",".","NA")

```

### Import data

```{r}


select.import.app(dir = paste0(current.dir,
                               "/data-processed/to_archive"), 
                  ext = c("txt", "csv", "xls*"),
                  multiple = TRUE,
                  type = "data")

data.list <- importer(
  path = selected.files$path[1],
  files = selected.files$filename,
  column.format = list(
    c(column = date.col, format = date.format),
    c(column = time.col, format = time.format)
  )
)


data.filenames<- tools::file_path_sans_ext(selected.files$filename)


```

### Filter data

```{r}

# Select data to transform
aou_code <- read.csv(list.files("reference/IBPAOU", full.names = TRUE)
                     , na = "")

# Remove NA rows
data.list <- data.list %>%
  purrr::map(.f = ~ .x %>%
               dplyr::mutate(dplyr::across(
                 where(is.character), ~ ifelse(. %in% na.if, NA, .)
               )) %>%
               dplyr::filter(!dplyr::if_all(colnames(.), is.na)))


if(any(c(species.col, date.col) %in% unlist(purrr::map(data.list, colnames)))) {
  # Make filter selections
  if (species.col %in% unlist(purrr::map(data.list, colnames))) {
    species <- unique(purrr::map(data.list, ~ .x %>%
                                   purrr::pluck(species.col)) %>%
                        unlist(.))
  } else {
    species<- character(0)
  }
  
  
  if (date.col %in% unlist(purrr::map(data.list, colnames))) {
    date.ranges <- data.list %>% purrr::map(
      .f = ~ .x %>%
        dplyr::mutate("{{date.col}}" := ifelse(!date.col %in% names(.), "", {{date.col}})) %>%
        dplyr::select(any_of(date.col)) %>%
        dplyr::filter(.data[[date.col]] != "") %>%
        dplyr::mutate_at(dplyr::vars(tidyselect::any_of(date.col)), as.Date, format = date.format) %>%
        # dplyr::mutate({{date.col}} := as.Date(.data[[date.col]], format = date.format)) %>%
        append(list(
          "Min" = min(.[[date.col]]), "Max" = max(.[[date.col]])
        ))
    )
    
  
    dates.df <- data.list %>% purrr::map(
      .f = ~ .x %>%
        dplyr::mutate({{date.col}} := ifelse(!date.col %in% names(.), "", .data[[date.col]])) %>%
        dplyr::select(all_of(date.col)) %>%
        dplyr::filter(.data[[date.col]] != "") %>%
        dplyr::mutate({{date.col}} := as.Date(.data[[date.col]], format = date.format))
    ) %>%
      dplyr::bind_rows(., .id = "id")
    
    min.date <- min(do.call("c", lapply(date.ranges, "[[", "Min")))
    max.date <- max(do.call("c", lapply(date.ranges, "[[", "Max")))
    
  } else {
    min.date <- as.Date("0001-01-01")
    max.date <- as.Date("0001-01-01")
  }
  
  


  filter.df.app(data.list, species.col, date.col, species, min.date, max.date)


# Apply filters
  id.var <- dplyr::last(make.names(c(unique(
    unlist(purrr::map(data.list, colnames))
  ), ".id"), unique = TRUE))
  
  
  data.list <- data.list %>%
    purrr::map(.f = ~ .x %>%
                 tibble::rowid_to_column(var = id.var))
  
  other.data.list <- data.list
  
  if (length(species) != 0) {
    data.list <- data.list %>%
      purrr::map(.f = ~ .x %>%
                   dplyr::filter(if_any(
                     matches(species.col), ~ .data[[species.col]] %in% selected.species
                   )))
  }
  
  
  if (min.date != "0001-01-01" |
      max.date != "0001-01-01") {
    data.list <- data.list %>%
      purrr::map(
        .f = ~ .x %>%
          dplyr::mutate_at(dplyr::vars(tidyselect::any_of(date.col)), as.Date, format = date.format) %>%
          dplyr::filter(if_any(
            matches(date.col),
            ~ data.table::between(.data[[date.col]], min(selected.dates), max(selected.dates), NAbounds =
                                    FALSE)
          ))
      )
  }
  
  rm.row <-  purrr::map(data.list, id.var)
  
  other.data.list <- lapply(seq_along(rm.row), function(i) {
    if (rlang::is_empty(rm.row[[i]]) == FALSE) {
      other.data.list[[i]] %>%
        dplyr::slice(-rm.row[[i]])
    } else {
      other.data.list[[i]]
    }
  }) %>%
    purrr::set_names(names(data.list))
  
  data.list <-
    purrr::map(data.list, ~ (.x %>%
                                   dplyr::select(-tidyselect::any_of(id.var))))
  
  other.data.list <- 
    purrr::map(other.data.list, ~ (.x %>% 
                                     dplyr::select(-tidyselect::any_of(id.var))))
   

# Remove empty dataframes
      data.list <- data.list %>%
        purrr::discard(., ~ nrow(.) == 0) %>%
        purrr::map(.f = ~ .x %>%
                           dplyr::mutate_at(dplyr::vars(tidyselect::any_of(date.col)), format, format = date.format))
                     
      
      other.data.list <- other.data.list %>%
        purrr::discard(., ~ nrow(.) == 0) %>%
         purrr::map(.f = ~ .x %>%
                           dplyr::mutate_at(dplyr::vars(tidyselect::any_of(date.col)), format, format = date.format))
  
}

```

### Biological samples inventory

#### Prepare data

```{r}

# Prep data
cap.data <- plyr::ldply(data.list)

cap.data %>%
  tibble::as_tibble() %>%
  dplyr::filter(if_all(c("Date", "Year"), is.na))

cap.data <- cap.data %>%
  dplyr::mutate_at("BandNumberOut", ~ as.character(.)) %>%
  dplyr::mutate_if(is.character, ~ dplyr::na_if(., "")) %>%
  dplyr::mutate_if(is.character, ~ dplyr::na_if(., "NULL")) %>%
  dplyr::mutate_if(is.character, ~ dplyr::na_if(., missing_values)) %>%
  tidyr::drop_na(BandNumberOut, Year)


# Load reference tables
ref_data <-
  lapply(list.files(path = "reference/biosample_inventory/", full.names = TRUE),
         utils::read.csv, na.strings = "")

names(ref_data) <-
  gsub(
    list.files(path = "reference/biosample_inventory/"),
    pattern = ".csv$",
    replacement = ""
  )

list2env(ref_data, globalenv())
  


```

#### Recode data

```{r}

# Correct colnames
cap.data <- cap.data %>%
  dplyr::rename_at(dplyr::vars(matches("Out$")), ~ stringr::str_remove(., "Out$")) %>%
  plyr::rename(
    replace = c(
      "Year" = "YearCollected",
      "Month" = "MonthCollected",
      "Day" = "DayCollected",
      "BanderID" = "Collector"
    )
  ) 
# %>%
# tidyr::unite("Feather",
#              c("FeatherType", "FeatherType1")[c("FeatherType", "FeatherType1") %in% names(cap.data)],
#              remove = TRUE,
#              na.rm = TRUE) %>%
# dplyr::mutate_at("Feather", ~ dplyr::na_if(., ""))

# Pivot sample columns

# Blood
  # c("BloodGenetics",
  # "BloodIsotope",
  # "BloodHg")
  
sample.cols <- c(
  "Feather",
  "AISwabO",
  "AISwabC",
  "Fecal",
  "Nails",
  "BloodFilter",
  "BloodCapTube",
  "BloodSmear"
)

cap.data <- cap.data %>%
  tidyr::pivot_longer(
    cols = tidyselect::any_of(sample.cols),
    names_to = "SampleType",
    values_to = "SampleTaken"
  ) %>%
  dplyr::mutate("FeatherType1" = ifelse(SampleType == "Feather", SampleTaken, NA)) %>%
  dplyr::mutate("SampleTaken" = ifelse(
    SampleType == "Feather" &
      is.na(FeatherType1) == FALSE,
    "Y",
    SampleTaken
  )) %>%
  dplyr::filter(SampleTaken %in% c("Y"))


```

#### Translate data

```{r}

# Auto fill sample details
cap.data<- cap.data %>%
  dplyr::left_join(.,sample_info,
                   by = "SampleType",
                   keep = FALSE,
                   relationship = "many-to-one")


# Add missing info
## OccurrenceID

cap.data %>%
  dplyr::filter_at("SpeciesCode", is.na)

cap.data %>%
  dplyr::filter(if_all(c("StateProvince", "City", "Area"), is.na))


cap.data<- cap.data %>%
  dplyr::mutate(
    "StorageMethod" = "frozen",
    "StorageLocation" = "Alaska MBM",
    "SampleStatus" = NA,
    "Contact" = NA,
    "Notes" = NA
  )


```

#### Join to inventory

```{r}

if(dir.exists(paste0("//ifw7ro-file.fws.doi.net/datamgt/mbm/"))==FALSE) {
  stop("Check the VPN connection and try again.")
}

selected.files<- fs::dir_ls(
 path =  "//ifw7ro-file.fws.doi.net/datamgt/mbm/mbmlb_012_Biological_Sample_Inventory/data/final_data"
)

if(length(selected.files)==1) {
  auth.data <- importer(path = dirname(selected.files),
                        files = basename(selected.files))[[1]]
  
  auth.filename <- tools::file_path_sans_ext(basename(selected.files))
  
} else {
  select.import.app(
    dir = "//ifw7ro-file.fws.doi.net/datamgt/mbm/mbmlb_012_Biological_Sample_Inventory/data/final_data",
    ext = c("txt", "csv", "xls*"),
    multiple = FALSE,
    type = "data"
  )
  
  auth.data <- importer(path = selected.files$path[1], files = selected.files$filename)[[1]]
  
   auth.filename <- tools::file_path_sans_ext(selected.files$filename)
  
}

# Remove inventoried data
  dplyr::inner_join(cap.data,
                    auth.data,
                    by = c("BandNumber",
                           "YearCollected"))

new.data <- auth.data %>%
  merge(cap.data, all = TRUE) %>%
  dplyr::select(colnames(auth.data))

```

#### Update sample status

```{r}


selected.files<- fs::dir_ls(
 path =  "reference/analysis_results"
)

if(length(selected.files)==1) {
  sex_results <- importer(path = dirname(selected.files),
                        files = basename(selected.files))[[1]]
  
  
} else {
  select.import.app(
    dir =  paste0(current.dir, "/reference/analysis_results"),
    ext = c("txt", "csv", "xls*"),
    multiple = FALSE,
    type = "data"
  )
  
  sex_results <- importer(path = selected.files$path[1], files = selected.files$filename)[[1]]

  
}


# filter for results
sex_results <- sex_results %>%
  dplyr::mutate_at("BandNumberOut", as.character) %>%
  # dplyr::filter(Sex %in% c("F", "M")) %>%
  dplyr::mutate_at("GenSex", ~ dplyr::recode(GenSex, "m" = "M", "f" = "F")) %>%
  dplyr::mutate(SampleTissue = "blood",
                SampleForm = "filter")


# Check for duplicate results
sex_results[c(
  anyDuplicated(sex_results$BandNumberOut, fromLast = TRUE),
  anyDuplicated(sex_results$BandNumberOut, fromLast = FALSE)
), ]


# new.data<- new.data %>%
#  dplyr::distinct(across(-OccurrenceID), .keep_all = TRUE)

# review<- new.data %>%
#   dplyr::group_by(BandNumber,
#                   SpeciesCode,
#                   SampleTissue,
#                   SampleForm,
#                   YearCollected) %>%
#   dplyr::filter(dplyr::n() > 1) %>%
#   dplyr::ungroup() 


# join sex results
new.data <- new.data %>%
  dplyr::left_join(
    .,
    dplyr::distinct(sex_results[, c("SpeciesCode",
                                    "BandNumberOut",
                                    "Year",
                                    # "YearCollected",
                                    "SampleTissue",
                                    "SampleForm",
                                    "GenSex")]),
    by = c(
      "SpeciesCode",
      "BandNumber" = "BandNumberOut",
      "YearCollected" = "Year",
      "SampleTissue",
      "SampleForm"
    ),
    keep = FALSE,
    relationship = "many-to-many"
  ) %>%
  dplyr::mutate_at(
    "SampleStatus",
    ~ dplyr::case_when(GenSex %in% c("F", "M") ~ "analysis complete", TRUE ~ SampleStatus)
  ) %>%
  dplyr::select(-"GenSex")

```

#### Export data

```{r}

# Save files - versioning
save.file.app(
  dir = paste0(
    dirname(rstudioapi::documentPath()),
    "/data-processed/to_archive"
  ),
  proposed.filename = paste0(auth.filename, " (1)"),
  ref.filename = auth.filename
)

write.csv(
  new.data,
  paste0(
    selected.dir,
    "\\",
    tools::file_path_sans_ext(new.name),
    ".csv"
  ),
  na = "",
  row.names = FALSE
)

```

### Bander Portal

```{r}

cap.data <- plyr::ldply(data.list)

# Reference tables
master_contacts <-
  read.csv(
    "reference/master_contacts.csv",
    na.strings = ""
  )

ref_data <-
  lapply(list.files(path = "reference/bander_portal/", full.names = TRUE),
         readxl::read_excel)

names(ref_data) <-
  gsub(
    list.files(path = "reference/bander_portal/"),
    pattern = ".xlsx$",
    replacement = ""
  )

list2env(ref_data, globalenv())


# Standardize missing values
cap.data <- cap.data %>%
  dplyr::mutate_if(is.character,  ~ dplyr::na_if(., "")) %>%
  dplyr::mutate_if(is.character,  ~ dplyr::na_if(., "NULL")) %>%
  dplyr::mutate_if(is.character,  ~ dplyr::na_if(., missing_values))

# Add missing columns
miss_cols <- setdiff(c(
  na.omit(capture_cols$my_col[!grepl("color", capture_cols$my_col)]),
  "TagType1",
  "TagType2",
  "TagStatus1",
  "TagStatus2"
),
colnames(cap.data))

cap.data[, (miss_cols) := as.character(NA)]

```

#### Recode data

```{r}

# How Obtained
recap.info <- cap.data %>%
  dplyr::filter(BBLDisposition == "R", !is.na(BBLCondition)) %>%
  dplyr::select(SpeciesCode, BBLDisposition, BBLCondition) %>%
  dplyr::distinct()

BBL_ref <-
  bander_portal_lookups %>%
  dplyr::filter(field == "How Obtained" &
                  status != "Discontinued") %>%
  dplyr::pull(code) %>%
  noquote() %>%
  unlist()

typical.num <- which(BBL_ref == "66")
  
  if (nrow(recap.info) != 0) {
    for (a in 1:nrow(recap.info)) {
      
    condition_desc <- bander_portal_lookups %>%
      dplyr::filter(field == "Bird/Band Condition" &
                      code == recap.info$BBLCondition[a]) %>%
      dplyr::pull(description) %>%
      unlist()
    
    obtained.choice <- 0
    while (obtained.choice == 0) {
      obtained.choice <-
        utils::menu(c(
          bander_portal_lookups %>%
            dplyr::filter(field == "How Obtained" &
                            status != "Discontinued") %>%
            dplyr::pull(description)
        ),
        title = cat(
          paste0(
            "\nThe bird/band condition of one or more recaptured ",
            recap.info$SpeciesCode[a] ,
            " was '",
            condition_desc,
            "'.\nHow were these birds obtained?\nSelect ",
            typical.num,
            " for typical recaptures (i.e., bird/band condition 'ALIVE - RELEASED/LEFT ON BIRD')\n"
          )
        ))
    }
    
    cap.data <- cap.data %>%
      dplyr::rowwise() %>%
      dplyr::mutate(
        "HowObtained" =
          ifelse(
            BBLDisposition == "R" &
              SpeciesCode == recap.info$SpeciesCode[a] &
              BBLCondition == recap.info$BBLCondition[a],
            BBL_ref[obtained.choice],
            
            HowObtained
          )
      )
    
  }
}

# Locations
loc.cols <- c("City", "Area")
loc.cols <- loc.cols[loc.cols %in% colnames(cap.data)]

cap.data <- cap.data %>%
  dplyr::mutate_at(
    dplyr::vars(which(names(.) %in% loc.cols)),
    ~ as.character(.) %>% dplyr::recode(
      .,
      "Eielson" = "EIEL",
      "Delta Junction" = "DELTA",
      "Eareckson" = "EARECK",
      "Beluga" = "BELUGA",
      "King Salmon" = "KING",
      "Nome" = "NOME",
      "Shemya" = "EARECK",
      "Anchorage" = "ANC",
      "Chicken" = "CHICKEN",
      "Fort Yukon" = "FORTYUK"
    )
  ) %>%
   dplyr::mutate(LocationID = dplyr::coalesce(!!!rlang::syms(loc.cols)))

BBL_locations <- c(
  "EIEL",
  "DELTA",
  "EARECK",
  "BELUGA",
  "KING",
  "NOME",
  "ANC",
  "JBER",
  "OTTER",
  "CHICKEN",
  "FORTYUK"
)

#!!# remove loc nums that correspond to the same locations
       
locs <- unique(cap.data$LocationID)[!unique(cap.data$LocationID) %in%
                                      BBL_locations]

for(a in locs) {
  loc.choice <-
    utils::menu(c(BBL_locations, "Enter a new location"),
                title = cat(
                  paste0(
                    "\n'",
                    a,
                    "' is not a listed BBL location ID.\nChoose an existing or new ID.\n"
                  )
                ))
  
  if (loc.choice == length(BBL_locations) + 1) {
    message(cat(
      paste0(
        "\nProvide a new location ID.\nThis will need to be added as a banding location in Bander Potal."
      )
    ))
    
    new_loc <-
      noquote(as.character(readline(prompt =)))
    
    
  } else {
    new_loc <- BBL_locations[loc.choice]
  }
  
  if (new_loc != "") {
    if (is.na(a)) {
      cap.data <-  cap.data %>%
        dplyr::mutate_at("LocationID",
                         ~ replace(.,
                                   is.na(.), new_loc))
    } else {
      cap.data <-  cap.data %>%
        dplyr::mutate_at("LocationID",
                         ~ replace(.,
                                   . == a, new_loc))
    }
  }
}

# Capture method
cap.data <- cap.data %>%
  dplyr::mutate_at(
    "CaptureMethod",
    ~ as.character(.) %>%
      dplyr::recode(
        .,
        "MN" = "Mist net",
        "WT" = "Walk-in trap",
        "HA" = "Hand capture",
        "BN" = "Bow net",
        "CN" = "Cannon net",
        "WN" = "Whoosh net",
        "NG" = "Net guns"
      )
  )

# Bander/Scribe ID
cap.data <- cap.data %>%
  dplyr::mutate_at(
    c("BanderID", "ScribeID"),
    ~ as.character(.) %>% plyr::mapvalues(
      .,
      from = master_contacts$FWSCode_FMLast,
      to = master_contacts$FWSCode_FML,
      warn_missing = FALSE
    )
  ) %>%
  dplyr::mutate_at(
    c("BanderID", "ScribeID"),
    ~ plyr::mapvalues(
      .,
      from = master_contacts$FWSCode_FLast,
      to = master_contacts$FWSCode_FL,
      warn_missing = FALSE
    )
  )

# Sex
cap.data <- cap.data %>%
  dplyr::mutate_at("FieldSex", ~ as.character(.)) %>%
  dplyr::rowwise() %>%
  dplyr::mutate_at("FieldSex",
                   ~ ifelse(
                     BBLDisposition == "R",
                     dplyr::recode(., "M" = "6", "F" = "7", "U" = "0"),
                     dplyr::recode(., "M" = "4", "F" = "5", "U" = "0")
                   ))


# Molt
cap.data <- cap.data %>%
  dplyr::mutate_at(
    "BodyMolt",
    ~ as.character(.) %>% dplyr::recode(
      .,
      "0" = "N",
      "1" = "Y",
      "2" = "Y",
      "3" = "Y",
      "4" = "Y"
    )
  ) %>%
  dplyr::mutate_at(
    "FlightFeatherMolt",
    ~ as.character(.) %>% dplyr::recode(
      .,
      "N" = "N",
      "S" = "Y",
      "A" = "Y",
      "J" = "Y"
    )
  )


# Leg attachments
leg_cols <-   c(
  "ULIn",
  "LLIn",
  "URIn",
  "LRIn",
  "ULOut",
  "LLOut",
  "UROut",
  "LROut"
)

leg_cols <- leg_cols[leg_cols %in% colnames(cap.data)]

color_cols<- paste0(leg_cols,"_colorband")


# test <- cap.data %>%
#   dplyr::filter_at(leg_cols, dplyr::any_vars(stringr::str_detect(., stringr::regex("-", ignore_case =
#                                                                                      TRUE))))

my_colors <-
  c(
    "Bk",
    "Br",
    "Db",
    "Lb",
    "Mb",
    "Dg",
    "Lg",
    "Gy",
    "Hp",
    "Mg",
    "Mv",
    "Or",
    "Pp",
    "Rd",
    "Wh",
    "Yl"
  )

BBL_colors <-
  c(
    "Black",
    "Brown",
    "Dark Blue",
    "Light Blue",
    "Blue",
    "Dark Green",
    "Light Green",
    "Gray",
    "Hot Pink",
    "Magenta",
    "Mauve",
    "Orange",
    "Purple",
    "Red",
    "White",
    "Yellow"
  )

cap.data <-  cap.data %>%
  dplyr::mutate_if(is.logical, as.factor) %>%
  dplyr::rowwise() %>%
  dplyr::mutate_at(
    c(leg_cols, "FlagColorIn", "FlagColorOut",  "FlagCodeColorIn", "FlagCodeColorOut"),
    ~ stringi::stri_replace_all_fixed(.,
                                      my_colors,
                                      BBL_colors,
                                      vectorize_all = FALSE)) %>%
  dplyr::mutate_at(leg_cols,
                   ~ stringi::stri_replace_all_fixed(.,
                                                     "-",
                                                     "/",
                                                     vectorize_all = FALSE)) %>%
  dplyr::mutate(across(.cols = all_of(leg_cols),
                       .names = "{paste0(color_cols[!grepl('Flag', color_cols)])}")) %>%
  dplyr::mutate_at(
    leg_cols,
    ~ stringi::stri_replace_all_fixed(.,
                                      BBL_colors,
                                      "01A",
                                      vectorize_all = FALSE)) %>%
  dplyr::mutate_at(leg_cols,
                   ~ stringi::stri_replace_all_fixed(.,
                                                     c("X", "Flag"),
                                                     c("00", "69"),
                                                     vectorize_all = FALSE)) %>%
  dplyr::mutate_at(
    color_cols,
    ~ ifelse(
      grepl("/", .),
      stringi::stri_replace_all_fixed(.,
                                      c("X", "Geo", "Flag"),
                                      "", vectorize_all = FALSE),
      stringi::stri_replace_all_fixed(.,
                                      c("X", "Geo", "Flag"),
                                      NA, vectorize_all = FALSE)
    )) %>%
  dplyr::mutate_at(color_cols,
                   ~ ifelse(
                     grepl("/$|.*/", .),
                     stringi::stri_replace_all_fixed(.,
                                                     "/",
                                                     "", vectorize_all = FALSE),
                     .
     
              ))


# ## Flag code colors by species - new/retained/replaced
# flag_info <- cap.data %>%
#   dplyr::filter(dplyr::if_any(c(leg_cols),
#                               ~ stringr::str_detect(., "69")))  %>%
#   dplyr::rowwise() %>%
#   dplyr::mutate(
#     FlagCodeChange = ifelse(FlagCodeIn != FlagCodeOut, 1, 0),
#     FlagIn = ifelse(is.na(FlagCodeIn) == FALSE, 1, 0),
#     FlagOut = ifelse(is.na(FlagCodeOut) == FALSE, 1, 0)
#   ) %>%
#   dplyr::select(any_of(c(
#     "SpeciesCode", "FlagIn", "FlagOut", "FlagCodeChange"
#   ))) %>%
#   dplyr::distinct() %>%
#   dplyr::mutate_all(as.character) %>%
#   dplyr::mutate_at("FlagCodeChange", ~ replace(., is.na(.), 0))
# 
# BBL_ref <- bander_portal_lookups$code[bander_portal_lookups$field ==
#                                         "Marker color"]
# BBL_ref<- BBL_ref[!grepl("/", BBL_ref)]
# 
# cap.data[, "FlagCodeIn_color"] <- NA
# cap.data[, "FlagCodeOut_color"] <- NA
# 
# if(nrow(flag_info)!=0) {
#   for (b in 1:nrow(flag_info)) {
#     flag_status <-
#       ifelse(flag_info$FlagIn[b] == 0,
#              "received a new",
#              "were recaptured with a")
#     
#     flag_status <-
#       ifelse(flag_info$FlagCodeChange[b] == 0,
#              flag_status,
#              "received a replacement")
#     
#     flag_status.2 <-
#       ifelse(flag_info$FlagCodeChange[b] == 0, "", "original ")
#     
#     
#     flag.choice <- 0
#     while (flag.choice == 0) {
#       flag.choice <-
#         utils::menu(c(BBL_ref), title = cat(
#           paste0(
#             "\nOne or more ",
#             flag_info$SpeciesCode[b] ,
#             " ",
#             flag_status,
#             " flag.\nWhat is the ",
#             flag_status.2 ,
#             "flag code color?\n"
#           )
#         ))
#     }
#     
#     if (flag_info$FlagIn[b] == 0) {
#       ### new
#       cap.data$FlagCodeOut_color[cap.data$SpeciesCode == flag_info$SpeciesCode[b] &
#                                    is.na(cap.data$FlagCodeIn) &
#                                    !is.na(cap.data$FlagCodeOut)] <-
#         BBL_ref[flag.choice]
#     } else {
#       if (flag_info$FlagCodeChange[b] == 1) {
#         flag.choice.2 <- 0
#         while (flag.choice.2 == 0) {
#           flag.choice.2 <-
#             utils::menu(c(BBL_ref), title = cat(paste0(
#               "\nWhat is the new flag code color?\n"
#             )))
#         }
#         
#         ### replaced
#         cap.data$FlagCodeIn_color[cap.data$SpeciesCode == flag_info$SpeciesCode[b] &
#                                     !is.na(cap.data$FlagCodeIn) &
#                                     cap.data$FlagCodeIn != cap.data$FlagCodeOut &
#                                     !is.na(cap.data$FlagCodeOut)] <-
#           BBL_ref[flag.choice]
#         
#         cap.data$FlagCodeOut_color[cap.data$SpeciesCode == flag_info$SpeciesCode[b] &
#                                      !is.na(cap.data$FlagCodeIn) &
#                                      cap.data$FlagCodeIn != cap.data$FlagCodeOut &
#                                      !is.na(cap.data$FlagCodeOut)] <-
#           BBL_ref[flag.choice.2]
#         
#       } else {
#         ### retained
#         cap.data$FlagCodeIn_color[cap.data$SpeciesCode == flag_info$SpeciesCode[b] &
#                                     !is.na(cap.data$FlagCodeIn) &
#                                     cap.data$FlagCodeIn == cap.data$FlagCodeOut] <-
#           BBL_ref[flag.choice]
#         
#         cap.data$FlagCodeOut_color[cap.data$SpeciesCode == flag_info$SpeciesCode[b] &
#                                      !is.na(cap.data$FlagCodeIn) &
#                                      cap.data$FlagCodeIn == cap.data$FlagCodeOut] <-
#           BBL_ref[flag.choice]
#       }
#       
#     }
#   }
# }


# Tags
tag_types <- c("Geo", "Radio", "GPS", "PTT")
tag_info <- cap.data %>%
  dplyr::filter_at(c("TagType1", "TagType2"),
                   dplyr::any_vars(stringr::str_detect(
                     ., stringr::regex(paste(tag_types, collapse = '|'), ignore_case =
                                         TRUE)
                   ))) %>%
  tidyr::pivot_longer(cols = starts_with("TagType"),
                      names_to = "TagType.col",
                      values_to = "TagType") %>%
  tidyr::pivot_longer(
    cols = starts_with("TagStatus"),
    names_to = "TagStatus.col",
    values_to = "TagStatus"
  ) %>%
  dplyr::filter(readr::parse_number(TagStatus.col) == readr::parse_number(TagType.col)) %>%
  dplyr::filter(!is.na(TagType)) %>%
  dplyr::select(SpeciesCode,
                TagType,
                TagStatus) %>%
  dplyr::distinct()

if(nrow(tag_info)!=0){
  for(a in 1:nrow(tag_info)) {
    BBL_tags <-
      unlist(strsplit(capture_recode$BBL_code[capture_recode$my_code == tag_info$TagType[a]], ","))
    
    BBL_ref <- bander_portal_lookups %>%
      dplyr::filter(field == "Marker Type", code %in% BBL_tags) %>%
      dplyr::pull(description) %>%
      stringr::str_extract_all("(?<=\\().+?(?=\\))") %>%
      unlist()
    
    tag.choice <- 0
    while (tag.choice == 0) {
      tag.choice <-
        utils::menu(c(BBL_ref), title = cat(
          paste0(
            "\nA ",
            tolower(tag_info$TagType[a]),
            " tag was ",
            tolower(tag_info$TagStatus[a]),
            " on one or more ",
            tag_info$SpeciesCode[a],
            ".\nHow was this marker attached?\n"
          )
        ))
    }
    
    cap.data <- cap.data %>%
      dplyr::rowwise() %>%
      dplyr::mutate_at(
        c(leg_cols, "TagType1"),
        ~ ifelse(
          !is.na(TagType1) &
            SpeciesCode == tag_info$SpeciesCode[a] &
            TagStatus1 == tag_info$TagStatus[a],
          stringi::stri_replace_all_regex(., tag_info$TagType[a], BBL_tags[tag.choice], vectorize_all = FALSE),
          
          .
        )
      ) %>%
      dplyr::mutate_at(
        c(leg_cols, "TagType2"),
        ~ ifelse(
          !is.na(TagType2) &
            SpeciesCode == tag_info$SpeciesCode[a] &
            TagStatus2 == tag_info$TagStatus[a],
          stringi::stri_replace_all_regex(., tag_info$TagType[a], BBL_tags[tag.choice], vectorize_all = FALSE),
          .
        )
      )
  }
}


```

#### Quality checks

```{r}

# Species measurements
if(class(bander_portal_species_measurements$bandSize)!= "list") {
  bander_portal_species_measurements <-
    bander_portal_species_measurements %>%
    dplyr::mutate_if(is.numeric, ~ dplyr::na_if(., 0)) %>%
    dplyr::rowwise() %>%
    dplyr::mutate_at("bandSize", ~ ifelse(grepl(", ", .), as.list(strsplit(., ", ")), as.list(.)))
}


mydata_col <- c("Mass_g", "Wing_mm", "Tail_mm")
BBL_col <-
  c(
    "femaleMass_min_g",
    "femaleMass_max_g",
    "maleMass_min_g",
    "maleMass_max_g",
    "femaleWingChord_min_mm",
    "femaleWingChord_max_mm",
    "maleWingChord_min_mm",
    "maleWingChord_max_mm",
    "femaleTailLength_min_mm",
    "femaleTailLength_max_mm",
    "maleTailLength_min_mm",
    "maleTailLength_max_mm"
  )

for (x in unique(cap.data$SpeciesCode)) {
  species.num <-
    which(bander_portal_species_measurements$alphaCode == x)
  if (length(species.num) == 0) {
    warning(paste0("SpeciesCode '", x, "' is invalid."))
  } else {
    sp_mydata <- cap.data %>%
      dplyr::filter(SpeciesCode == x) %>%
      dplyr::rowwise() %>%
      dplyr::mutate_at("FieldSex",
                       ~ dplyr::recode(
                         .,
                         "4" = "M",
                         "5" = "F",
                         "6" = "M",
                         "7" = "F" ,
                         "0" = "U"
                       ))
    
    sp_BBLdata <- bander_portal_species_measurements %>%
      dplyr::filter(alphaCode == x)
    
    ### Band size
    if (length(setdiff(unique(na.omit(
      sp_mydata$BandSize
    )), unlist(na.omit(
      sp_BBLdata$bandSize
    ))))
    != 0) {
      warning(
        paste0(
          "One or more ",
          sp_BBLdata$commonName,
          " has a band size inconsistent with BBL recommendations."
        )
      )
    }
    
    ## Mass/Wing chord/Tail
    for (a in c("Mass", "Wing", "Tail")) {
      for (b in  c("F", "M", "U")) {
        for (c in c("min", "max")) {
          mydata_col.match <- mydata_col[grep(a, mydata_col)]
          
          if (b == "U") {
            BBL_col.match <-
              BBL_col[grepl(a, BBL_col) & grepl(c, BBL_col)]
            sex_code <- c("U", NA)
            sex <- "unknown sex"
            
          } else {
            BBL_col.match <-
              BBL_col[grepl(a, BBL_col) &
                        grepl(paste0("^", b), BBL_col, ignore.case = TRUE) &
                        grepl(c, BBL_col)]
            sex_code <-  b
            sex <-
              grep(
                paste0("^", b),
                c("female", "male"),
                ignore.case = TRUE,
                value = TRUE
              )
            
          }
          fun <- match.fun(c)
          sign <- c("<", ">")[grep(c, c("min", "max"))]
          comparative <-
            c("smaller", "greater")[grep(c, c("min", "max"))]
          measurement <-
            c("mass", "wing chord", "tail")[grep(a, c("Mass", "Wing", "Tail"))]
          
          
          if (is.na(sum(sp_BBLdata[which(colnames(sp_BBLdata) %in% BBL_col.match)])) == FALSE) {
            if (length(na.omit(sp_mydata[[mydata_col.match]][sp_mydata$FieldSex %in% sex_code])) !=
                0) {
              if (get(sign)(fun(sp_BBLdata[which(colnames(sp_BBLdata) %in% BBL_col.match)]), fun(na.omit(sp_mydata[[mydata_col.match]][sp_mydata$FieldSex %in% sex_code]))) ==
                  FALSE)
              {
                {
                  warning(
                    paste0(
                      "One or more ",
                      sp_BBLdata$commonName,
                      " (",
                      sex,
                      ") have a ",
                      comparative,
                      " ",
                      measurement,
                      " than expected for the species (",
                      c,
                      ": ",
                      fun(sp_BBLdata[which(colnames(sp_BBLdata) %in% BBL_col.match)]),
                      " ",
                      sub('.*\\_', '', BBL_col.match[1]),
                      ")"
                    )
                  )
                }
              }
            }
          }
        }
      }
    }
  }
}


# lookup tables

lookups <- c("Age",
             "Sex",
             "How Aged",
             "How Sexed",
             "Marker Type",
             "Marker color")

for (a in lookups) {
  b <- which(capture_cols$BBL_col[capture_cols$BBL_template=="band_aux"] == a)
  for (c in b) {
    invalid.num <-
      which(!na.omit(cap.data[[capture_cols$my_col[c]]]) %in% bander_portal_lookups$code[bander_portal_lookups$field == a])
    
    invalid.index <-
      setdiff(1:nrow(cap.data), which(is.na(cap.data[[capture_cols$my_col[c]]])))[invalid.num]
    
    entry.skip <- c()
    
    if (length(invalid.index) != 0) {
      for (d in invalid.index) {
        if (d %in% entry.skip) {
          next
        } else{
          entry.choice <-
            utils::menu(c("replace these entries",
                          "do nothing"),
                        title = cat(
                          paste0(
                            "\n'",
                            cap.data[[capture_cols$my_col[c]]][d],
                            "' (",
                            capture_cols$my_col[c],
                            ") is not listed as a BBL ",
                            a,
                            "\nHow would you like to proceed?"
                          )
                        ))
          
          descriptions <- bander_portal_lookups$description[bander_portal_lookups$field == a]
          if (all(is.na(descriptions))) {
            descriptions <- ""
          } else {
            descriptions<- paste0(" - ", descriptions)
          }
          
          if (entry.choice == 1) {
            replace.choice <-
              utils::menu(c(
                paste0(
                  bander_portal_lookups$code[bander_portal_lookups$field == a],
                  descriptions
                )
              ),
              title = cat(paste0(
                "\nChoose a replacement value."
              )))
            
            if (replace.choice != 0) {
              entry.skip <- c(entry.skip,
                              which(cap.data[[capture_cols$my_col[c]]] == cap.data[[capture_cols$my_col[c]]][d]))
              
              cap.data[[capture_cols$my_col[c]]] <-
                replace(cap.data[[capture_cols$my_col[c]]],
                        cap.data[[capture_cols$my_col[c]]] %in% cap.data[[capture_cols$my_col[c]]][d],
                        bander_portal_lookups$code[bander_portal_lookups$field == a][replace.choice])
            }
            
          } else if (entry.choice == 2) {
            entry.skip <- c(entry.skip, which(cap.data[[capture_cols$my_col[c]]] == cap.data[[capture_cols$my_col[c]]][d]))
          }
        }
      }
    }
  }
}

# Bird status
## Status codes
status_info <- cap.data %>%
  dplyr::rowwise() %>%
  dplyr::mutate_at("TagType1", ~ ifelse(TagStatus1 %in% c("Removed", "Lost"), TagType1 <-
                                          NA
                                        , .)) %>%
  dplyr::mutate_at("TagType2", ~ ifelse(TagStatus2  %in% c("Removed", "Lost"), TagType2 <-
                                          NA
                                        , .)) %>%
  dplyr::select(
    "OccurrenceID",
    "BandNumberOut",
    "SpeciesCode",
    "BBLStatus",
    leg_cols[grepl("Out", leg_cols)],
    "TagType1",
    "TagType2",
    "Blood",
    "BBLStatus"
  ) %>%
  dplyr::mutate_at(dplyr::vars(c(
    leg_cols[grepl("Out", leg_cols)], "TagType1", "TagType2", "BBLStatus"
  )), ~as.character(.) %>% dplyr::na_if(.,"")) %>%
  dplyr::mutate(NumMarkers = sum(!unique(dplyr::c_across(
    c(leg_cols[grepl("Out", leg_cols)], "TagType1", "TagType2")
  ))  %in% c("00", NA))) %>%
  dplyr::mutate_at(
    "BBLStatus",
    ~ dplyr::case_when(
      NumMarkers >= 2 &
        Blood != "Y" ~ "325",
      NumMarkers >= 1 &
        Blood == "Y" ~ "319",
      NumMarkers == 0 &
        Blood == "Y" ~ "318",
      NumMarkers == 0 &
        Blood != "Y" ~ "300",
      NumMarkers == 1 &
        dplyr::if_any(c(leg_cols[grepl("Out", leg_cols)], "TagType1", "TagType2"), ~
                        . == "69") &
        Blood != "Y" ~ "369",
      NumMarkers == 1 &
        dplyr::if_any(c(leg_cols[grepl("Out", leg_cols)], "TagType1", "TagType2"), ~
                        grepl("^81", .)) &
        Blood != "Y" ~ "381",
      NumMarkers == 1 &
        dplyr::if_any(c(leg_cols[grepl("Out", leg_cols)], "TagType1", "TagType2"), ~
                        grepl("^80", .)) &
        Blood != "Y" ~ "380",
      NumMarkers == 1 &
        dplyr::if_any(c(leg_cols[grepl("Out", leg_cols)], "TagType1", "TagType2"), ~
                        grepl("^90", .)) &
        Blood != "Y" ~ "390",
      NumMarkers == 1 &
        dplyr::if_any(c(leg_cols[grepl("Out", leg_cols)], "TagType1", "TagType2"), ~
                        grepl("^01", .)) &
        Blood != "Y" ~ "301",
      TRUE ~ BBLStatus
    )
  ) %>%
  dplyr::rename(`temp:BBLStatus` = BBLStatus)


cap.data <- cap.data %>%
  dplyr::left_join(
    .,
    status_info[, c("OccurrenceID", "temp:BBLStatus")],
    by = c("OccurrenceID" = "OccurrenceID"),
    keep = FALSE,
    relationship = "one-to-one"
  ) %>%
  dplyr::mutate_at("BBLStatus", ~ as.character(.))

invalid.index <- which((cap.data$BBLStatus != cap.data$`temp:BBLStatus`) |
                         is.na(cap.data$BBLStatus))


if (length(invalid.index) != 0) {
  for (a in invalid.index) {
    message(cat(
      paste0(
        "The status for the following individual appears invalid.\nSpecies: ",
        cap.data$SpeciesCode[a],
        "\nBand number: ",
        cap.data$BandNumberOut[a],
        "\nStatus (reported): ",
        cap.data$BBLStatus[a],
        " [",
        bander_portal_lookups$description[grepl(stringr::str_sub(cap.data$BBLStatus[a], 1, 1),
                                                bander_portal_lookups$code) == TRUE &
                                            bander_portal_lookups$field == "Bird Status"],
        " ",
        bander_portal_lookups$description[grepl(stringr::str_sub(cap.data$BBLStatus[a], 2, 3),
                                                bander_portal_lookups$code) == TRUE &
                                            bander_portal_lookups$field == "Bird Status Extra Info"],
        "]",
        "\nStatus (detected): ",
        cap.data$`temp:BBLStatus`[a],
        " [",
        bander_portal_lookups$description[grepl(
          stringr::str_sub(cap.data$`temp:BBLStatus`[a], 1, 1),
          bander_portal_lookups$code
        ) == TRUE &
          bander_portal_lookups$field == "Bird Status"],
        " ",
        bander_portal_lookups$description[grepl(
          stringr::str_sub(cap.data$`temp:BBLStatus`[a], 2, 3),
          bander_portal_lookups$code
        ) == TRUE &
          bander_portal_lookups$field == "Bird Status Extra Info"],
        "]",
        "\nBlood: ",
        cap.data$Blood[a],
        "\nULOut: ",
        cap.data$ULOut[a],
        "\nLLOut: ",
        cap.data$LLOut[a],
        "\nUROut: ",
        cap.data$UROut[a],
        "\nLROut: ",
        cap.data$LROut[a],
        "\nTagStatus1: ",
        cap.data$TagStatus1[a],
        "\nTagStatus2: ",
        cap.data$TagStatus2[a]
      )
    ))
    
    status.choice <- utils::menu(
      choice = c("update as detected", "manually edit", "leave as is"),
      graphics = TRUE,
      title = cat("\nHow would you like to proceed?\n")
    )
    
    if (status.choice == 1) {
      cap.data$BBLStatus[a] <- cap.data$`temp:BBLStatus`[a]
      
      
    }
    if (status.choice == 2) {
      message(
        "\nIMPROTANT: you must click 'syncronise' (arrow loop icon) and then 'Done' to enact changes."
      )
      
      invalid.entries <- cap.data %>%
        dplyr::ungroup() %>%
        dplyr::filter(dplyr::row_number() == a)
      
      invalid.entries.edit <-
        DataEditR::data_edit(
          invalid.entries,
          col_readonly = colnames(invalid.entries)[!colnames(invalid.entries) %in% "BBLStatus"],
          viewer = "pane"
        ) %>%
        dplyr::mutate_at("BBLStatus", as.character)
      
      cap.data[match(invalid.entries.edit$OccurrenceID,
                     cap.data$OccurrenceID), "BBLStatus"] <-
        invalid.entries.edit$BBLStatus
    }
    
  }
}


cap.data <-
  cap.data %>%
  dplyr::select(-tidyselect::starts_with("temp:"))


# Other
## Band number
for(a in c("In", "Out")) {
  invalid_code.num <-
    which(!c(nchar(na.omit(
      gsub("-", "", cap.data[[paste0("BandNumber", a)]])
    ))) %in% c(8, 9))
  
  dir <-
    c("incoming", "outgoing")[grep(a, c("BandNumberIn", "BandNumberOut"))]
  
  if (length(invalid_code.num) != 0) {
    warning(
      paste0(
        "The following ",
        dir,
        " band numbers appear invalid. BBL issued federal metal bands should have 8 or 9 digits.\n"
      ),
      paste0(na.omit(cap.data[[paste0("BandNumber", a)]])[invalid_code.num], sep = "\n")
    )
  }
}

invalid_bands <- cap.data %>%
  dplyr::group_by(BandNumberOut, BBLDisposition) %>%
  dplyr::filter(BBLDisposition == "1",
                dplyr::n() > 1)

for(a in unique(invalid_bands$BandNumberOut)) {
  unique(invalid_bands$SpeciesCode[invalid_bands$BandNumberOut == a])
  warning(
    paste0(
      "Band number ",
      a,
      " is duplicated for the following species/flag codes.\n"
    ),
    paste0(
      paste0(
        invalid_bands$SpeciesCode[which(invalid_bands$BandNumberOut == a)],
        " (",
        invalid_bands$FlagCodeOut[which(invalid_bands$BandNumberOut == a)],
        ")"
      ),
      sep = "\n"
    )
  )
}

## Date
if(NA %in% cap.data$Day |
   NA %in% cap.data$Month |
   NA %in% cap.data$Year) {
  if ("Date" %in% colnames(cap.data)) {
    if (!NA %in% suppressWarnings(chron::chron(cap.data$Date, format = gsub("%", "",date.format)))) {
      
      date.choice <-
        utils::menu(c("please do", "absolutely not"),
                    title = cat(
                      paste0(
                        "\nThe variables Day, Month, and/or Year are incomplete. Would you like to parse this information from Date?\n"
                      )
                    ))
    
      if(date.choice==1) {
        cap.data <-
          cap.data %>%
          dplyr::mutate(
            Date = as.Date(Date, date.format),
            Day = lubridate::day(Date),
            Month = lubridate::month(Date),
            Year = lubridate::year(Date)
          )
      }
    } else {
      warning("The variables Day, Month, and/or Year are incomplete.")
    }
  } else {
    warning("The variables Day, Month, and/or Year are incomplete.")
  }
}

## Time
time.num <-
  which(is.na(lubridate::hm(cap.data$CaptureTime_2400, quiet = TRUE)) |
          grepl("[A-Za-z]", cap.data$CaptureTime_2400) == TRUE)

if (length(time.num) != 0) {
  warning(
    paste0(
      "One or more time entries (CaptureTime_2400) are invalid. Time must be formatted as 24:00.\n"
    ),
    paste0(unique(cap.data$CaptureTime_2400[time.num]), sep = "\n")
  )
}

              
## Bander ID/Scribe ID
for (a in c("BanderID", "ScribeID")) {
  invalid_code.num <- which(nchar(unique(cap.data[[a]])) > 3)
  
  for (b in invalid_code.num) {
    message(cat(
      paste0(
        "'",
        unique(cap.data[[a]])[b],
        "' (",
        a,
        ") appears invalid.\nProvide a replacement value. Press Enter to retain the current value."
      )
    ))
    
    new_code <-
      noquote(as.character(readline(prompt =)))
    
    if (new_code != "") {
      cap.data <-  cap.data %>%
        dplyr::mutate_at(c("BanderID", "ScribeID"),
                         ~ replace(., . == unique(cap.data[[a]])[b], new_code))
    }
  }
}


# Check bands are issued under permit
browseURL("https://www.pwrc.usgs.gov/bbl/Bander_portal/login/page0.php")

band_inventory<- read.csv("reference/bander_portal/BInv4Bandit_24111.csv")

cap.data %>%
  dplyr::filter(!BandNumberOut %in% band_inventory$BAND.NUMBER)


```

#### Translate data

```{r}

# Partition by disposition
cap.data_aux <-
  cap.data %>% dplyr::filter(dplyr::if_any(leg_cols, ~ !. %in% c(NA, "00")))

cap.data_no_aux <-
  cap.data %>% dplyr::filter(dplyr::if_all(leg_cols, ~ . %in% c(NA, "00")))

cap.data_aux <- split(cap.data_aux, cap.data_aux$BBLDisposition)
cap.data_no_aux <- split(cap.data_no_aux, cap.data_no_aux$BBLDisposition)

#!!# Improve process for no data
#!!# 8 just reports band number in band_no_aux
for (a in c("cap.data_aux", "cap.data_no_aux")) {
  for (b in c("5")) {            #"1", "R"
    if (!b %in% names(eval(as.name(a)))) {
      temp <-  list(rbind(tibble::as_tibble(
        matrix(
          ncol = length(cap.data),
          nrow = 0,
          dimnames = list(NULL, colnames(cap.data))
        ), check.names = FALSE
      )))
      
      names(temp) <- b
      
      temp<- c(eval(as.name(a)),temp)
      
      assign(a, temp)
      
    }
  }
}


# Merge 1, 5, 8, & F dispositions
#!# If original band not in inventory, 5 -> 1 & manually fill marker at capture
cap.data_no_aux$'1' <-
  tibble::as_tibble(data.table::rbindlist(cap.data_no_aux[c('1', '8')]))
cap.data_no_aux <- suppressWarnings(within(cap.data_no_aux, rm('8')))

cap.data_aux$'1' <-
  tibble::as_tibble(data.table::rbindlist(cap.data_aux[c('1', '8')]))
cap.data_aux <- suppressWarnings(within(cap.data_aux, rm('8')))

cap.data_no_aux$'R' <-
  tibble::as_tibble(data.table::rbindlist(cap.data_no_aux[c('R', 'F', '5')]))
cap.data_no_aux <- suppressWarnings(within(cap.data_no_aux, rm('F', '5')))

cap.data_aux$'R' <-
  tibble::as_tibble(data.table::rbindlist(cap.data_aux[c('R', 'F', '5')]))
cap.data_aux <- suppressWarnings(within(cap.data_aux, rm('F', '5')))

## Correct variable classes
cap.data_aux <- lapply(cap.data_aux, function(x) {
  x %>%
    dplyr::mutate_if(is.character, ~ dplyr::na_if(., ''))
})
cap.data_no_aux <- lapply(cap.data_no_aux, function(x) {
  x %>%
    dplyr::mutate_if(is.character, ~ dplyr::na_if(., ''))
})


# Complete templates
# Create empty dataframes
template_filenames <-
  list.files("templates/bander_portal",
             pattern = ".xlsx",
             full.names = TRUE)
template_colnames <-
  lapply(template_filenames, function(x)
    suppressWarnings(colnames(readxl::read_excel(x))))
names(template_colnames) <- basename(template_filenames)


band_no_aux_col <-
  unlist(template_colnames[names(template_colnames)
                    [grepl("band", names(template_colnames), ignore.case = TRUE) &
                     grepl("no", names(template_colnames), ignore.case = TRUE)]])
band_no_aux <-
  data.frame(matrix(
    ncol = length(band_no_aux_col),
    nrow = nrow(cap.data_no_aux$'1'),
    dimnames = list(NULL, band_no_aux_col)
  ),
  check.names = FALSE)

recap_no_aux_col <-
  unlist(template_colnames[names(template_colnames)
                    [grepl("recap", names(template_colnames), ignore.case = TRUE) &
                     grepl("no", names(template_colnames), ignore.case = TRUE)]])
recap_no_aux <-
  data.frame(matrix(
    ncol = length(recap_no_aux_col),
    nrow = nrow(cap.data_no_aux$'R'),
    dimnames = list(NULL, recap_no_aux_col)
  ),
  check.names = FALSE)

band_aux_col <-
  unlist(template_colnames[names(template_colnames)
                    [grepl("band", names(template_colnames), ignore.case = TRUE) &
                     grepl("aux", names(template_colnames), ignore.case = TRUE)]])
band_aux <-
  data.frame(matrix(
    ncol = length(band_aux_col),
    nrow = nrow(cap.data_aux$'1'),
    dimnames = list(NULL, band_aux_col)
  ),
  check.names = FALSE)

recap_aux_col <-
  unlist(template_colnames[names(template_colnames)
                    [grepl("recap", names(template_colnames), ignore.case = TRUE) &
                     grepl("aux", names(template_colnames), ignore.case = TRUE)]])
recap_aux <-
  data.frame(matrix(
    ncol = length(recap_aux_col),
    nrow = nrow(cap.data_aux$'R'),
    dimnames = list(NULL, recap_aux_col)
  ),
  check.names = FALSE)

band_no_aux_ref<- capture_cols %>% dplyr::filter(BBL_template=="band_no_aux")
recap_no_aux_ref<- capture_cols %>% dplyr::filter(BBL_template=="recap_no_aux")
band_aux_ref<- capture_cols %>% dplyr::filter(BBL_template=="band_aux")
recap_aux_ref<- capture_cols %>% dplyr::filter(BBL_template=="recap_aux")



## Fill banding - no aux
if(nrow(cap.data_no_aux$'1')>1) {
  auto.match.num <-
    which(
      band_no_aux_ref$BBL_col %in% colnames(band_no_aux) &
        !is.na(band_no_aux_ref$my_col)
    )
  
  
  for (a in na.omit(auto.match.num)) {
    band_no_aux[[band_no_aux_ref$BBL_col[a]]] <-
      cap.data_no_aux$'1'[[band_no_aux_ref$my_col[a]]]
  }
  
  band_no_aux[["Replaced Band Number"]] <-
    unlist(cap.data_no_aux$'1'["BandNumberIn"])
  
  banded_legs <- c()
  banded_legs <-
    c(banded_legs, sapply(1:nrow(cap.data_no_aux$'1'), function(x)
      substring(colnames(cap.data_no_aux$'1'[leg_cols])[cap.data_no_aux$'1'[x, leg_cols] %in% "00"][1], 2, 2)))
  
  band_no_aux[["Banded Leg"]] <- banded_legs
}

## Fill recaptures - no aux
if(nrow(cap.data_no_aux$'R')>1) {
auto.match.num <-
  which(recap_no_aux_ref$BBL_col %in% colnames(recap_no_aux) &
          !is.na(recap_no_aux_ref$my_col))


for (a in na.omit(auto.match.num)) {
  recap_no_aux[[recap_no_aux_ref$BBL_col[a]]] <-
    cap.data_no_aux$'R'[[recap_no_aux_ref$my_col[a]]]
}

banded_legs <- c()
banded_legs <-
  c(banded_legs, sapply(1:nrow(cap.data_no_aux$'R'), function(x)
    substring(colnames(cap.data_no_aux$'R'[leg_cols])[cap.data_no_aux$'R'[x, leg_cols] %in% "00"][1], 2, 2)))

recap_no_aux[["Banded Leg"]]<- banded_legs
}

## Fill banding - aux
if(nrow(cap.data_aux$'1')>1) {
  auto.match.num <-
    which(band_aux_ref$BBL_col %in% colnames(band_aux) &
            !is.na(band_aux_ref$my_col))
  
  manual.match.num <-
    which(!band_aux_ref$BBL_col %in% colnames(band_aux))
  
  
  for (a in na.omit(auto.match.num)) {
    band_aux[[band_aux_ref$BBL_col[a]]] <-
      cap.data_aux$'1'[[band_aux_ref$my_col[a]]]
  }
  
  
  for (a in 1:nrow(cap.data_aux$'1')) {
    n <- 0
    marker_string <- c()
    for (b in c(leg_cols[grepl("Out", leg_cols)])) {
      if (!is.na(cap.data_aux$'1'[a, b])) {
        for (c in unlist(strsplit(paste0(cap.data_aux$'1'[a, b]), "/"))) {
          n <- n + 1
          marker_cols <-
            colnames(band_aux)[grepl(n, colnames(band_aux))]
          
          band_aux[[marker_cols[1]]][a] <-
            c
          
          code_col <-
            "FlagCodeOut"[sapply("69", function(x)
              grepl(x, c))]
          if (length(code_col) != 0) {
            band_aux[[marker_cols[2]]][a] <-
              cap.data_aux$'1'[[code_col]][a]
          }
          
          color_col <-
            c("FlagColorOut", paste0(b, "_colorband"))[sapply(c("69", "01A"), function(x)
              grepl(x, c))]
          if (length(color_col) != 0) {
            band_aux[[marker_cols[3]]][a] <-
              cap.data_aux$'1'[[color_col]][a]
          }
          
          code_color_col <-
            "FlagCodeColorOut"[sapply("69", function(x)
              grepl(x, c))]
          if (length(code_color_col) != 0) {
            band_aux[[marker_cols[4]]][a] <-
              cap.data_aux$'1'[[code_color_col]][a]
          }
          
          band_aux[[marker_cols[5]]][a] <-
            c("L", "R")[sapply(c(".{1}L", ".{1}R"), function(x)
              grepl(x, b))]
          
          band_aux[[marker_cols[6]]][a] <-
            paste0("L", c("B", "A")[sapply(c("^L", "^U"), function(x)
              grepl(x, b))])
          
          if (cap.data_aux$'1'[a, b] == "00") {
            band_aux[["Banded Leg"]][a] <-
              c("L", "R")[sapply(c(".{1}L", ".{1}R"), function(x)
                grepl(x, b))]
          }
        }
      }
      
    }
    
    for (b in c("TagType1", "TagType2")) {
      if (!is.na(cap.data_aux$'1'[a, b])) {
        n <- n + 1
        
        status <-
          cap.data_aux$'1'[a, paste0("TagStatus", c("1", "2")[sapply(c("1", "2"), function(x)
            grepl(x, b))])]
        
        if (!TRUE %in% sapply(leg_cols, function(x)
          grepl(cap.data_aux$'1'[a, b], cap.data_aux$'1'[a, x])) &
          status != "Removed") {
          band_aux[[colnames(band_aux)[grepl(n, colnames(band_aux))][1]]][a] <-
            unlist(cap.data_aux$'1'[a, b])
          
        }
      }
    }
    
    end.num <-
      c(27, 33, 39, 45, 51, 57)[sapply(c(1, 2, 3, 4, 5, 6), function(x)
        grepl(x, n))]
    
    marker_string <-
      tidyr::replace_na(unlist(band_aux[a, 22:end.num]), "")
    
    for (b in
         names(marker_string)[grepl("Side", names(marker_string))]) {
      marker_string <-
        append(marker_string, c("", ""), after = which(names(marker_string) == b))
    }
    
    band_aux[['All Marker Info At Release']][a] <-
      paste0(paste0(marker_string, collapse = "|"), "|")
  }
  
  band_aux[["Replaced Band Number"]] <-  unlist(cap.data_aux$'1'["BandNumberIn"])
}

## Fill recaptures - aux
if(nrow(cap.data_aux$'R')>1) {
  auto.match.num <-
    which(recap_aux_ref$BBL_col %in% colnames(recap_aux) &
            !is.na(recap_aux_ref$my_col))
  
  for (a in na.omit(auto.match.num)) {
    recap_aux[[recap_aux_ref$BBL_col[a]]] <-
      cap.data_aux$'R'[[recap_aux_ref$my_col[a]]]
  }
  
  
  for (a in 1:nrow(cap.data_aux$'R')) {
    in.num <- 0
    out.num <- 0
    
    for (b in leg_cols) {
      if (!is.na(cap.data_aux$'R'[a, b])) {
        for (c in unlist(strsplit(paste0(cap.data_aux$'R'[a, b]), "/"))) {
          if (grepl("In", b)) {
            in.num <- in.num + 1
            n <- in.num
            marker_cols <-
              colnames(recap_aux)[grepl(n, colnames(recap_aux)) &
                                    grepl("capture", colnames(recap_aux))]
            
          } else if (grepl("Out", b)) {
            out.num <- out.num + 1
            n <- out.num
            marker_cols <-
              colnames(recap_aux)[grepl(n, colnames(recap_aux)) &
                                    grepl("release", colnames(recap_aux))]
          }
          
          recap_aux[[marker_cols[1]]][a] <-
            c
          
          dir <-  c("In", "Out")[sapply(c("In", "Out"), function(x)
            grepl(x, b))]
          
          code_col <-
            paste0("FlagCode", dir)[sapply(c("69"), function(x)
              grepl(x, c))]
          if (length(code_col) != 0) {
            recap_aux[[marker_cols[2]]][a] <-
              cap.data_aux$'R'[[code_col]][a]
          }
          
          color_col <-
            c(paste0("FlagColor", dir), paste0(b, "_colorband"))[sapply(c("69", "01A"), function(x)
              grepl(x, c))]
          if (length(color_col) != 0) {
            recap_aux[[marker_cols[3]]][a] <-
              cap.data_aux$'R'[[color_col]][a]
          }
          
          code_color_col <-
            paste0("FlagCodeColor", dir)[sapply(c("69"), function(x)
              grepl(x, c))]
          if (length(code_color_col) != 0) {
            recap_aux[[marker_cols[4]]][a] <-
              cap.data_aux$'R'[[code_color_col]][a]
          }
          
          recap_aux[[marker_cols[5]]][a] <-
            c("L", "R")[sapply(c(".{1}L", ".{1}R"), function(x)
              grepl(x, b))]
          
          recap_aux[[marker_cols[6]]][a] <-
            paste0("L", c("B", "A")[sapply(c("^L", "^U"), function(x)
              grepl(x, b))])
          
          if (cap.data_aux$'R'[a, b] == "00") {
            recap_aux[["Banded Leg"]][a] <-
              c("L", "R")[sapply(c(".{1}L", ".{1}R"), function(x)
                grepl(x, b))]
          }
          
        }
      }
    }
    
    for (b in c("TagType1", "TagType2")) {
      if (!is.na(cap.data_aux$'R'[a, b]) &
          !TRUE %in% sapply(leg_cols, function(x)
            grepl(cap.data_aux$'R'[a, b], cap.data_aux$'R'[a, x]))) {
        status <-
          cap.data_aux$'R'[a, paste0("TagStatus", c("1", "2")[sapply(c("1", "2"), function(x)
            grepl(x, b))])]
        
        dir <-
          list("Out", "In", c("In", "Out"), c("In", "Out"))[c("Deployed", "Removed", "Replaced", "Retained") %in% status]
        
        for (c in unlist(dir)) {
          direction  <-
            c("capture", "release")[sapply(c("In", "Out"), function(x)
              grepl(x, c))]
          
          if (grepl("In", c)) {
            in.num <- in.num + 1
            n <- in.num
          } else if (grepl("Out", c)) {
            out.num <- out.num + 1
            n <- out.num
          }
          
          recap_aux[[colnames(recap_aux)[grepl(n, colnames(recap_aux)) &
                                           grepl(direction, colnames(recap_aux))][1]]][a] <-
            unlist(cap.data_aux$'R'[a, b])
          
        }
      }
    }
    
    if (n == 0) {
      n <- 1
    }
    
    capture.end.num <-
      c(29, 35, 41, 47, 53, 59)[sapply(c(1, 2, 3, 4, 5, 6), function(x)
        grepl(x, n))]
    
    marker_string <-
      tidyr::replace_na(unlist(recap_aux[a, 24:capture.end.num]), "")
    
    for (b in
         names(marker_string)[grepl("Side", names(marker_string))]) {
      marker_string <-
        append(marker_string, c("", ""), after = which(names(marker_string) == b))
    }
    
    if (paste0(paste0(marker_string, collapse = "|"), "|") == "00||||R|||LB|") {
      recap_aux[a, colnames(recap_aux)[grepl("1", colnames(recap_aux)) &
                                         grepl("capture", colnames(recap_aux))]] <- NA
      
    } else {
      recap_aux[['All Marker Info At Capture']][a] <-
        paste0(paste0(marker_string, collapse = "|"), "|")
    }
    
    
    release.end.num <-
      c(65, 71, 77, 83, 89, 95)[sapply(c(1, 2, 3, 4, 5, 6), function(x)
        grepl(x, n))]
    
    marker_string <-
      tidyr::replace_na(unlist(recap_aux[a, 60:release.end.num]), "")
    
    for (b in
         names(marker_string)[grepl("Side", names(marker_string))]) {
      marker_string <-
        append(marker_string, c("", ""), after = which(names(marker_string) == b))
    }
    
    recap_aux[['All Marker Info At Release']][a] <-
      paste0(paste0(marker_string, collapse = "|"), "|")
    
  }
}

```

#### Export files

```{r}

message(cat(
  paste0(
    "The following data files have been processed and are ready for Bander Portal.\n"
  ),
  paste0(names(data.list), sep = "\n"),
  paste0("\n\nProvide a prefix for the newly written files.\n")
))

prefix <-
  gsub(" ", "_", noquote(as.character(readline(prompt =))))

if (!TRUE %in% sapply(c("_", ".", "-"), function(x)
  endsWith(prefix, x))) {
 prefix<- paste0(prefix,"_")
}

for(a in c("band_no_aux", "band_aux", "recap_no_aux", "recap_aux")) {
  if (nrow(get(a)) > 0) {
  # if (rowSums(is.na(get(a)))[1] != ncol(get(a))) {
    openxlsx::write.xlsx(get(a),
                         paste0(
                           "data-processed/bander_portal/",
                           prefix,
                           a,
                           "_",
                           format(Sys.time(), '%Y%m%d-%H%M%S'),
                           ".xlsx"
                         ))
    
  }
}

message(cat(
  paste0(
    "The files were sucessfully written to '" ,
    "data-processed/bander_portal",
    "'"
  )
))

utils::browseURL("data-processed/bander_portal")

```

# Survey data transformations

### ebird

```{r}
# survey.to.ebird
library("auk")

# ebird
# usfwslandbirds
# fwslandbirds
# API key: e2m7tt6sfo4n


```

# Rename photos

### Provide inputs

```{r}
photo.file.path <- "C:/Users/hvincelette/OneDrive - DOI/Documents/Fieldwork_wrap/2024/Eielson/photos/ALMS_points/CFGesmundo/"

ref.file.path<- "data-processed/to_archive/"

date.col<- "Date"
date.format <- "%m/%d/%Y"
time.col<- "StartTime"
time.format<- "%H:%M"
obs.col<- "ObserverID"
ref.col<- "UnitID"
prefix<- "EIEL"
missing.photo<- c("221","181")

variation <- c("")

```

### Import data

```{r}
# Import ref data file(s)
ref.file <- importer(
  path = ref.file.path,
  multiple = FALSE,
  graphics = TRUE,
  column.format = list(
    c(column = date.col, format = date.format),
    c(column = time.col, format = time.format)
  )
)

```

### Filter data

```{r}

ref.file <- ref.file %>%
  dplyr::mutate({{date.col}} := lubridate::as_date(.data[[date.col]], format = date.format)) %>%
  dplyr::mutate({{time.col}} := lubridate::as_datetime(.data[[time.col]], format = time.format)) %>%
  dplyr::mutate("temp:Year" := format(.data[[date.col]], "%Y"))

year.options <-
  ref.file %>%
  dplyr::select("temp:Year") %>%
  dplyr::filter("temp:Year" != "") %>%
  na.omit() %>%
  purrr::pluck("temp:Year") %>%
  unique() %>%
  sort()

year.choice <- utils::select.list(
  year.options,
  graphics = TRUE,
  multiple = FALSE,
  title = cat(
    paste0("\nSelect a year.\nEnter 0 to proceed with all years.\n")
  )
)
 
if(year.choice !=""){
   ref.file <- ref.file %>% dplyr::filter(`temp:Year` == year.choice)
}

obs.choice <- utils::select.list(
  c(unique(na.omit(ref.file[[obs.col]]))),
  multiple = FALSE,
  graphics = TRUE,
  title = cat(
    paste0(
      "\nSelect an observer.\nNote, photos taken by each observer must be processed individually due to the possibility of concurrent timestamps.\n"
    )
  )
)

if (length(obs.choice) != 0) {
  ref.file <- ref.file %>% dplyr::filter(.[[obs.col]] %in% obs.choice)
}

```

### Formulate names

```{r}

ref.file <- ref.file %>%
  dplyr::select(c({{date.col}}, {{time.col}}, {{ref.col}})) %>%
  dplyr::distinct() %>%
  dplyr::arrange(.[[date.col]], .[[time.col]]) 

ref.order <- ref.file %>%
  purrr::pluck(ref.col) %>%
  sub(paste0(prefix, "_"), "", .) %>%
  subset(!(. %in% missing.photo)) %>%
  rep(., each=length(variation))

variation.vector <- rep(variation, length(fs::dir_ls(path = photo.file.path, type = "file")) / 4)

date.taken <- fs::dir_ls(path = photo.file.path, type = "file") %>%
  exifr::read_exif(tags =
                     c("DateTimeOriginal")) %>%
  dplyr::arrange(DateTimeOriginal) %>%
  dplyr::mutate_at("SourceFile", basename) %>%
  dplyr::mutate(Ext= tools::file_ext(SourceFile)) %>%
  dplyr::mutate_at(
    "DateTimeOriginal",
    ~ lubridate::as_date(., format = "%Y:%m:%d %H:%M:%S") %>% format(., "%Y%m%d")
  )


if (length(ref.order) != nrow(date.taken)) {
  temp <- ref.file %>%
    dplyr::group_by(.[[date.col]]) %>%
    dplyr::summarize(RefTotal = dplyr::n())
  
  message(
    "Images and reference points are not of equal length.\nCheck missing data are accurately reported before continuing.\n"
    
  )
  
  message(huxtable::print_screen(
    huxtable::hux(
      date.taken %>%
        dplyr::group_by(DateTimeOriginal) %>%
        dplyr::summarize(PhotoTotal = dplyr::n() / length(variation)) %>%
        dplyr::mutate(ReferenceTotal = temp$RefTotal)
    ),
    colnames = FALSE
  ))
  
}

if (any(is.na(date.taken$DateTimeOriginal)) == TRUE) {
  browseURL(photo.file.path)
  
  for (a in which(is.na(date.taken$DateTimeOriginal))) {
    pass <- 0
    while (pass == 0) {
      message(cat(
        paste0(
          "\nThe image ",
          date.taken$SourceFile[a],
          " is missing a capture date.\nProvide a date formatted as %Y%m%d (e.g., 20240630).\n"
        )
      ))
      
      new.date <- as.character(readline(prompt = ""))
      
      if (nchar(new.date) != 8 |
          new.date == "" |
          is.na(suppressWarnings(as.numeric(new.date)))) {
        message("Incorrect date format.\n")
      } else {
        date.taken$DateTimeOriginal[date.taken$SourceFile == date.taken$SourceFile[a]] <- new.date
        pass <- 1
      }
    }
  }
}

date.taken <- date.taken %>%
  dplyr::mutate(
    "FileName" = paste0(
      prefix,
      "_",
      ref.order,
      # "_",
      # variation.vector,
      "_",
      date.taken$"DateTimeOriginal",
      ".",
      date.taken$"Ext"
    )
  )

file.rename(
  paste0(photo.file.path, date.taken$SourceFile),
  paste0(photo.file.path, date.taken$FileName)
)

```

# Movebank download

### Provide inputs

Verify the stored login credentials are correct, and add/remove credentials as needed.

```{r warning=FALSE, include=FALSE, results=FALSE}

# Check existing login credentials
keyring::key_list()

# Add login credentials
move2::movebank_store_credentials(username = "mbmlandbirds",
                                  password = "",
                                  key_name = "movebank")

# Remove login credentials
move2::movebank_remove_credentials(key_name = getOption("move2_movebank_key_name"))


```

*Optional: If you want to import Movebank data from a secondary account, paste the associated service name listed in the key list (e.g., "myOtherAccount") and run the following code. Note the account with the service name "movebank" is the default, so you only need to run this code if you want to access a different account. Restart R or execute* `options("move2_movebank_key_name" = "movebank")` *to* *return to the default account. See* `vignette("movebank", package="move2")` *for more details.*

```{r}

keyring::key_list()

options("move2_movebank_key_name" = "movebank")

```

### Import data

Run the following code to import location data from one or more studies into R. Single study imports have the option of selecting a time range of location data, whereas multiple study imports include all location data. By default, the location and reference data are both imported

```{r warning=FALSE, include=FALSE, results=FALSE}
          
# Import Movebank location and/or reference data
import <- get.move.data(
  location = TRUE,
  reference = TRUE,
  remove_movebank_outliers <- FALSE,
  omit_derived_data <- FALSE
)

save(import, file = "reference/movebank/import.RData")

Movebank_dictionary <- readxl::read_excel("reference/movebank/Migratory_Tracking_data_dictionary.xlsx", na = "")

# load("reference/movebank/banding_files.RData")

# download.file(url = "https://www.birdpop.org/docs/misc/IBPAOU.zip", destfile = file.path("reference", "IBPAOU.zip"))
# 
# path <- file.path("reference", "IBPAOU.zip")
# utils::unzip(file.path("reference", "IBPAOU.zip"),
#              exdir = file.path("reference", "IBPAOU"))
# unlink(file.path("reference", "IBPAOU.zip"), recursive = TRUE)

aou_code <- read.csv(list.files("reference/IBPAOU", full.names = TRUE), na = "")

import.dfs <- import

# Unlist dataframes (if necessary)
# if(any(sapply(import, function(x)
#   any(class(x) == "list")))==TRUE) {
#   import.dfs <- rrapply::rrapply(
#     import,
#     classes = "data.frame",
#     how = "flatten",
#     options = list(namesep = "-",
#                    simplify = FALSE)
#   )
# } else if (inherits(import, "data.frame") == TRUE) {
#   import.dfs <- list(import)
# 
# } else{
#   import.dfs <- import
# }


```

### Prepare data

```{r}

# Extract coordinates
for(x in 1:length(import.dfs)) {
  loc.cols <-
    colnames(dplyr::select(import.dfs[[x]], matches("location")))
  for (a in loc.cols) {
    if ("sfc_POINT" %in% class(import.dfs[[x]][[a]])) {
      import.dfs[[x]][[a]] <- sf::st_cast(import.dfs[[x]][[a]], to = "POINT")
            
      names <- names(unlist(import.dfs[[x]][[a]]))
      names<- stringi::stri_remove_empty(names, na_empty = TRUE)
      if (rlang::is_empty(names) == TRUE) {
        names <- c(gsub("location", "long", a), 
                   gsub("location", "lat", a))}
      names <- sapply(names, function(x)
        sub(pattern = "^argos_", replacement = "argos:", x), USE.NAMES = FALSE)
      
      suppressWarnings(
        import.dfs[[x]] <-  import.dfs[[x]] %>%
          tibble::add_column(
            !!names[1] := sf::st_coordinates(import.dfs[[x]][[a]])[, 1],
            !!names[2] := sf::st_coordinates(import.dfs[[x]][[a]])[, 2],
            .before = paste0(a)
          ) %>%
          dplyr::select(!a)
      )
      
    }
  }
}

# Update field names
import.dfs <- import.dfs %>%
  purrr::map(
    .f = ~ .x %>%
      setNames(., gsub(
        pattern = "_", replacement = "-", names(.)
      )) %>%
      setNames(., sub(
        pattern = "^argos-", replacement = "argos:", names(.)
      )) %>%
      data.table::setnames(
        .,
        old = c(
          "individual-id",
          "nick-name",
          "weight",
          "individual-comments",
          "individual-number-of-deployments"
        ),
        new = c(
          "animal-id",
          "nickname",
          "tag-mass",
          "animal-comments",
          "animal-number-of-deployments"
        )
        ,
        skip_absent = TRUE
      )
  )


movebank_attributes<- Movebank_dictionary$codeName[Movebank_dictionary$domainItem_value=="dataField"]

# extract import colnames
all_colnames <- import.dfs %>%
  purrr::map_df(~ names(.x) %>%
                  tibble::enframe(name = NULL)) %>% dplyr::pull(value)

# find dictionary field that matches unknown colnames
new_colnames <-
  sapply(all_colnames[!all_colnames %in% movebank_attributes], function(x) {
    grep(paste0("-", x),
         movebank_attributes,
         value = TRUE)
  }) %>% purrr::compact(.)


# rename unknown colnames
if (length(new_colnames) != 0) {
  import.dfs <- import.dfs %>%
    purrr::map(.f = ~ data.table::setnames(
      .x,
      old = names(new_colnames),
      new = unlist(new_colnames),
      skip_absent = TRUE
    ))
}

  
# Add required columns (if necessary)
all_req_cols <- c(
  "study-name",
  "study-id",
  "deployment-id",
  "tag-id",
  "animal-id",
  "individual-taxon-canonical-name",
  "animal-taxon-detail",
  "animal-ring-id",
  "tag-local-identifier",
  "animal-marker-id"
)

import.dfs <- import.dfs %>%
  purrr::map(.f = ~ .x %>%
               dplyr::mutate(!!!setNames(
                 rep(NA, length(setdiff(
                   all_req_cols, colnames(.x)
                 ))), setdiff(all_req_cols, colnames(.x))
               )) %>% dplyr::relocate(
                 c("animal-ring-id", "animal-marker-id", "tag-local-identifier"),
                 .after = "animal-id"
               ))


# create deploy year column in reference data
import.dfs[grep("reference-data$", names(import.dfs))] <-
  import.dfs[grep("reference-data$", names(import.dfs))] %>%
  purrr::map(
    .f = ~ .x %>%
      dplyr::mutate(
        "deploy-date" = lubridate::as_date(.$"deploy-on-timestamp"),
        .after = "deploy-on-timestamp"
      ) %>%
      dplyr::mutate("deploy-year" = as.character(
        lubridate::year(.$"deploy-on-timestamp"),
        .after = "deploy-date"
      ))
  )

# evaluate data
## missing band numbers
view <-
  import.dfs[["reference-data"]] %>%
  dplyr::filter(is.na(`animal-ring-id`))

## duplicate band numbers
view <-
  import.dfs[["reference-data"]] %>%
  dplyr::filter(!is.na(`animal-ring-id`)) %>%
  dplyr::group_by(`animal-ring-id`) %>%
  dplyr::filter(dplyr::n() > 1)

# duplicate local identifiers and band numbers
view <-
  import.dfs[["reference-data"]] %>%
  dplyr::group_by(`tag-local-identifier`, `animal-ring-id`) %>%
  dplyr::filter(dplyr::n() > 1)

## missing deploy dates
view <-
  import.dfs[["reference-data"]] %>%
  dplyr::filter(is.na(`deploy-date`))

## duplicate studies
# view <-
#   import.dfs[["reference-data"]] %>%
#   dplyr::filter(`individual-taxon-canonical-name`== "Limnodromus griseus") %>%
#   dplyr::group_by(`animal-ring-id`) %>%
#   dplyr::filter(dplyr::n() > 1)
# 
# view <-
#   import.dfs[["reference-data"]] %>%
#   dplyr::filter(`individual-taxon-canonical-name`== "Pluvialis fulva") %>%
#   dplyr::group_by(`animal-ring-id`, `tag-local-identifier`) %>%
#   dplyr::filter(dplyr::n() > 1)


# specific study missing deploy dates
# view <-
#   import.dfs[["reference-data"]] %>%
#   dplyr::filter(`individual-taxon-canonical-name`== "Calidris canutus") %>%
#   dplyr::filter(lubridate::year(`deploy-date`)<min(all_band_data$Year))
#   # dplyr::filter(is.na(`deploy-date`)) %>%
#   # dplyr::filter(dplyr::n() > 1)
             
# view <-
#   import.dfs[["location-data"]] %>%
#   dplyr::filter(
#     `study-name` == "Red Knot (C. canutus roselaari) migratory movements",
#   visible == TRUE) %>%
#   dplyr::arrange(`animal-id`, timestamp) %>%
#   dplyr::group_by(`animal-id`) %>%
#   dplyr::slice(1, dplyr::n())


# Load banding data
missing_values<- c(NA)
species.col <- "SpeciesCode"
date.col <- "Date"
date.format <-  "%m/%d/%Y"
time.col <- "CaptureTime_2400"
time.format <- "%H:%M"
na.if<- c("","-","—",".","NA")

select.import.app(dir = paste0(dirname(rstudioapi::documentPath()),"/reference/movebank/banding_files"),
                  ext = c("txt", "csv", "xls*"),
                  multiple = TRUE,
                  type = "data")

banding.file.list <- importer(
  path = selected.files$path[1],
  files = selected.files$filename,
  column.format = list(
    c(column = date.col, format = date.format),
    c(column = time.col, format = time.format)
  )
)


# Add missing columns
all_band_data <-  data.table::rbindlist(banding.file.list, fill = TRUE)
miss_cols <- setdiff(
  c("FlagCodeOut", "LLOut", "ULOut", "LROut", "UROut"),
  colnames(all_band_data)
)


if (length(miss_cols) > 0) {
  all_band_data <- all_band_data %>%
    dplyr::mutate(!!!setNames(rep(list(as.character(NA)), length(miss_cols)), miss_cols))
}

all_band_data <- all_band_data %>%
  dplyr::mutate_if(is.character,
                   ~ dplyr::na_if(., "")) %>%
  dplyr::mutate_if(is.character, 
                   ~ dplyr::na_if(., "NULL")) %>%
  dplyr::mutate_if(is.character, 
                   ~ dplyr::na_if(., missing_values)) %>%
  dplyr::mutate_at(
    dplyr::vars("BandNumberOut", "Year"), 
    as.character) %>%
  dplyr::mutate_at("Date",
                   ~ format(
                     as.Date(.,format = "%m/%d/%Y"), 
                            "%Y-%m-%d")) %>%
  dplyr::mutate_at("Year", 
                   ~ ifelse(
                     is.na(.) & !is.na("Date"),
                     format(
                       as.Date(Date, format = "%d/%m/%Y"),
                       "%Y"), .)) %>%
  dplyr::mutate_at("Mass_g",  
                   ~ units::set_units(
                     ., "g", mode = "standard")) %>%
  dplyr::mutate_at(c("FieldSex", "CHDSex"), 
                   ~ tolower(.) %>%
                     dplyr::na_if(., "u")) %>%
  dplyr::mutate_at(
    "Age",
    ~ dplyr::recode(
      .,
      HY = "juvenile",
      L = "juvenile",
      AHY = "adult",
      SY = "adult",
      ASY = "adult",
      TY = "adult",
      ATY = "adult"
    ) %>%
      dplyr::na_if(., "U")
  ) %>% 
  dplyr::mutate("AuxMarker" =
                  dplyr::case_when(
                    !is.na(FlagCodeOut) ~ FlagCodeOut,

                    TRUE ~ paste0(
                      stringr::str_replace_na(UROut, replacement = ""),
                      "/",
                      stringr::str_replace_na(LROut, replacement = ""),
                      ":",
                      stringr::str_replace_na(ULOut, replacement = ""),
                      "/",
                      stringr::str_replace_na(LLOut, replacement = "")
                    )
                  )) %>%
  dplyr::inner_join(
    .,
    aou_code[, c("SCINAME", "SPEC")],
    by = c("SpeciesCode" = "SPEC"),
    keep = FALSE,
    relationship = "many-to-many"
  )




# attach(all_band_data)
# unique(Age)
# unique(FieldSex)
# unique(CHDSex)
# which(is.na(Year))
# detach(all_band_data)

view<- all_band_data %>%
  dplyr::filter(is.na(CHDSex))


# Load sex results
select.import.app(dir = paste0(dirname(rstudioapi::documentPath()),"/reference/analysis_results"),
                  ext = c("txt", "csv", "xls*"),
                  multiple = FALSE,
                  type = "data")

sexing_results <- importer(
  path = selected.files$path[1],
  files = selected.files$filename,
  column.format = list(
    c(column = date.col, format = date.format),
    c(column = time.col, format = time.format)
  )
)[[1]]
 
# Join sex results (usually in data postprocessing)
all_band_data <- all_band_data %>%
  dplyr::left_join(
    .,
    dplyr::distinct(sexing_results[, c("BandNumber", "Sex")]) %>%
    dplyr::mutate_at(dplyr::vars("BandNumber", "Sex"), as.character),
    by = c("BandNumberOut" = "BandNumber"),
    keep = FALSE,
    relationship = "many-to-many"
  ) %>%
  dplyr::mutate_at("CHDSex", ~ dplyr::case_when(is.na(.) ~ Sex, TRUE ~ `CHDSex`)) %>%
  dplyr::select(-"Sex") %>%
  dplyr::group_by(Year, BandNumberOut, SCINAME) %>% 
  dplyr::slice_head(n = 1) %>%
  dplyr::ungroup() # remove recpatures within the same year


view <- all_band_data %>%
  dplyr::group_by(
    Year,
    BandNumberOut,
    SCINAME
  ) %>%
  dplyr::summarise(count = dplyr::n()) %>%
  dplyr::filter(count > 1)
    

save(all_band_data, file = "reference/movebank/banding_files.RData")



# Join missing band numbers
na_animal_ring <- import.dfs$`reference-data` %>%
  dplyr::filter(is.na(`animal-ring-id`)) %>%
  dplyr::mutate_at("tag-local-identifier", ~as.character(.)) %>%
  dplyr::left_join(
    .,
    all_band_data[, c("Year", 
                      "BandNumberOut", 
                      "SCINAME", 
                      # "FlagCodeOut", 
                      "TagID1")] %>%
       dplyr::mutate_at("TagID1", ~as.character(.)),
    by = c(
      "individual-taxon-canonical-name" = "SCINAME",
      "deploy-year"  = "Year",
      "tag-local-identifier" = "TagID1"
      # ,
      # "individual-local-identifier" = "FlagCodeOut"
    ),
    keep = FALSE,
    relationship = "many-to-many"
  ) %>%
  dplyr::mutate("animal-ring-id" = `BandNumberOut`
                # ,
                # "animal-marker-id" = `individual-local-identifier`
                ) %>%
  dplyr::mutate_at(
    "animal-ring-id",
    ~ dplyr::case_when(
      is.na(.) ~ `individual-local-identifier`,
      TRUE ~ `animal-ring-id`
    )
  ) %>% 
  dplyr::select(
    "deploy-year",
    "individual-taxon-canonical-name",
    "animal-id",
    "animal-ring-id",
    "tag-local-identifier"
  ) %>%
  dplyr::distinct() %>%
  dplyr::mutate_at(c("animal-id", "animal-ring-id"),
                   ~ as.character(.)) %>%
  dplyr::rename("temp:animal-ring-id" = "animal-ring-id") %>%
  dplyr::distinct()

# awaiting sexing results
view <- all_band_data %>%
  dplyr::filter(is.na(CHDSex), 
                BandNumberOut %in% 
                  BandNumberOut[BandNumberOut %in% sexing_results$BandNumber])

# missing tag status
view<- all_band_data %>%
  dplyr::filter(!is.na(TagID1),
                is.na(TagStatus1))

# associated tag ids missing from imported banding files
view<- import.dfs[["reference-data"]] %>%
  dplyr::filter(!`tag-local-identifier` %in% 
                  c(all_band_data$TagID1, all_band_data$TagID2),
                `animal-ring-id` %in% all_band_data$BandNumberOut)

# view2<- all_band_data %>%
#   dplyr::filter(BandNumberOut == view$`animal-ring-id`[1])
# 
# view$`tag-local-identifier`[1]


# tagged individuals missing from imported banding files
view<- import.dfs[["reference-data"]] %>%
  dplyr::filter(!`tag-local-identifier` %in% 
                  c(all_band_data$TagID1, all_band_data$TagID2),
                !`animal-ring-id` %in% all_band_data$BandNumberOut)


# Create animal id key
ref_req_cols<- 
    c("deploy-year",
      "capture-timestamp",
      "capture-longitude",
      "capture-latitude",
      "animal-sex",
      "animal-life-stage",
      "animal-reproductive-condition",
      "animal-mass",
      "attachment-type",
      "tag-manufacturer-name",
      "tag-model",
      "tag-mass")


suppressWarnings(
animal_key<- import.dfs$`reference-data` %>%
  dplyr::mutate(!!!setNames(
                 rep(NA, length(setdiff(
                   c(all_req_cols,ref_req_cols),
                   colnames(.)
                 ))), 
                 setdiff(c(all_req_cols,ref_req_cols),
                         colnames(.))
               )) %>%
  dplyr::mutate(".id" = seq_along(1:nrow(.))) %>%
  dplyr::mutate_at(dplyr::vars(
          all_req_cols
        ), as.character) %>%
    dplyr::left_join(
    .,
    na_animal_ring[, c("animal-id", "temp:animal-ring-id")],
    by = c(
      "animal-id" = "animal-id"
    ),
    keep = FALSE,
    relationship = "many-to-many"
  ) %>%
   dplyr::mutate_at(
          "animal-ring-id",
          ~ dplyr::case_when(
            is.na(.) ~ `temp:animal-ring-id`,
            TRUE ~ `animal-ring-id`
          )
        ) %>%
  dplyr::left_join(
    .,
    all_band_data[, c("Year", 
                      "BandNumberOut", 
                      "SCINAME", 
                      "AuxMarker",
                      "Latitude_dec",
                      "Longitude_dec",
                      "Age",
                      "CHDSex", 
                      "FieldSex",
                      "Mass_g")],
    by = c(
      "individual-taxon-canonical-name" = "SCINAME",
      "deploy-year"  = "Year",
      "animal-ring-id" = "BandNumberOut"
    ),
    keep = FALSE,
    relationship = "many-to-one"
  ) %>%
  dplyr::mutate_at(
          "animal-marker-id",
          ~ dplyr::case_when(
            is.na(.) ~ AuxMarker,
            TRUE ~ `animal-marker-id`
          )
        ) %>%
  dplyr::mutate_at(
          "animal-marker-id",
          ~ dplyr::case_when(
            !is.na(as.numeric(.)) ~ format(as.numeric(.),
             scientific = FALSE), 
            TRUE ~ `animal-marker-id`)
        ) %>%
   dplyr::mutate_at(
          "animal-mass",
          ~ dplyr::case_when(
            is.na(.) ~ Mass_g,
            TRUE ~ `animal-mass`
          )
        ) %>%
   dplyr::mutate_at(
          "animal-life-stage",
          ~ dplyr::case_when(
            is.na(.) ~ Age,
            TRUE ~ `animal-life-stage`
          )
        ) %>%
  dplyr::mutate_at(
    "animal-reproductive-condition",
    ~ dplyr::case_when(
      `animal-life-stage` %in% c("brood rearing", 
                                 "brooding", 
                                 "incubating") ~ `animal-life-stage`,
      TRUE ~ `animal-reproductive-condition`
    )
  ) %>%
    dplyr::mutate_at(
    "animal-life-stage",
    ~ dplyr::case_when(
      . %in% c("brood rearing",
               "brooding",
               "incubating") ~ "adult",
      TRUE ~ `animal-life-stage`
    )
  ) %>%
   dplyr::mutate_at(
          "animal-sex",
          ~ dplyr::case_when(
            is.na(.) ~ CHDSex,
            TRUE ~ `animal-sex`
          )
        ) %>%
   dplyr::mutate_at(
          "animal-sex",
          ~ dplyr::case_when(
            is.na(.) ~ "u",
            TRUE ~ `animal-sex`
          )
        ) %>%
    dplyr::mutate_at(
    "animal-sex", ~ tolower(.)
  ) %>% 
  dplyr::select(c(tidyselect::any_of(
    c(all_req_cols, ref_req_cols)
  ))) %>%
  dplyr::distinct() %>%
  dplyr::mutate_at(
    "animal-taxon-detail",
    ~  dplyr::case_when(
      `animal-taxon-detail` %in% c("Lesser Yellowlegs",
                                   "LesserYellowlegs") ~ NA,
      TRUE ~ `animal-taxon-detail`
    )
  ) %>%
  dplyr::mutate_at("animal-ring-id", ~stringr::str_remove_all(., "-"))
)
      

# attach(animal_key)
# unique(`animal-sex`)
# unique(`animal-life-stage`)
# unique(`animal-reproductive-condition`)
# detach(animal_key)


# isolate tag deployment/removal events
tag_cols <- colnames(all_band_data)[grep("Tag", colnames(all_band_data))]

tagging_events <- all_band_data %>%
  dplyr::filter(TagID1 %in% animal_key$`tag-local-identifier`) %>%
  dplyr::arrange(BandNumberOut, TagID1, Date) %>%
  tidyr::pivot_longer(
    cols = starts_with("TagID"),
    names_to = "TagNum",
    values_to = "TagID",
    values_drop_na = TRUE
  ) %>%
  dplyr::mutate(
    "TagStatus" = dplyr::case_when(
      TagNum ==  "TagID1"~ TagStatus1, 
      TRUE ~ 
        NA
        # TagStatus2
      ),
    "TagMass_g" = dplyr::case_when(
      TagNum ==  "TagID1"~ TagMass1_g, 
      TRUE ~ 
        NA
        # TagMass2_g
      ),
    "TagAttachment" = dplyr::case_when(
      TagNum ==  "TagID1"~ TagAttachment1, 
      TRUE ~ 
        NA
        # TagAttachment2
      ),
    "TagManufacturer" = dplyr::case_when(
      TagNum ==  "TagID1" ~ TagManufacturer1, 
      TRUE ~ 
        NA
        # TagManufacturer2
      ),
    "TagModel" = dplyr::case_when(
      TagNum ==  "TagID1" ~ TagModel1, 
      TRUE ~ 
        NA
        # TagModel2
      )
  ) %>%
  dplyr::select(-any_of(tag_cols)) %>%
  dplyr::mutate_at("TagStatus",
                   ~ ifelse(. %in% c("Deployed"), 
                            "On", .) %>%
                     ifelse(. %in% c("Removed",
                                     "Replaced"), 
                            "Off", .)) %>%
  dplyr::filter(!TagStatus %in% c("Retained", "Lost", "Recovered")) %>%
  dplyr::select(
    c(
      "SCINAME",
      "Year",
      "Date",
      "BandNumberOut",
      "Longitude_dec",
      "Latitude_dec",
      "BBLDisposition",
      "AuxMarker",
      "TagID",
      "TagStatus",
      "TagAttachment",
      "TagMass_g",
      "TagModel",
      "TagManufacturer"
    )
  ) %>%
  dplyr::mutate_at("TagMass_g", 
                   ~ units::set_units(
                     ., "g", mode = "standard")) %>%
  dplyr::filter(TagStatus == "On")

# attach(tag_events)
# unique(TagModel)
# unique(TagAttachment)
# detach(tag_events)
  
 
# join tagging event info to animal id key
animal_key<- animal_key %>%
  dplyr::left_join(
    .,
    tagging_events %>%  dplyr::mutate_at("TagID", ~as.character(.)),
    by = c("individual-taxon-canonical-name" = "SCINAME",
           "animal-ring-id" = "BandNumberOut",
           "tag-local-identifier" = "TagID",
           "deploy-year" = "Year"),
    keep = FALSE,
    relationship = "many-to-one"
  ) %>%
  dplyr::mutate_at("capture-longitude",
                   ~ dplyr::case_when(
                     is.na(.) ~ Longitude_dec,
                     TRUE ~ `capture-longitude`)) %>%
  dplyr::mutate_at("capture-latitude",
                   ~ dplyr::case_when(
                     is.na(.) ~ Latitude_dec, 
                     TRUE ~ `capture-latitude`)) %>%
  dplyr::mutate_at("capture-timestamp",
                   ~ dplyr::case_when(
                     is.na(.) ~ Date, 
                     TRUE ~ `capture-timestamp`)) %>%
  dplyr::mutate_at("tag-model",
                   ~ dplyr::case_when(
                     is.na(.) ~ TagModel, 
                     TRUE ~ `tag-model`)) %>%
  dplyr::mutate_at("tag-manufacturer-name",
                   ~ dplyr::case_when(
                     is.na(.) ~ TagManufacturer, 
                     TRUE ~ `tag-manufacturer-name`)) %>%
  dplyr::mutate_at("tag-mass",
                   ~ dplyr::case_when(
                     is.na(.) ~ TagMass_g, 
                     TRUE ~ `tag-mass`)) %>%
  dplyr::mutate_at("attachment-type",
                   ~ dplyr::case_when(
                     is.na(.) ~ TagAttachment, 
                     TRUE ~ `attachment-type`)) %>%
  dplyr::mutate_at("tag-model",
                   ~ dplyr::case_when(
                     . %in% c("Argos PinPoint75", 
                              "PinPoint Argos 75",
                              "Pinpoint Argos 75",
                              "PinPoint75") ~ 
                       "Argos PinPoint 75", 
                     TRUE ~ `tag-model`)) %>%
  dplyr::mutate_at("tag-manufacturer-name",
                   ~ dplyr::case_when(
                     . %in% c("Lotek Wireless, Inc.")
                     ~ "Lotek",
                     TRUE ~ `tag-manufacturer-name`)) %>% 
  dplyr::mutate_at("attachment-type",
                   ~ dplyr::case_when(
                     . %in% c("harness",
                              "Leg-Loop") ~ 
                       "leg-loop-harness", 
                     TRUE ~ `attachment-type`)) %>%
  dplyr::select(any_of(c(all_req_cols, ref_req_cols))) %>%
  dplyr::relocate(all_req_cols)
  

# view band records of individuals missing capture lat/long
view <- animal_key %>%
  dplyr::rowwise() %>%
  dplyr::filter(is.na(`capture-latitude`),
                `animal-ring-id` %in%
                  all_band_data$BandNumberOut[!is.na(all_band_data$Latitude_dec) &
                all_band_data$Year == `deploy-year`])



# view2 <-
#   all_band_data %>%
#   dplyr::filter(BandNumberOut == view$BandNumberOut[5])
# 
# view3 <-
#   animal_key %>%
#   dplyr::filter(`animal-ring-id` == view$BandNumberOut[5])

# if tag id appears incorrect..
# view4 <-
#   animal_key %>%
#   dplyr::filter(`tag-local-identifier` %in% c(view3$`tag-local-identifier`,
#                                               view2$TagID1))
  
# attach(animal_key)
# unique(`tag-model`)
# unique(`attachment-type`)
# detach(animal_key)

# duplicate animal ids
view <-
  animal_key %>%
  dplyr::group_by(`animal-id`) %>%
  dplyr::filter(dplyr::n() > 1)

# missing aux markers
view <-
  animal_key %>%
  dplyr::filter(is.na(`animal-marker-id`))

# missing CHDSex
view <-
  animal_key %>%
  dplyr::filter(is.na(`animal-sex`)) # sexually dimorphic can use FieldSex

  
# investigate discrepancies
# bands<- animal_key %>%
#   dplyr::filter(is.na(`animal-marker-id`),
#                 `individual-taxon-canonical-name` == "Tringa flavipes")
# 
# bands$`animal-ring-id`[!bands$`animal-ring-id` %in%  all_band_data$BandNumberOut]
# 
# band_data <-
#   all_band_data %>%
#   dplyr::filter(
#     BandNumberOut == bands$`animal-ring-id`[1] |
#       TagID1 == bands$`tag-loca-identifier`[1]) %>%
#   dplyr::arrange(Year)
# 
# band_data <-
#   all_band_data %>%
#   dplyr::filter(BandNumberOut == "122274480")
# 
# loc_data <-
#   import.dfs[["location-data"]] %>% 
#   dplyr::filter(
#     `tag-local-identifier` == "175320",
#   visible == TRUE) %>%  
#   dplyr::arrange(`animal-id`, timestamp) %>% 
#   dplyr::group_by(`animal-id`) %>%
#   dplyr::slice(1, dplyr::n())

  
# review tag events unassociated with animals (timestamp after deployment ends)
view <- import.dfs[["location-data"]] %>%
  dplyr::filter(is.na(`animal-id`)) %>%
  dplyr::select(c( `study-name`,
                   `tag-local-identifier`, 
                   `timestamp`)) %>%
  dplyr::distinct()

view <- animal_key %>%
  dplyr::left_join(
    .,
    view,
    by = c("tag-local-identifier" = "tag-local-identifier", "study-name" = "study-name"),
    keep = FALSE,
    relationship = "one-to-many"
  ) %>%
  dplyr::filter(!is.na(timestamp)) %>%
  dplyr::mutate_at("capture-timestamp", ~
              as.POSIXct(`capture-timestamp`)) %>%
  dplyr::filter(`capture-timestamp`<timestamp,
                timestamp<Sys.time())
    

# fill animal marker and ring id in location data
suppressWarnings(
  locations <-
    import.dfs[["location-data"]] %>%
    dplyr::mutate(!!!setNames(rep(NA, length(
      setdiff(c(all_req_cols), colnames(.)))), 
      setdiff(c(all_req_cols), colnames(.)))) %>%
    dplyr::mutate(".id" = seq_along(1:nrow(.))) %>%
    dplyr::mutate_at("animal-ring-id", ~ gsub(
      pattern = "-", replacement = "", .)) %>%
    dplyr::mutate_at(dplyr::vars(all_req_cols), as.character) %>%
    dplyr::filter(!is.na(`animal-id`)) %>% # remove unaffliated events
    # dplyr::mutate_at("animal-taxon-detail", ~ "roselaari") %>%
    dplyr::left_join(
      .,
      animal_key %>%
        dplyr::select(all_req_cols) %>%
          dplyr::rename("temp:animal-ring-id" = "animal-ring-id",
                        "temp:animal-marker-id" = "animal-marker-id"),
      by = c("study-name" = "study-name", 
             "study-id" = "study-id", 
             "deployment-id" = "deployment-id", 
             "tag-id" = "tag-id", 
             "tag-local-identifier" = "tag-local-identifier",
             "animal-id" = "animal-id", 
             "individual-taxon-canonical-name" = "individual-taxon-canonical-name", 
             "animal-taxon-detail"  = "animal-taxon-detail"
             ),
      keep = FALSE,
      relationship = "many-to-one"
    ) %>%
    dplyr::mutate_at(
      "animal-ring-id",
      ~ `temp:animal-ring-id`
    ) %>%
    dplyr::mutate_at(
      "animal-marker-id",
      ~ `temp:animal-marker-id`
      
    ) %>%
    dplyr::select(-c(tidyselect::any_of(
      c(
        "deployment-local-identifier",
        "individual-local-identifier",
        ".id",
        grep(
          "temp:",
          colnames(.),
          value = TRUE
        )
      )
    ))) %>%
    dplyr::relocate(all_req_cols) %>%
    dplyr::mutate_at(
    "animal-taxon-detail",
    ~  dplyr::case_when(
      `animal-taxon-detail` %in% c("Lesser Yellowlegs",
                                   "LesserYellowlegs") ~ NA,
      TRUE ~ `animal-taxon-detail`
    )
  )
)

    

# duplicate ids
# view <-
#   locations %>%
#   dplyr::group_by(`.id`) %>%
#   dplyr::filter(dplyr::n() > 1)


# missing required fields
view<- locations %>%
   dplyr::filter(if_any(
     all_req_cols[!all_req_cols %in% c("animal-taxon-detail")], 
                        ~is.na(.x)))


# colnames not in dictioanry
# colnames(animal_key)[!colnames(animal_key) %in% movebank_attributes]
# colnames(locations)[!colnames(locations) %in% movebank_attributes]


```

### Export files

```{r}

# Save dataframes as separate objects 
export_ref <- animal_key %>%
  split(animal_key$`individual-taxon-canonical-name`)

export_loc <- locations %>%
  split(locations$`individual-taxon-canonical-name`)

# if(any(sapply(export, function(x)
#   any(class(x) == "list")))==TRUE) {
#   export.dfs <- rrapply::rrapply(
#     export,
#     classes = "data.frame",
#     how = "flatten",
#     options = list(namesep = "-",
#                    simplify = FALSE)
#   )}



names(export_ref) <- sapply(1:length(export_ref), function(x) {
  paste0(
    gsub(" ", "_", aou_code$COMMONNAME[aou_code$SCINAME %in% names(export_ref[x])]),
    "-reference-data-",
    format(Sys.time(), '%Y%m%d-%H%M%S')
  )
})

names(export_loc) <- sapply(1:length(export_loc), function(x) {
  paste0(
    gsub(" ", "_", aou_code$COMMONNAME[aou_code$SCINAME %in% names(export_loc[x])]),
    "-location-data-",
    min(export_loc[[x]][["event-id"]]),
    "-",
    max(export_loc[[x]][["event-id"]])
  )
})


for (i in seq_along(export_ref)) {
  write.table(
    export_ref[[i]],
    paste0("data-processed/movebank/", names(export_ref[i]), ".txt"),
    na = "",
    row.names = FALSE,
    quote = FALSE,
    sep = "\t"
  )
}


for (i in seq_along(export_loc)) {
  write.table(
    export_loc[[i]],
    paste0("data-processed/movebank/", names(export_loc[i]), ".txt"),
    na = "",
    row.names = FALSE,
    quote = FALSE,
    sep = "\t"
  )
}


# for (i in seq_along(export_ref)) {
#   write.csv(
#     export_ref[[i]],
#     paste0("data-processed/movebank/", names(export_ref[i]), ".csv"),
#     na = "",
#     row.names = FALSE,
#     quote = FALSE
#   )
# }

# for (i in seq_along(export_loc)) {
#   write.csv(
#     export_loc[[i]],
#     paste0("data-processed/movebank/", names(export_loc[i]), ".csv"),
#     na = "",
#     row.names = FALSE,
#     quote = FALSE
#   )
# }
#   

# Get updated movebank vocabulary
move.vocab <-
  move2::movebank_get_vocabulary(xml = "http://vocab.nerc.ac.uk/collection/MVB/current/",
                                 return_type = "list",
                                 omit_deprecated = FALSE)

move.vocab <- rrapply::rrapply(
  move.vocab,
  how = "bind",
  options = list(
    simplify = FALSE,
    coldepth = 3,
    namecols = TRUE
  )
) %>%
  tidyr::pivot_wider(
    names_from = "L2",
    values_from = "1",
    values_fill = NA,
    values_fn = unique
  ) %>%
  tidyr::unnest(colnames(.)) %>%
  dplyr::rename("codeName" = "L1") %>%
  dplyr::mutate_at(c("codeName", "altLabel"), ~ stringr::str_replace_all(., c("argos " = "argos:", " " = "-"))) %>%
  dplyr::mutate(
    "units" =
      stringr::str_match(definition, "Units: \\s*(.*?)\\s*;") %>% .[, 2] %>%
      replace(.,  . %in% c(
        "none", "not defined", "unitless", "not specified"
      ), NA) %>%
      gsub(".*none \\((.*)\\).*", "\\1", .)
  ) %>% dplyr::select("codeName", "definition", "units")

write.csv(move.vocab, "move.vocab.csv", na = "", row.names = FALSE)


#!# remove trailing " " & ";"

# move.vocab <- move.vocab %>%
#   dplyr::mutate_at("definition", ~
#                      stringr::str_replace_all(
#                        .,
#                        c(
#                          "Units: \\s*(.*?)\\s*;" = "",
#                          "Values are chosen from a controlled list: \\s*(.*?)\\s*\\." = "",
#                          "Allowed values are \\s*(.*?)\\s*\\." = ""
#                        )
#                      ))

```
